---
title: "Calculation of fisheries sustainability score"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true

---

```{r message = FALSE}
library(readxl)
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggthemes)
library(scales)
library(zoo)
```

# Importing formatted catches and stocks data
```{r }
ices <- read.csv("data/ices_stock_stats.csv")
catch <- read.csv("data/fish_catch_by_municipality_update2018.csv")
seaarea <-read_excel("/Volumes/ftp.imr.no/Fishery/DeliveryDataFiskDir/KommArea.xlsx")
```

# Update ICES data with the 2018 stock info 
```{r}
bmsy_blim <- ices %>% 
  group_by(Species, Bmsy, Blim) %>% 
  summarize(count = n()) %>% 
  select(-count) %>% arrange(Species) 



ices_2018 <- data.frame(Year = rep(2018, 6),
                        Species = c("Haddock", 
                                    "Cod", 
                                    "Herring", 
                                     "Saithe",
                                    "Capelin",
                                    "Kingcrab"),
                        SSB = c(401586,
                                1504567, 
                                3691000, 
                                454042, 
                                462000, 
                                NA))


ices_2018_prep <- ices_2018 %>%
  arrange(Species) %>%
  mutate(Bmsy = bmsy_blim$Bmsy) %>%
  mutate(SSB.Bmsy = ifelse(Species == "Kingcrab", 1, SSB/Bmsy)) %>%
  mutate(Blim = bmsy_blim$Blim) %>%
  mutate(Spp_year = paste(Species, Year, sep = "_"))

ices_updated <-bind_rows(ices, ices_2018_prep)

#write_csv(ices_updated, "data/ices_updated_2018.csv")
```


# Merging catch data with ICES stocks data
```{r warning=FALSE}
fish_nor <-catch %>% 
           mutate(English = as.character(English)) %>% 
           mutate(English = replace(English, English == "Atlantic cod", "Cod")) %>% 
           mutate(English = replace(English, English == "Atlantic herring", "Herring")) %>% 
           mutate(Spp_year = paste(English,Year, sep = "_")) %>% 
           left_join(ices_updated[,c(3,4,5,6,7)], by = "Spp_year") %>% 
           rename(Species = English)
```

Checking the number of data points per species
```{r}
fish_nor %>% group_by(Species) %>% 
             summarize(n = n())
```


# Further data cleaning: removing and re-arranging locations
Deleting municipalities that are not in the study (Beiarn county and Guovdageaidnu-Kautokeino)
Re-estimating landings for the municipalities that were merged (Bodo + Skjerstad_2005, Bjarkoy_2013 + Harstad_2013) - I sum the landings for these pairs of municipalities for the years before they've merged. 
```{r}
fish_nor_prep <- fish_nor %>%
  mutate(Municip_name = as.character(Municip_name)) %>%
  filter(!Municip_name %in% c("Beiarn", "Guovdageaidnu-Kautokeino")) %>% # mind the correct spelling of Guo... county
  mutate(newid = case_when(
    Municip_name %in% c("Harstad_2013", "Bjarkoy_2013") ~ "1",
    Municip_name %in% c("Bodo", "Skjerstad_2005") ~ "2",
    TRUE ~ Municip_name
  )) %>%
  group_by(newid, Year, Species) %>%
  mutate(Total_catch_new = sum(Total_catch))
```

And now update the whole table, aggregate landings over counties that were merged, remove Skjerstad_2005 and Bjarkoy_2013.
In this table, *Municip_name_new* are the names of the municipalities after they've merged. *Total_catch_new* is the total fish landings after municipalities merged.
```{r}
fish_nor_prep2<- fish_nor_prep %>% 
                  select(-c("Municip_name", "Total_catch")) %>% 
                  mutate(Municip_name_new = newid) %>% 
                  mutate(Municip_name_new = replace(Municip_name_new, Municip_name_new == "1", "Harstad")) %>% 
                  mutate(Municip_name_new = replace(Municip_name_new, Municip_name_new == "2", "Bodo")) %>% 
                  ungroup() %>% 
                  select(c(1,2,11,3:8,10))
```


# Estimating sustainability score for each of the fish stocks
```{r}
fish_nor_estim <-fish_nor_prep2 %>% 
                 mutate(Blim.Bmsy = Blim/Bmsy)

fish_nor_score <-fish_nor_estim %>% #estimating sustainability score for each stock (each species) - variable  "stock_score"
                 mutate(stock_score = case_when(
                                    SSB.Bmsy >= 1 ~ 1,
                                    SSB.Bmsy < 1 & SSB.Bmsy > Blim.Bmsy ~ (SSB.Bmsy - Blim.Bmsy)/Blim.Bmsy,
                                    SSB.Bmsy <= Blim.Bmsy ~ 0
                                    )) 

write_csv(fish_nor_score, "data/sustain_score.csv")
```

# Estimating catch weighted score  
(Check fisheries README. for details)
```{r}
fish_nor_weight <- fish_nor_score %>% #adding total weight of the catch for all stock per municipality and proportion of each stock
                   group_by(Year, Municip_number, Municip_name_new) %>%
                   mutate(total_tons = sum(Total_catch_new)) %>%
                   ungroup() %>%
                   mutate(catch_prop = Total_catch_new/total_tons,
                   catch_weighted_score = catch_prop*stock_score) 

write_csv(fish_nor_weight, "data/catch_weighted_score.csv")


```


