---
title: "Visualizing aquaculture production"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('~/github/aqua_fish_paper/src/common.r')
```

```{r include = FALSE}
library(janitor)
library(ggthemes)
```

# Load data on fisheries catches and aquaculture production per municipality and year
```{r}
fishcatch <- vroom::vroom(file.path("~/github/nor-prep/prep/food_provision/fisheries/data/catch_total.csv"))
aquaprod <-vroom::vroom(file.path("~/github/nor-data/aquaculture/gapfilled_aquaculture_production.csv"))
komnames_norw <- read_excel(file.path("../data/raw/komlist_norw.xlsx"))
```

# Clean the datasets and merge by municipality and year
```{r}
fishcatch_prep <- fishcatch %>% 
  clean_names(.) %>% 
  rename(municip = municip_name) %>% 
  mutate(municip = replace(municip, municip == "Guovdageaidnu-Kautokeino", "Kautokeino")) %>% 
  mutate(municip = replace(municip, municip %in% c("Bjarkoy_2013", "Harstad_2013"), "Harstad")) %>% 
  mutate(municip = replace(municip, municip == "Skjerstad_2005", "Bodo")) %>% 
  mutate(municip_number = replace(municip_number, municip_number %in% c(1915, 1901), 1903)) %>% 
  mutate(municip_number = replace(municip_number, municip_number == 1842, 1804)) %>% 
  group_by(year, municip_number, municip) %>% 
  summarize(catch = sum(total_catch)) 

aquaprod_prep <- aquaprod %>% 
  select(-c(feed_kg, county, gapfilled))

total_prod <- fishcatch_prep %>% 
  filter(year >= 2005) %>% 
  filter(!municip_number == 2011) %>% 
  full_join(aquaprod_prep, by = c("year", "municip_number", "municip")) 
```

Aquaculture data is in kg, so I will divide it by 1000 to make tons
```{r}
convert_to_ton <- function(x){
  tons <- x/1000
  tons
}

totprod_prep <- total_prod %>% 
  mutate_at(vars(production_final), 
            ~ convert_to_ton(.))  %>% 
  rename(aquaprod = production_final) %>% 
  pivot_longer(cols = c(catch, aquaprod), 
               names_to = "production_type",
               values_to = "tons") %>% 
  mutate(county = case_when(
    municip_number >= 1800 & municip_number < 1900 ~ "Nordland",
    municip_number >= 1900 & municip_number < 2000 ~ "Troms",
    municip_number >= 2000  ~ "Finnmark",
  )) 
```

Adding names of municipalities with Norwegian special characters (probably, for manuscript we should use municipalities names as they are in Norwegian)
```{r}
totprod_prep2 <- totprod_prep %>% 
  left_join(komnames_norw[,c("Komnum", "Norw_name")], by = c("municip_number" = "Komnum")) %>% 
  clean_names(.)
```

## Creating ggplot theme 
```{r}
theme_food <- function(...) {
  theme_stata(scheme = "s1color") +
    theme(
      axis.text.x = element_text(
        size = 8, angle = 90, hjust = 1, vjust = 0.5,
        face = "bold"
      ),
      axis.text.y = element_text(size = 10, angle = 0, face = "bold"),
      axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0), face = "bold", size = 10),
      plot.title = element_text(hjust = 0.5, vjust = 0.5, face = "bold"),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2, "cm"),
      strip.text = element_text(size = 10, face = "bold"),
      strip.background = element_rect(fill = "bisque", color ="bisque"),
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      panel.grid.major.y = element_line(color = "grey", size = 0.7),
      legend.background = element_rect(color = "white")
    )
}
```


# Plots
```{r}

breaks1 <- seq(2005, 2020, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1 %% 2 == 0] <- ""


county_names <- list("Finnmark", "Nordland", "Troms")

counties_df <- split(totprod_prep2, totprod_prep2$county)

plots_food <- map2(
  counties_df,
  names(counties_df),
  ~ ggplot(
    data = .x,
    mapping = aes(x = year, y = tons, fill = production_type)
  ) +
    geom_bar(stat = "identity", position = "fill") +
    ggtitle(.y) +
    facet_wrap(norw_name ~ .) +
    scale_x_continuous(breaks = breaks1, labels = labels1) +
    scale_y_continuous(labels = scales::percent) +
    labs(y = "Proportion of seafood production", x = "") +
    theme_food() +
    scale_fill_manual(values = c("mediumblue", "orange3"),  
      name = "" ,               
      labels = c(
        "Aquaculture yield",
        "Fisheries yield"
      ))) 
    


file_names <- paste0("../figs/compare_fish_aqua_prod", "_", county_names, ".pdf")
map2(file_names, plots_food, ggsave)

```