---
title: "Your Document Title"
author: "Document Author"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
---
###Loading libraries
```{r include = FALSE}
library(readxl)
library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggthemes)
library(scales)
library(zoo)
```


###Load the data tables
```{r}
fish_nor_weight <- read.csv("data/catch_weighted_score.csv")
fish_nor_score <-read.csv ("data/sustain_score.csv")
ices_clean <- read.csv("data/ices_stock_stats")
final_scores <-read.csv("data/fishery_final_score.csv")
```

###Overview of fisheries data and scores with ggplots:
Compare total proportions of catch per stock  
```{r, warning=FALSE}
breaks1 <- seq(1994, 2018, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1%%2==0]<- ""

plot_catches <- fish_nor_weight %>% 
                group_by(Year, Species) %>% 
                summarize(catch_spp = sum(Total_catch_new)) %>% 
                group_by(Year) %>% 
                mutate(catch_grandtot = sum(catch_spp)) %>% 
                ungroup() %>% 
                mutate(catch_prop = (catch_spp/catch_grandtot)*100)
                
ggplot(plot_catches, aes(x = Year, y = catch_prop, fill= Species))+
  geom_bar(stat = "identity")+
  scale_x_continuous(breaks = breaks1, labels = labels1)+
  labs(x = "", y = "Total catch proporiton, %")+
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 0.5, face = "bold", colour = "black"),
        axis.text.y = element_text(size = 10, face = "bold",colour = "black"),
        legend.text = element_text(face = "italic"),
        legend.title = element_blank())+
  theme_hc()+
  scale_fill_viridis_d(option = "D") 
```



###Total catch per species and municipality (figures are uploaded as pdf) 
```{r message=FALSE} 
stocks_bycounty <-fish_nor_score %>% 
             mutate(county = case_when(
                                    Municip_number >= 1800 &  Municip_number < 1900 ~ "Nordland",
                                    Municip_number >= 1900 &  Municip_number < 2000 ~ "Troms",
                                    Municip_number >= 2000 ~ "Finnmark")) %>% 
                                    mutate(log_catch = log(Total_catch_new + 1))


counties <- unique(stocks_bycounty$county) 

#redefine breaks if needed, but they are already defined in lines 209 - 211
# breaks1 <- seq(1994, 2018, by = 1)
# labels1 <- as.character(breaks1)
# labels1[!breaks1%%2==0]<- ""

for(i in 1:length(counties))
  {
  cty <- counties[i]
  
  df <- stocks_bycounty %>%
    filter(county == cty)
  
  species <- unique(df$Species)
  
  for(j in 1:length(species)){
  sp <- species[j]
    
  df2 <- df %>% 
    filter(Species == sp)
  
  plot <- ggplot(df2)+
      geom_col(aes(x = Year, y = log_catch), fill = "dodgerblue4")+
      facet_wrap(Municip_name_new  ~ .) +
     # scale_y_continuous(limits = c(0,10))+
      scale_x_continuous(breaks = breaks1, labels = labels1)+
      labs(y = "log(Total catch)", x = "") +
      ggtitle(paste(cty, "_", sp))+
      theme_stata(scheme = "s1color") +
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 6, angle = 90, hjust = 1, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 7, face = "bold", margin = margin(t = 0, r = 5, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))
           
            
ggsave(paste0("figs/catch_", cty, "_", sp, ".pdf"), plot)
  
}
  
}


```

###How SSB per species varied accross years in relation to Blim and Bmsy?
```{r message=FALSE}
#breaks2 <- seq(1994, 2018, by = 1)
# labels1 <- as.character(breaks1)
# labels1[!breaks1%%2==0]<- ""

ices_spp <-unique(ices_clean$Species)

 ices_plotprep <- ices_clean %>% 
       filter(Year == c("1994":"2017")) %>% 
       select(Year, Species, SSB, Bmsy, Blim) %>% 
       mutate(SSB = as.numeric(as.character(SSB))) %>% 
       mutate(Year = as.numeric(as.character(Year)))

 for (i in 1:length(ices_spp)) {
   
 sp <-species[i]  
 df <- ices_plotprep %>% 
       filter(Species == sp)
 bmsy = df$Bmsy[1]
 blim = df$Blim[1]
 
options(scipen = 999) 

plot2 <- ggplot(df)+
      geom_col(aes(x = Year, y = SSB), fill = "dodgerblue")+
      scale_x_continuous(breaks = breaks1, labels = labels1)+
      scale_y_continuous(limits = c(0,max(df$SSB, na.rm = T)))+
      labs(y = "SSB", x = "") +
      geom_hline(aes(yintercept = Blim), col = "darkred", linetype = "dashed", size = 1) +
      geom_hline(aes(yintercept = Bmsy), col = "darkgreen", linetype = "dashed", size = 1) +
      ggtitle(sp) +
      annotate("text", x = 1994, y = (bmsy + 20000), label = "Bmsy", hjust = 0, color = "darkgreen", size = 7)+
      annotate("text", x = 1994, y = (blim + 20000), label = "Blim", hjust = 0, color = "darkred", size = 7)+
      theme_stata(scheme = "s1color") +
      theme(legend.position = "none", 
             axis.title = element_text(size = 15),
             axis.text.x = element_text(size = 12, angle = 45, hjust = 1, vjust = 0.5, face = "bold"),
             axis.text.y = element_text(size = 12, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 0),
                                        angle = 0, hjust = 1, vjust = 0),
             plot.title = element_text(hjust=0.5, vjust=0.5),
             axis.ticks.x = element_line(size = 0.5),
             axis.ticks.length = unit(0.2, "cm"),
             strip.text = element_text(size = 15))  

ggsave(paste0("figs/SSB_", "_", sp, ".pdf"), plot2)      
        
 } 

```


###How did the fisheries sustainability score varied in Norway temporally in each municipality?
```{r message=FALSE}
plot_fscores <-final_scores %>% 
             mutate(county = case_when(
                                    Municip_number >= 1800 &  Municip_number < 1900 ~ "Nordland",
                                    Municip_number >= 1900 &  Municip_number < 2000 ~ "Troms",
                                    Municip_number >= 2000 ~ "Finnmark"
                                    ))

for(i in 1:length(counties))
  {
  cty <- counties[i]
  
  df <- plot_fscores %>%
    filter(county == cty)
  
  plot <- ggplot(df)+
      geom_line(aes(x = Year, y = fish_score))+
      facet_wrap(Municip_name_new  ~ .) +
      scale_x_continuous(breaks = breaks1, limits = c(1994,2018), labels = labels1)+
      labs(y = "Fisheries sustainability score", x = "") +
      ggtitle(cty)+
      theme_stata(scheme = "s1color") +
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 6, angle = 90, hjust = 1, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 7, face = "bold", angle = 0, hjust = 1, vjust = 0),
            axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.1, "cm"))
           
           
ggsave(paste0("figs/fishery_score","_", cty, ".pdf"), plot)
  
}

```

###Re-estimating fisheries sustainability score by smoothing it with a rolling average (zoo package)
```{r}
municipalities <-unique(final_scores$Municip_name_new)

finallist <-list()
for (i in 1:length(municipalities)) 
  {
  mp <- municipalities[i]
  df <- final_scores %>% 
      filter(Municip_name_new == mp)
   runmean <-data.frame(runmean = rollapply(df$fish_score, width = 3, FUN = mean,  fill=NA))
   df2 <- bind_cols(df, runmean)
  finallist[[i]] <- df2  
}

smoothed_fish_scores <- do.call("rbind", finallist)
colSums(is.na(smoothed_fish_scores)) #a lot of NA were introduced with a rolling mean, because some municipailities have too few data lines. Does it make sense to use rolling 
```

