---
title: "Mapping fisheries and aquaculture production"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, 
    out.width = "100%",
    dpi = 300,
    fig.width = 8,
    fig.height = 6,
    strip.white = T)
source('~/github/aqua_fish_paper/src/common.r')
```

```{r}
library(mapdata)
library(maptools)
library(rgdal)
library(rgeos)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
```


# Loading data
## Municipalities shape files
```{r}
allkommuner <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
komlist <-read.csv("/Users/marinaespinasse/github/nor-prep/prep/administrative/komlist.csv", sep = ";")
```

# Fisheries annual production data
Loading production data for fisheries and also Norwegian names of municipalities (with special Norwegian characters)
```{r}
fishcatch <- vroom::vroom(file.path("~/github/nor-prep/prep/food_provision/fisheries/data/catch_total.csv"))

komnames_norw <- read_excel(file.path("../data/raw/komlist_norw.xlsx"))
```

# Preparing mapping layers 
```{r }
komlist <-filter(komlist, !Name == "NonCoast")
ohikom <-allkommuner[allkommuner$KOMM %in% komlist$Komnum,] 
```

```{r }
world <- ne_countries(scale = "medium", returnclass = "sf")
kommuner.pol <- spTransform(ohikom, CRS("+proj=longlat +ellps=GRS80"))
kommunersf <- sf::st_as_sf(ohikom, coords= c("long", "lat"),crs = 4326, agr="identity")
```


```{r}
library(sf)
norway <-world[world$name_long %in% c("Norway", "Sweden", "Finland", "Russian deferation"),]
norway <-st_transform(norway,crs = 32633)
st_crs(norway)
```

## Preparing the data for plotting
I first calculate the total catch of all species per year and municipality. Then I join the resultant table with a polygon geometries for all municipalities.
```{r }
fishcatch05 <-filter(fishcatch, Year == "2005")

plot_fishcatch05_prep <- fishcatch05 %>% 
  group_by(Year, Municip_number,Municip_name) %>% 
  summarize(catch_all_sp = sum(Total_catch)) 


plot_fishcatch05_prep2 <- left_join(
  plot_fishcatch05_prep,
  kommunersf, 
  by = c("Municip_number"="KOMM")) %>% 
  st_as_sf(.)
```

In 2005, 13 municipalities did not have any catch registered (no info on their catches in the table). 


Create intervals of total fisheries catch, it would be easier to visualize then. These intervals I apply to the values of total catch, such as category 0-4,000 ton, or 4,000-8,000 ton.
```{r}

intervals05 <-c(0, 4000, 8000, 12000, 16000, 20000,24000, 28000,46000)#intervals for year 2005 only

labels2 <- imap_chr(intervals05, function(., id){
  return(paste0(formatC(intervals05[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals05[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels2 <- labels2[1:length(labels2) - 1]


plot_fishcatch05_prep3  <- plot_fishcatch05_prep2 %>% 
  filter(!is.na(catch_all_sp)) %>% 
  mutate(catch_quantiles = cut(catch_all_sp,
                               breaks = intervals05,
                               labels = labels2,
                               include.lowest = T))
```


## Map for year 2005
```{r}
base_map <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "gray", 
   size = 0.2) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_fishcatch05_prep3,
    mapping = aes(fill = catch_quantiles),
    color = "white",
    size = 0.2
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_viridis_d(
    option = "magma",
    name = "Fisheries catch ranges, tons",
    alpha = 0.7, # make fill a bit brighter
    begin = 0.1, end = 0.9,
    direction = 1, # dark is lowest, yellow is highest
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top",
      reverse = T
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2005"
  ) 


```

# Adding municipalities borders on land
```{r, dpi = 250, fig.height=8, fig.width=8}
grDevices::cairo_pdf("../figs/fisheries_catch_2005.pdf", width = 10, height = 10)
main_map <- base_map +geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4)
main_map
```
ggsave did not work well - dashes in the legend disapper
```{r}
ggsave("../figs/fisheries_catch_2005.pdf", main_map)
```
## Map for year 2018
I do all the same manipulations for catch in yaer 2018 as for year 2005. I calculate total catch of all species,
and then join the catches table with a geometry of municipalities.
```{r }
fishcatch18 <-filter(fishcatch, Year == "2018")

plot_fishcatch18_prep <- fishcatch18 %>% 
  group_by(Year, Municip_number,Municip_name) %>% 
  summarize(catch_all_sp = sum(Total_catch)) 


plot_fishcatch18_prep2 <- left_join(
  plot_fishcatch18_prep,
  kommunersf, 
  by = c("Municip_number"="KOMM")) %>% 
  st_as_sf(.)
```
Check, for which municipalities did not have registered catch in 2018. None, all municipalities had catch registered in 2018.
```{r}
filter(plot_fishcatch18_prep2, is.na(catch_all_sp)) %>% pull("NAVN")
```

Again, creating intervals for 2018 total catch and creating labels for the intervals.
```{r}

intervals18 <-c(70000, 1000000, 25000000, 55000000,115000000,145000000, 230000000,
                330000000,740000000)#intervals for year 2018 only

labels3 <- imap_chr(intervals18, function(., id){
  return(paste0(formatC(intervals18[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals18[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels3 <- labels3[1:length(labels3) - 1]

plot_fishcatch18_prep3  <- plot_fishcatch18_prep2 %>% 
  filter(!is.na(catch_all_sp)) %>% 
  mutate(catch_quantiles = cut(catch_all_sp,
                               breaks = intervals18,
                               labels = labels3,
                               include.lowest = T))
```

The plot for year 2018
```{r}
base_map2 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "gray", 
   size = 0.2) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_fishcatch18_prep3,
    mapping = aes(fill = catch_quantiles),
    color = "white",
    size = 0.2
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_alpha(
    name = "",
    range = c(0.6, 0),
    guide = F
  ) +
  scale_fill_viridis_d(
    option = "magma",
    name = "Fisheries catch ranges, tons",
    alpha = 0.9, # make fill a bit brighter
    begin = 0.1, end = 0.9,
    direction = 1, # dark is lowest, yellow is highest
    guide = guide_legend(
      keyheight = unit(10, units = "mm"),
      title.position = "top",
      reverse = T
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2018"
  ) 

```

```{r, dpi = 250, fig.height=8, fig.width=8}
grDevices::cairo_pdf("../figs/fisheries_catch_2018.pdf", width = 10, height = 10)
main_map2 <- base_map2 +geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4)
main_map2
```

# Same maps but with different colors,  with municipalities's numbers and modified breaks
## 2005 maps
Likely also, it is better that both maps has similar number of intervals for legend, like 6.
I also divide the catches by 1000, to label them as thousand tons on the map.


```{r}

intervals05 <-c(0, 8, 16, 24, 32,40,46)#intervals for year 2005 only

labels2 <- imap_chr(intervals05, function(., id){
  return(paste0(formatC(intervals05[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals05[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels2 <- labels2[1:length(labels2) - 1]


plot_fishcatch05_prep4  <- plot_fishcatch05_prep2 %>% 
  filter(!is.na(catch_all_sp)) %>% 
  mutate(catch_all_sp_ths = catch_all_sp/1000) %>% 
  mutate(catch_quantiles = cut(catch_all_sp_ths,
                               breaks = intervals05,
                               labels = labels2,
                               include.lowest = T)) 


```
For adding municipalities names, we need centroids of each municipality's polygon.
I will also remove the muncipalities 1920 and 1919 because they are too small, and numbers don't fit, i'll add number manually afterwards.
```{r}
kom.names05 <- cbind(plot_fishcatch05_prep2,st_coordinates(st_centroid(plot_fishcatch05_prep2)))
kom.names_prep05 <- filter(kom.names05, !Municip_number %in% c(1919, 1920))
```


Add transparency to map colors, as suggested here: https://gotellilab.github.io/GotelliLabMeetingHacks/NickGotelli/ColorPalettes.html

```{r}

addalpha <- function(colors, alpha=0.7) {
  r <- col2rgb(colors, alpha=T)
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  return(rgb(r[1,], r[2,], r[3,], r[4,]))
}

```

```{r}
cols3 <-addalpha(brewer.pal(6,"Greens"))
```


```{r}
base_map3 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_fishcatch05_prep4,
    mapping = aes(fill = catch_quantiles),
    color = "white",
    size = 0.2
  ) +
  
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols3,
    name = "Fisheries catch ranges, 1000 tons",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top"
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2005"
  ) 

```

```{r, dpi = 250, fig.height=8, fig.width=8}
grDevices::cairo_pdf("../figs/fisheries_catch_2005.pdf", width = 10, height = 10)
main_map3 <- base_map3 +geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4) +geom_sf_text(data = kom.names_prep05, aes(X,Y, label= Municip_number),fontface = "bold", size = 2) 
main_map3
```
## 2018 map version2
Similarly, prepare municipalities numbers - only for those, that had registered catches in 2018.
```{r}
kom.names18 <- cbind(plot_fishcatch18_prep2,st_coordinates(st_centroid(plot_fishcatch18_prep2)))
kom.names_prep18 <- filter(kom.names18, !Municip_number %in% c(1919, 1920))
```

```{r}

intervals18 <-c(70, 125000,250000, 375000,500000,625000,740000)#intervals for year 2018 only

labels3 <- imap_chr(intervals18, function(., id){
  return(paste0(formatC(intervals18[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals18[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels3 <- labels3[1:length(labels3) - 1]

plot_fishcatch18_prep4  <- plot_fishcatch18_prep2 %>% 
  filter(!is.na(catch_all_sp)) %>% 
   mutate(catch_all_sp_ths = catch_all_sp/1000) %>% 
  mutate(catch_quantiles = cut(catch_all_sp_ths,
                               breaks = intervals18,
                               labels = labels3,
                               include.lowest = T))
```

```{r}
base_map4 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_fishcatch18_prep4,
    mapping = aes(fill = catch_quantiles),
    color = "white",
    size = 0.2
  ) +
  
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols3,
    name = "Fisheries catch ranges, 1000 tons",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top"
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2018"
  ) 

```

```{r, dpi = 250, fig.height=8, fig.width=8}
grDevices::cairo_pdf("../figs/fisheries_catch_2018.pdf", width = 10, height = 10)
main_map4 <- base_map4 +geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4) +geom_sf_text(data = kom.names_prep18, aes(X,Y, label= Municip_number),fontface = "bold", size = 2) 
main_map4
```

# Aligning both plots together
```{r}
library(patchwork)
```

```{r fig.height=5, fig.width=4, dpi = 300}
main_map3/main_map4
```




