---
title: "Calculation of fisheries sustainability score"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r message = FALSE}
library(stringr)
library(ggthemes)
library(scales)
library(zoo)
library(ggridges)
library(kableExtra)
```

# Importing formatted catches and stocks data
```{r }
stock_scores <- read.csv("data/stock_scores.csv")
catch <- read.csv("data/catch_total.csv")
seaarea <-read_excel(file.path(dir_M[2], "Fishery/DeliveryDataFiskDir/KommArea.xlsx"))
```

# Merging catch data with ICES stocks data

```{r warning=FALSE}
stock_scores_prep <- stock_scores[,-1]

fish_nor <-catch %>% 
           left_join(stock_scores_prep, 
                  by = c("Year", "English"
                  )) 
```

Checking the number of data points per species
```{r}
fish_nor %>% group_by(English) %>% 
            tally()
```


# Further data cleaning: removing and re-arranging locations 
Deleting municipalities that are not in the study (Beiarn county and Guovdageaidnu-Kautokeino)
Re-estimating landings for the municipalities that were merged (Bodo + Skjerstad_2005, Bjarkoy_2013 + Harstad_2013) - I sum the landings for these pairs of municipalities for the years before they've merged. 
Using this approach to merge cathes will create duplicated rows, I remove them in the chunk after.
```{r}

fish_nor_prep <- fish_nor %>% 
                  rename(Species = English) %>% 
                  mutate(Municip_name = as.character(Municip_name)) %>% 
                  filter(!Municip_name %in% c("Beiarn", "Guovdageaidnu-Kautokeino")) %>% #mind the correct spelling of Guo... county
                  mutate(newid = case_when(Municip_name %in% c("Harstad_2013", "Bjarkoy_2013") ~ "1",
                                           Municip_name %in% c("Bodo", "Skjerstad_2005") ~ "2",
                                           TRUE ~ Municip_name)) %>% 
                  group_by(newid, Year, Species) %>%  
                  mutate(Total_catch_new = sum(Total_catch))
```


In this table, *Municip_name_new* are the names of the municipalities after they've merged. *Total_catch_new* is the total fish landings after municipalities merged.

```{r}
fish_nor_prep2 <- fish_nor_prep %>% 
                  ungroup() %>% 
                  select(-c("Municip_name", "Total_catch")) %>% 
                  mutate(Municip_name_new = newid) %>% 
                  mutate(Municip_name_new = replace(Municip_name_new, Municip_name_new == "1", "Harstad")) %>% 
                  mutate(Municip_number = replace(Municip_number, Municip_name_new == "Harstad", "1903")) %>% 
                  mutate(Municip_name_new = replace(Municip_name_new, Municip_name_new == "2", "Bodo")) %>% 
                  mutate(Municip_number = replace(Municip_number, Municip_name_new == "Bodo", "1804")) %>% 
                  distinct_all() %>% 
                  select(-c(newid,  species_name))
```



# Examining the distribution of observations of catches by species, regardless of municipality
This grapth shows how the number of municipalities, where the catch was delivered, varied fro, 1994 to 2018
```{r}
catch_dist <- ggplot(data = fish_nor_prep2) +
  geom_density_ridges(aes(y = Species, x = Year, fill = Species)) +
  scale_x_continuous(breaks = seq(1994, 2018, 4)) +
  labs(x = "") +
  scale_fill_viridis_d(option = "D", alpha = 0.6) +
  theme_igray()
catch_dist
```

# Estimating cacth per area and maximal weighted catch
(Check fisheries README. for details). This approach however, has a danger of positive  outliers dominating 
the maximal catch per area. Due to multiple reasons, catch will vary in every municipality for each stock from year to year. Thus, a maximal catch (biomass) per area can be just an outlier, a seldomly seen very high catch. Therefore most of the biomass.area/max.biomass.area ratios are very low (75% are less than 0.6).
The largest catch per sea area was observed in Oksnes in 2018, for Atlantic cod.

We do not have MSY-reference points, but we can assume it is sustainable.
https://www.hi.no/hi/temasider/arter/kyst-og-fjordreke 
Similarly, for Kingcrab, I will assume its catches sustainable.

I remove redfish, blue whiting and greenland halibut from the table.
17.02.2020 we decided not to include these three species into the Fisheries sub-goal.

```{r}
seaarea <- seaarea %>% 
          mutate(Komnum = as.character(Komnum))   

weighted_catch_score <- fish_nor_prep2 %>%
  filter(!Species %in%
    c(
      "Redfish",
      "Greenland halibut",
       "Blue whiting"
    )) %>%
  mutate(stock_score = replace(
    stock_score,
    Species == "Shrimp",
    1
  )) %>%
  mutate(stock_score = replace(
    stock_score,
    Species == "Kingcrab",
    1
  )) %>%
  left_join(seaarea[, c("Havflate", "Komnum")], by = c("Municip_number" = "Komnum")) %>%
  rename("Seaarea" = "Havflate") %>%
  mutate(catch_per_area = Total_catch_new / Seaarea) %>%
  group_by(Species) %>%
  mutate(max_catch_area = max(catch_per_area)) %>%
  ungroup() %>%
  mutate(weighted_catch =  catch_per_area / max_catch_area) %>%
  mutate(weighted_catch_score = weighted_catch * stock_score)
#Check the basic stats for weighted biomass score:
mosaic::fav_stats(weighted_catch_score$catch_per_area)

#write.csv(weighted_catch_score, "data/area_weighted_catch_and_catchscore.csv", row.names = F )

```


# Re-estimate weighted catch per area based on smoothed catches

## Apply 3-years rolling mean on catch data
Rolling mean is a remedy against inter-annual variability in catches and against influential outliers. 
I apply 3-year rolliing mean to catch data. However,rolling mean does not help to smooth variability in catches between the municipalities.
```{r}
municipalities <-unique(fish_nor_prep2$Municip_name_new)
species <-unique(fish_nor_prep2$Species)

splist <-list()
finallist <-list()
for (i in 1:length(municipalities)) 
  {
   mp <- municipalities[i]
   df.mp <- fish_nor_prep2 %>% 
      filter(Municip_name_new == mp)
    for (j in 1:length(species))
     {
        sp <- species[j]
        df.sp <- df.mp %>% 
            filter(Species == sp)
  
      runmean <-data.frame(catch_runmean = rollapply(df.sp$Total_catch_new, width = 3, FUN = mean,  fill=NA))
      df.sp.full <- bind_cols(df.sp, runmean)
      splist[[j]] <- df.sp.full
      smoothed_catches_spp <- do.call("rbind", splist)
      finallist[[i]] <-smoothed_catches_spp
    
    }
   
 }
 
smoothed_catches <- do.call("rbind", finallist)
colSums(is.na(smoothed_catches)) 
```

## Re-esitmate weighted catch per area based on smoothed catches
```{r}
smoothed_weighted_catch_score <- smoothed_catches %>%
  filter(!Species %in%
    c(
      "Redfish",
      "Greenland halibut",
      "Blue whiting"
    )) %>%
  mutate(stock_score = replace(
    stock_score,
    Species == "Shrimp",
    1
  )) %>%
  mutate(stock_score = replace(
    stock_score,
    Species == "Kingcrab",
    1
  )) %>% 
left_join(seaarea[, c("Havflate", "Komnum")], by = c("Municip_number" = "Komnum")) %>%
  rename("Seaarea" = "Havflate") %>%
  filter(!is.na(catch_runmean)) %>%
  mutate(catch_per_area = catch_runmean / Seaarea) %>%
  group_by(Species) %>%
  mutate(max_catch_area = max(catch_per_area)) %>%
  ungroup() %>%
  mutate(weighted_catch = catch_per_area / max_catch_area) %>%
  mutate(weighted_catch_score = weighted_catch * stock_score)

#write.csv(smoothed_weighted_catch_score, "data/ smoothed_area_weighted_catch_and_catchscore.csv", row.names = F)

```

# Check the distribution of biomass/area for each stock
## Plot smoothed catches on a boxplot
```{r}
plot2<- ggplot(smoothed_weighted_catch_score) +
  geom_boxplot(aes(Species, catch_per_area), color = "darkgreen", alpha = 0.7, fill = "transparent", outlier.color = "red") +
  labs(x = "") +
  theme(
  axis.text = element_text(angle = 90) 
  )


#And a truncated version
plot3<- ggplot(smoothed_weighted_catch_score) +
  geom_boxplot(aes(Species, catch_per_area), color = "darkgreen", alpha = 0.7, fill = "transparent", outlier.color = "red") +
  scale_y_continuous(limits = c(0,30)) +
  labs(x = "") +
  theme(
    axis.text = element_text(angle = 90)
  )
  
plot2
plot3
  
```

## Summarize smoothed catches in a table
```{r}  
species <-unique(smoothed_weighted_catch_score$Species)

outputdf <- list()
for (i in 1:length(species))
  {
   sp <- species[i]
   spdf <- filter(smoothed_weighted_catch_score, Species == sp)
   stats <- mosaic::fav_stats(spdf$catch_per_area)
   allstats <-cbind(stats, Species = rep(sp, 1))
   outputdf[[i]] <- allstats
}

catch_area_favstats <- do.call("rbind", outputdf)

newtable <- catch_area_favstats %>%  
    mutate(Species = cell_spec(
    Species, color = "white", bold = T,
    background = spec_color(1:8, end = 0.9, option = "A", direction = -1))) %>% 
    mutate(max = round(max, 2)) %>% 
    mutate(max = cell_spec(
    as.numeric(max), "html", 
    color = ifelse(max > 1000, "red", "blue"))) %>% 
    kable(escape = F, align = "c", digits = 4) %>%
    kable_styling(c("striped", "condensed"), full_width = F) %>% 
    column_spec(1:10,bold = T,background = "azure", color = "darkgreen")
newtable
```


# Calculate final fisheries scores based on smoothed weighted catches per area

```{r}
final_scores_smoothed_area <- smoothed_weighted_catch_score %>%
  filter(!is.na(catch_runmean)) %>%
  group_by(Year, Municip_number, Municip_name_new) %>%
  mutate(total_tons = sum(catch_runmean)) %>%
  mutate(catch_prop = catch_runmean / total_tons) %>%
  mutate(fish_score = sum(weighted_catch_score * catch_prop * 100))

```

# Quick-check of the distribution of area-weighted final score 
```{r}
mosaic::fav_stats(final_scores_smoothed_area$fish_score)       
plot<- ggplot(final_scores_smoothed_area) +
        geom_density(aes(fish_score), fill = "blue", color = "navy", alpha = 0.7)
plot 
```

# Using 0.75th quartile of catch per area as a target
```{r}
smoothed_weighted_catch_score_quantile <- smoothed_catches %>%
  filter(!Species %in%
    c(
      "Redfish",
      "Greenland halibut",
      "Blue whiting"
    )) %>%
  mutate(stock_score = replace(
    stock_score,
    Species == "Shrimp",
    1
  )) %>%
  mutate(stock_score = replace(
    stock_score,
    Species == "Kingcrab",
    1
  )) %>%
  left_join(seaarea[, c("Havflate", "Komnum")], by = c("Municip_number" = "Komnum")) %>%
  rename("Seaarea" = "Havflate") %>%
  filter(!is.na(catch_runmean)) %>%
  mutate(catch_per_area = catch_runmean / Seaarea) %>%
  group_by(Species) %>%
  mutate(catch_area_q75 = quantile(catch_per_area, probs = 0.75)) %>%
  ungroup() %>%
  mutate(weighted_catch = ifelse(
    catch_per_area > catch_area_q75, 1, catch_per_area / catch_area_q75
  )) %>%
  mutate(weighted_catch_score = weighted_catch * stock_score)
```

# Recalcualte fisheries score based on 0.75th quartile of catch as a reference
```{r}
final_scores_smoothed_area_q75 <- smoothed_weighted_catch_score_quantile %>%
                  group_by(Year, Municip_number, Municip_name_new) %>% 
                  mutate(total_tons = sum(catch_runmean)) %>%
                  mutate(catch_prop = catch_runmean / total_tons) %>% 
                  summarize(fish_score = sum(weighted_catch_score*catch_prop*100))

#write.csv(final_scores_smoothed_area_q75, "data/fish_score_smoothed_seaarea.csv")
```

# Quick-check the distribution of area-weighted final score based on 0.75th qunatile
```{r}
mosaic::fav_stats(final_scores_smoothed_area_q75$fish_score)       
plot4<- ggplot(final_scores_smoothed_area_q75) +
        geom_density(aes(fish_score), fill = "blue", color = "navy", alpha = 0.7)
plot4 
```