---
title: "Calculating fish stock scores"
author: "Jamie Afflerbach"
date: "2/11/2020"
output: html_document
---

This script shows how to use the F/B lim and MSY metrics to score a stock on a scale from 0 to 1. Importantly, this method does not penalize for underfished stocks (SSB > MSYBtrigger) or underfishing (F < Fmsy). 

If you want to consider those penalties, you can expand the methods. But for now I'll just start with this simple method of calculating stock scores.

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(readxl)
```

Read in data
```{r}
data <- read_excel("~/github/nor-prep/prep/food_provision/Fisheries/data/sample_data_fmsy_bmsy_norway.xlsx")
```

Clean data
```{r}
# Remove columns we don't need

data_clean <- data %>%
  filter(AssessmentYear == 2019) %>% #if we don't filter for this we have duplicate years
  mutate(F = as.numeric(F),
         FMSY = as.numeric(FMSY),
         FLim = as.numeric(FLim)) %>% #not sure why but these came in as character
  select(year = Year, SSB, F, FLim, Blim, FMSY, MSYBtrigger)# %>%
  #pivot_longer(cols = c(b_bmsy, f_fmsy), names_to = "metric", values_to = "value")

```
# Calculate stock scores between 0 and 1

## Scenario 1: no underfishing penalty

In this case since we have both a target (MSYBtrigger) and Blim we don't need to first calculate B/Bmsy, we can use these two reference points to score each stock. If a stock is above MSYBtrigger, it gets a perfect score since we don't apply an overfishing penalty. 

First calculate the B'
```{r}
data_b_score <- data_clean %>%
  mutate(b_score = case_when(
    SSB >= MSYBtrigger ~ 1,
    SSB <= Blim ~ 0,
    SSB > Blim & SSB < MSYBtrigger ~ 1-((MSYBtrigger - SSB)/(MSYBtrigger - Blim))
  ))
```

Second calculate F'

```{r}
data_b_f_score <- data_b_score %>%
  mutate(f_score = case_when(
    F <= FMSY ~ 1,
    F >= FLim ~ 0,
    F < Blim & F > FMSY ~ 1-((FMSY - F)/(FMSY - FLim))
  ))
```

Take a look
```{r}
plot_data <- data_b_f_score %>%
  pivot_longer(cols = c(b_score, f_score), names_to = "metric", values_to = "score")

ggplot(plot_data) +
  geom_line(aes(x = year, y = score, color = metric)) +
  theme_bw() +
  labs(y = "Score")
```

Combine to get your final stock score

```{r}
data_score <- data_b_f_score %>%
  rowwise() %>%
  mutate(stock_score = mean(c(b_score, f_score), na.rm = T))

ggplot(data_score) +
  geom_line(aes(x = year, y = stock_score)) +
  theme_bw() +
  labs(y = "Stock Score") + 
  expand_limits(y = 0)
```

