---
title: "Calculation of fisheries sustainability score"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true

---

```{r }
library(sf) #run this first
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,
    out.width = "100%",
    dpi = 300,
    fig.width = 8,
    fig.height = 6,
    strip.white = T)
source('~/github/nor-prep/prep/src/common.R')
```


```{r}
library(mapdata)
library(maptools)
library(rgdal)
library(maps)
```

# Loading data
## Municipalities shape files
```{r}
fylker <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_Fylker")
allkommuner <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
komlist <-read.csv("../../administrative/komlist.csv", sep = ";")
```

## Fisheries score data and municipalities
```{r }
fishscore <-read.csv("./data/fishery_final_score_smoothed_by_area.csv")#read the relevant file, this is jsut an example
```

# Trial basic maps for all needs
## Get the map of Norway and neighbour countries
```{r }
norway<- map_data("worldHires", c("Norway", "Sweden", "Finland", "USSR")) 
```

## Simple map of Northern Norway with municipalities
```{r fig.width = 6, fig.height=6, dpi=200}
norway.map <- ggplot(norway)  + 
  geom_polygon(aes(x=long, y=lat, group=group),colour="gray35", fill="gray85") + 
  coord_cartesian(xlim = c(10, 32), ylim=c(65, 75)) +
  labs(x = "Longitude",y = "Latitude")+
  theme_bw()+
  theme(axis.text = element_text(face = "bold", size = 15 ), 
        axis.title =  element_text(face = "italic", size = 15, margin = margin(t = 0, r = 0, b = 0, l = 10)))
norway.map
```

## Filtering out municipallities that are not in the project and changing CRS
```{r }
komlist <-filter(komlist, !Name == "NonCoast")
ohikom <-allkommuner[allkommuner$KOMM %in% komlist$Komnum,] 
kommuner.pol <- spTransform(ohikom, CRS("+proj=longlat +ellps=GRS80"))
```

```{r fig.width = 6, fig.height=6, dpi=200 }
map.komm <- norway.map + geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'steelblue',  size = .4)
map.komm
```

# Load additional libraries and shapes
```{r }
library(rnaturalearth)
library(rnaturalearthdata)
```

```{r }
world <- ne_countries(scale = "medium", returnclass = "sf")
kommunersf <- sf::st_as_sf(ohikom, coords= c("long", "lat"),crs = 4326, agr="identity")
```

# Plot year fisheries score for year 1994
```{r }
fishscore95 <-filter(fishscore, Year == "1995")
foo <- left_join(kommunersf, fishscore95, by = c("KOMM" ="Municip_number")) %>% 
  filter(!is.na(Municip_name_new))
plot_fishscore95 <- st_as_sf(foo)
```


##  Map version 1: basic 
(jsut an example of quick and dirty solution)
```{r }
coloredplot3 <- ggplot(data = world) +
  ggtitle("1995") + 
  geom_sf()+
   geom_sf(fill = "antiquewhite1") +
   geom_sf(data = plot_fishscore95, aes(fill = fish_score), color = "black", size = 0.5) + 
  scale_fill_viridis_c(option = "C", alpha = 0.85, begin=0, end = 0.8, name= "Fisheries \n score ")+
   coord_sf(xlim = c(10, 32), ylim = c(65,72), expand = FALSE)+
   labs(x = "", y = "")+
   scale_y_continuous(breaks = 66:72)+
   theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), panel.background = element_rect(fill = "aliceblue"),
        axis.ticks= element_line(size = 1),
        axis.ticks.length = unit(0.2, "cm"),
        axis.text = element_text(size = 15, colour = "black"),
        plot.title = element_text(size = 15, hjust = 0.5, vjust = 0.5),
        legend.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"))
coloredplot3
```

##  Prepare cosmetics for the map version 2
```{r message=F}

no_classes <- 5

# intervals based on quantiles
quantiles <-plot_fishscore %>% 
  filter(!is.na(fish_score)) %>% 
  pull("fish_score") %>%
  quantile(probs = seq(0, 1, length.out = no_classes + 1)) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric


#create custom labels
labels <- imap_chr(quantiles, function(., id){
  return(paste0(round(quantiles[id], 0),
                             "",
                             " – ",
                             round(quantiles[id + 1], 0),
                             ""))
})

labels <- labels[1:length(labels) - 1]

#create new variable in the plot_data dataframe
plot_fishscore95  <- plot_fishscore95 %>% 
  mutate(score_quantiles = cut(fish_score,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T))



################################################################
##Custom labels version2, based on intervals by 14.

labels2 <- imap_chr(intervals, function(., id){
  return(paste0(round(intervals[id], 0),
                             "",
                             " – ",
                             round(intervals[id + 1], 0),
                             ""))
})

labels2 <- labels2[1:length(labels2) - 1]
  
plot_fishscore95_2  <- plot_fishscore95 %>% 
  mutate(score_quantiles = cut(fish_score,
                               breaks = intervals,
                               labels = labels2,
                               include.lowest = T))
```


##  Defining the general map theme (some items will be customized later)
```{r}
theme_map <- function(...) {
  theme_minimal() +
  theme(
    text = element_text(family = "sans", color = "#22211d"),
    # add a subtle grid
    panel.grid.major = element_line(color = "#dbdbd9", size = 0.3),#dbdbd9 - previous grid color
   # panel.grid.minor = element_blank(),
    # background colors
    panel.background = element_rect(fill = "#f5f5f2", color = NA), 
    panel.spacing = unit(c(-.1, 0.2, .2, 0.2), "cm"),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    # borders and margins
    #plot.margin = unit(c(.2, .2, .2, .2), "cm"),
    panel.border = element_blank(),
    # titles
    legend.text = element_text(size = 15, hjust = 0, color = "#4e4d47"),
    #plot.title = element_blank(),
    legend.title = element_text(size = 12)
  )
}
```

## Map version 1 - only study municipalities are present
```{r}
ggplot(data = plot_fishscore95_2) +
  # use the "alpha hack" (as the "fill" aesthetic is already taken)
  scale_alpha(name = "",
              range = c(0.6, 0),
              guide = F) + # suppress legend
  # add main fill aesthetic
  # use thin white stroke for municipality borders
      geom_sf(
         mapping = aes(fill = score_quantiles),
         color = "white",
          size = 0.6) +
  # use the Viridis color scale
      scale_fill_viridis_d(
      option = "magma",
      name = "Fisheries score",
      alpha = 0.8, # make fill a bit brighter
      begin = 0.1, end = 0.9,
      direction = 1, # dark is lowest, yellow is highest
     guide = guide_legend(
     keyheight = unit(10, units = "mm"),
     title.position = "top",
     label.theme = element_text(size = 20),
     reverse = T )) +     # display highest score on top
  # use thicker white stroke for cantonal borders
     geom_sf(
       data = ,
       fill = "transparent",
       color = "white",
       size = 0.5) +
  # add titles
    labs(x = NULL,
         y = NULL,
         title = "Fisheries score")
  # add theme
  theme_map()
```

###  Make projections for world layer and fisheries score layer similar
```{r}
norway <-world[world$name_long %in% c("Norway", "Sweden", "Finland", "Russian deferation"),]
norway <-st_transform(norway,crs = 32633)
st_crs(norway)
```

## Map version 2: add world layer with borders between Norway, Sweden and Finland
```{r}
norway.layer2<- ggplot()  + 
  geom_sf(data = world, fill = "antiquewhite1",color = "gray") +
  #coord_sf(xlim = c(10, 32), ylim = c(65,72))+
  labs(x = "Longitude",y = "Latitude")+
  geom_sf(data = plot_fishscore95_2,
         mapping = aes(fill = score_quantiles),
         color = "white",
          size = 0.4) +
   coord_sf(xlim = c(7, 32), ylim = c(65,72))+
   scale_alpha(name = "",
              range = c(0.6, 0),
              guide = F) + 
      scale_fill_viridis_d(
      option = "magma",
      name = "Fisheries score",
      alpha = 0.8, # make fill a bit brighter
      begin = 0.1, end = 0.9,
      direction = 1, # dark is lowest, yellow is highest
     guide = guide_legend(
     keyheight = unit(10, units = "mm"),
     title.position = "top",
     reverse = T )) +     
    labs(x = NULL,
         y = NULL,
         title = "Fisheries score")
  # add theme
  theme_map()  
 
  
norway.layer2   
```

# Map verion 3: the whole Norway without neighbour countries
```{r}
norway_admin <- read_sf("../../administrative/raw/shapefiles/NO_Fylker/NO_Fylker_pol.shp")
```

```{r}

norway.layer3 <- ggplot()  + 
  geom_sf(data = norway_admin, fill = "antiquewhite1",color = "gray") +
  #coord_sf(xlim = c(10, 32), ylim = c(65,72))+
  labs(x = "Longitude",y = "Latitude")+
  geom_sf(data = plot_fishscore95_2,
         mapping = aes(fill = score_quantiles),
         color = "white",
          size = 0.2) +
   #coord_sf(xlim = c(10, 32), ylim = c(65,72))+
   scale_alpha(name = "",
              range = c(0.6, 0),
              guide = F) + 
      scale_fill_viridis_d(
      option = "magma",
      name = "Fisheries score",
      alpha = 0.8, # make fill a bit brighter
      begin = 0.1, end = 0.9,
      direction = 1, # dark is lowest, yellow is highest
     guide = guide_legend(
     keyheight = unit(5, units = "mm"),
     title.position = "top",
     reverse = T )) +     
    labs(x = NULL,
         y = NULL,
         title = "Fisheries score") 
  # add theme
  theme_map()  
 
  
norway.layer3  
```


# Map verion 4: using world data
```{r}
norway.layer4 <-ggplot() +
    ggtitle("1995") +
    geom_sf(data = world, fill = "antiquewhite1",color = "gray", size = 1) +
    #coord_sf(xlim = c(4, 32), ylim = c(58, 72), expand = FALSE) +
  theme(axis.text = element_text(size = 17),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 15, face = "bold")) +
  geom_sf(data = plot_fishscore95_2,
         mapping = aes(fill = score_quantiles),
         color = "white",
          size = 0.8) +
   coord_sf(xlim = c(8, 32), ylim = c(64,72), expand = FALSE)+
   scale_alpha(name = "",
              range = c(0.6, 0),
              guide = F) + 
      scale_fill_viridis_d(
      option = "magma",
      name = "Fisheries score",
      alpha = 0.75, # make fill a bit brighter
      begin = 0.1, end = 0.9,
      direction = 1, # dark is lowest, yellow is highest
     guide = guide_legend(
     keyheight = unit(7, units = "mm"),
     title.position = "top",
     title.theme = element_text(size = 17, face  = "bold"),
     label.theme = element_text(size = 15),
     reverse = T )) +     
     labs(x = NULL,
         y = NULL)
  # add theme
  theme_map()  
norway.layer4
```


#  Trying 3D tricks to plot the fishereis score
```{r}
library(rayshader)
```

```{r}
plot_gg(coloredplot3, raytrace = TRUE, width = 10, height = 8, scale = 200, windowsize = c(1400, 866), 
         multicore = TRUE, fov = 50, zoom = 0.7, phi = 45, theta = 30,  sunangle = 300)
#render_camera(fov = 50, zoom = 0.9, theta = 50, phi = 45)
render_snapshot()
```


# Map where land is over the filled polygons of municipalities
```{r}
norway.layer5 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "gray", 
   size = 0.2) +
   coord_sf(xlim = c(10, 32), ylim = c(65, 72)
    ) +
  theme(axis.text = element_text(size = 17),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 15, face = "bold")
  ) +
  geom_sf(
    data = plot_fishscore95_2,
    mapping = aes(fill = score_quantiles),
    color = "white",
    size = 0.2
  ) +
  labs(x = "Longitude", y = "Latitude") +
  scale_alpha(
    name = "",
    range = c(0.6, 0),
    guide = F
  ) +
  scale_fill_viridis_d(
    option = "magma",
    name = "Fisheries score",
    alpha = 0.8, # make fill a bit brighter
    begin = 0.1, end = 0.9,
    direction = 1, # dark is lowest, yellow is highest
    guide = guide_legend(
      keyheight = unit(10, units = "mm"),
      title.position = "top",
      reverse = T
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "1995"
  )
# add theme
theme_map()
  
norway.layer6 <- norway.layer5  + geom_sf(data = world, fill = "antiquewhite1", color = "gray") +
   coord_sf(xlim = c(10, 32), ylim = c(65, 72)) 
```

# Adding municipalities borders on land
```{r}
norway.layer7 <- norway.layer6 +geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4)
norway.layer7
  
```

