---
title: "Estimation of stock scores"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r libraries, results="hide"}
library(ggthemes)
library(janitor)
library(DT)
library(readxl)
```

# Loading ICES stocks data
```{r }
ices_stocks_all <-read.csv("data/ices_stocks_ssb.csv")
ices_msy <-read.csv("data/ices_stock_msy.csv")
```

Below I remove some of the stocks, that do not belong to North-East Atlantic. Also, i homogenize stock names and replace MSYBtrigger for saithe with NA (it is not defined, but 0 value is there instead). 
```{r}
msy_prep <- ices_msy %>% 
  filter(!StockKeyLabel %in% c(
    "cap.27.2a514", 	
    "cod.2127.1f14",
    "cod-arct",
    "cod-coas",
    "cod.27.1-2coast",
    "had-arct",
    "sai-arct",
    "smr-arct",
    "ghl-arct"
  )) 
  
```
There is no MSYBtrigger assessment for redfish (except Blim in 2018), greenland halibut, saithe  and capelin.


# Removing unnessesary rows and small-cleaning the ICES stocks data

Remove unused stocks.
I also suspect that the same stock were uploaded to the database under different names, such as cod-arct and cod.27.1-2 have similar stock description (both are NEA cod). Same problem occurred for haddock  (had-arct and had.27.1-2(), saithe (pok.27.1-2 and sai-arct),  redfish (	reg.27.1-2 and smr-arct), and halibut (ghl-arct and  ghl.27.1-2) which are the same stocks.

I observed also that stock stats for mackerel are duplicated, but SSB and F vary slightly between the duplicates. Likely, this is because data came from two databases, or two assessments. I cannot see how to distinguish between the two assessments, so I just keep the first row of the each pair of duplicated rows. The difference in SSB and F values is small and will not affect our analysis. 

```{r}
stocks_prep <- ices_stocks_all %>% 
  filter(!fishstock %in% c(
    "cap.27.2a514", 	
    "cod.2127.1f14",
    "cod-arct",
    "cod-coas",
    "cod.27.1-2coast",
    "had-arct",
    "sai-arct",
    "smr-arct",
    "ghl-arct"
)) %>% 
  select(c(Year, 
           SSB,
           F,
           fishstock,
           SpeciesName,
           StockDescription
  )) %>%
  distinct_at(vars(
    Year,
    fishstock,
    SpeciesName, 
    StockDescription
  ), .keep_all = TRUE
  ) %>% 
  clean_names(.) 
  
```


For MSY-table, keep only the latest year of assessment (usually, 2016-2018) for each species.

```{r}
msy_prep2 <- msy_prep %>%
  group_by(StockKeyLabel) %>% 
  arrange(AssessmentYear) %>% 
  slice(n()) %>% 
  ungroup()
```

# Merging the MSY and SSB tables
```{r}
stock_df_full <- stocks_prep %>% 
  left_join(msy_prep2, 
              by = c("fishstock" = "StockKeyLabel"))
```

For saithe, no MSYBtrigger  and Fmsy are undefined, to replace missing MSYBtrigger, we use SSB trigger  (220000), reported here:
http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/pok.27.1-2.pdf 


For redfish (*S. norvegicus*), ICES advice states that "there is no direct fisheries" in the region (North-East Atlantic), and no MSYBtrigger assessemnt exist. In the northern Norway, the proportion of redfish in total catches in negligible as well.
http://ices.dk/sites/pub/Publication%20Reports/Advice/2018/2018/reg.27.1-2.pdf

For capelin, MSYBtrigger and Fmsy are also undefined. We can instead take the lowest observed SSB since 1994.
http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/cap.27.1-2.pdf


Greenland halibut does not have any stock assessments, but we can assume that since 1994 the stock is sustainable since fisheries was reduced in 1990's.
http://ices.dk/sites/pub/Publication%20Reports/Advice/2019/2019/ghl.27.1-2.pdf

Find a minimum SSB of capeling observed since 1994:
```{r}
stocks_prep %>% 
  filter(fishstock == "cap.27.1-2") %>% 
  filter(year >= 1994) %>% 
  summarise_at(vars(ssb), min, na.rm = T)
  
```

Replacing missing MSYBtrigger for saithe and capelin.
Also, for some species, the mortality (f) is unkown for the recent year (e.g., 2018), I replace NA with the values of the nearest year (e.g., 2017). 

```{r}
stock_df_final <- stock_df_full %>% 
  select(-c(
    AssessmentYear, 
    AssessmentKey, 
    stock_description
            )) %>% 
  filter(year >= 1994 & year < 2019) %>% 
  mutate(MSYBtrigger = replace(
  MSYBtrigger, 
  fishstock == "pok.27.1-2",
  220000
  )) %>% 
 mutate(MSYBtrigger = replace(
  MSYBtrigger, 
  fishstock == "cap.27.1-2",
  94000	
 )) %>% 
  group_by(fishstock) %>% 
  mutate(f_gapfilled = zoo::na.fill(f,"extend")) %>% 
  ungroup()
```

# Calculating stock scores
```{r}
stock_scores_prep <- stock_df_final %>% 
  select(-c(n,f)) %>% 
  rename(Year = year,
         SSB = ssb,
         F = f_gapfilled,
         Flim = FLim,
         Fmsy = FMSY,
         Stock = fishstock
      ) %>% 
  rowwise() %>% 
  mutate(
    b_score = ifelse(
    any(is.na(c(MSYBtrigger, Blim))),
    NA,
    case_when(
    SSB >= MSYBtrigger ~ 1,
    SSB <= Blim ~ 0,
    SSB > Blim & SSB < MSYBtrigger ~ 1-((MSYBtrigger - SSB)/(MSYBtrigger - Blim))
  )
  )) %>% 
     rowwise() %>% 
  mutate(
    f_score = ifelse(
    any(is.na(c(Fmsy, Flim))),
    NA, 
    case_when(
    F <= Fmsy ~ 1,
    F >= Flim ~ 0,
    F < Flim & F > Fmsy ~ 1-((Fmsy - F)/(Fmsy - Flim))
    )
  )) %>% 
  rowwise() %>%
  mutate(stock_score = ifelse(
    all(is.na(c(f_score, b_score))), 
         NA,
        mean(c(b_score, f_score),na.rm = T))
  )
```
  
Adding English names to species
```{r}
stock_scores_final <- stock_scores_prep %>% 
  mutate(English = case_when(
    Stock == "cap.27.1-2" ~ "Capelin",
    Stock == "cod.27.1-2" ~ "Atlantic cod",
    Stock == "ghl.27.1-2" ~ "Greenland halibut",
    Stock == "had.27.1-2" ~ "Haddock",
    Stock == "her.27.1-24a514a" ~ "Atlantic herring",
    Stock == "pok.27.1-2" ~ "Saithe",
    Stock == "reg.27.1-2" ~ "Redfish",
    Stock == "whb.27.1-91214" ~ "Blue whiting",
    Stock == "mac.27.nea" ~ "Atlantic mackerel"
  )) 

write.csv(stock_scores_final, "data/stock_scores.csv")

```


# Take a look at the stock scores
```{r}
ggplot(stock_scores_final) +
  geom_line(aes(x = Year, y = stock_score)) +
  theme_bw() +
  facet_wrap(~ English) +
  expand_limits(y = 0) +
  labs(y = "Stock Score") 
```

  
  
  


