---
title: "Cleaning fish catch data for fisheries goal estimation"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true

---

```{r, message=FALSE}
library(readxl)
library(RCurl)
library(dplyr)
library(stringr)
library(ggplot2)
library(ggthemes)
library(scales)
library(zoo)
```

# Loading ICES fishery parameters data, landings in Northern Norway, and species names table
```{r include = FALSE}
temp_ices <- read_excel("/Volumes/ftp.imr.no/Fishery/DeliveryDataFiskDir/ices_bmsy.xlsx")
temp_catch17 <- read.csv("/Volumes/ftp.imr.no/Fishery/DeliveryDataFiskDir/Fisk0017.csv", sep = ";")
temp_catch99 <- read.csv("/Volumes/ftp.imr.no/Fishery/DeliveryDataFiskDir/Fisk9499.csv", sep = ";")
names<-read.csv("/Volumes/ftp.imr.no/Fishery/DeliveryDataFiskDir/Fiskid.csv", sep = ";")#fish species names in English and Norwegian
komlist<-read.csv("/Volumes/ftp.imr.no/Administrative/komlist.csv", sep = ";")#list of municipalities names and codes
seaarea<-read_excel("/Volumes/ftp.imr.no/Fishery/DeliveryDataFiskDir/KommArea.txt")#fishing area (sea area) 
```

# Data cleaning
Removing unnessesary rows and small-cleaning the ices Bmsy and SSB data
```{r, warning=FALSE}
foo <- temp_ices[-c(1:7),]; colnames(foo) <-temp_ices[7,];colnames(foo)[13] <-"SSB"
ices_clean <- data.frame(Year = rep(foo$Yrs, 6), 
                         Species = rep(c("Cod", "Haddock", "Herring", "Saithe","Capelin", "Kingcrab"), each = 72),
                         SSB = rbind(foo[,2], foo[,5],foo[,7],foo[,9],foo[,11],foo[,13]),
                         Bmsy = rep(c(460000, 80000,3184000,220000,427000,NA), each = 72),
                         SSB_Bmsy = rbind(foo[,3],foo[,6],foo[,8],foo[,10],foo[,12],foo[,14]),
                         Blim = rep(c(220000, 50000,2500000,136000,200000,NA), each = 72) )
ices_fin <- ices_clean %>% 
            mutate(SSB = as.numeric(SSB),SSB.Bmsy = as.numeric(SSB.Bmsy)) %>% 
            mutate(Spp_year = paste(Species,Year, sep = "_"))
```

## Cleaning and formatting fish landings data.
**NB!** Coastal cod (skrei) and atlantic cod (annen torsk) are all under one name "Cod".
I also corrected municipalities names so that they are similar in all data layers, by joining catch tables with komlist table.
```{r include=FALSE}
temp_catch17 <-select(temp_catch17, -c(5,10)) %>% rename("Lengdegr"= "Lengdegr_l.l.")
temp_catch99 <-select(temp_catch99, -5) #remove species code, which is different between 1990s and 2000s
temp_catch<-bind_rows(temp_catch99, temp_catch17)
colnames(temp_catch) <- c("Year", "Month", "Municip_number", "Municip_name","Species",
                          "Coast_ocean", "Region", "Fleet_length", "Catch_weight", "Payed_NOK")
#several norwegian names were used for the same species, let's check all variation
fishnames<-unique(temp_catch$Species)
str_view(fishnames, regex("torsk", ignore_case = TRUE))#Annen Torsk, Nordostarktisk torsk, Torsk, Torsk (oppdrett), Polartorsk
str_view(fishnames, regex("hyse", ignore_case = TRUE))#Hyse, Nordostarktisk hyse , Annen hyse
str_view(fishnames, regex("lodde", ignore_case = TRUE))#Lodde,Barentshavslodde, Lodde - Island/O Gronl./Jan M
str_view(fishnames, regex("sei", ignore_case = TRUE))#Sei, Sei (oppdrett)
str_view(fishnames, regex("sild", ignore_case = TRUE))#Sild, Feitsild, Nordsjosild,Norsk vaargytende sild, Skagerraksild
str_view(fishnames, regex("kongekrabbe", ignore_case = TRUE))#Kongekrabbe, han-, Kongekrabbe, hun-, Kongegrabbe
```

## Finallly creating clean catches data
```{r}
#I assume that we don't need maricultured cod and saithe (oppdrett torsk, oppdrett sild), and capeling from Iceland and Greenalnd, Skagerraksild, Nordsjosild
catch_clean <- temp_catch %>% 
              select(-c(7,10)) %>% #remove columns "Region" and "Payed_NOK", we don't need them for now 
              filter(Coast_ocean == "8") %>% #%>% #select only coastal areas
              filter(!Species %in% c("Torsk (oppdrett)", "Polartorsk", "Lodde - Island/O Gronl./Jan M", 
                                     "Sei (oppdrett)", "Skagerraksild" )) %>% 
              filter(Species %in% c("Annen Torsk","Annen torsk","Nordostarktisk torsk","Skrei","Hyse","Nordostarktisk hyse", 
                                    "Annen hyse", "Sei", "Lodde","Barentshavslodde","Sild","Feitsild", "Kongekrabbe", "Norsk vaargytende sild")) %>% 
              mutate(Species = replace(Species, Species %in% c("Annen Torsk","Annen torsk" , "Nordostarktisk torsk", "Skrei"), "Torsk")) %>% 
              mutate(Species = replace(Species, Species %in% c("Nordostarktisk hyse", "Annen hyse"), "Hyse")) %>% 
              mutate(Species = replace(Species, Species == "Barentshavslodde", "Lodde")) %>% 
              mutate(Species = replace(Species, Species %in% c("Kongekrabbe, hun-", "Kongekrabbe, han-"), "Kongekrabbe")) %>%   
              mutate(Species = replace(Species, Species %in% c("Feitsild", "Norsk vaargytende sild"), "Sild")) %>% 
              left_join(names[,c(4,6)], by = c("Species" = "Common (Norw)")) %>% 
              rename("English" = "Common (Eng)") %>% 
              left_join(komlist[,c(1,2)], c("Municip_number" = "Komnum")) %>% 
              select(-4) %>% 
              rename("Municip_name" = "Name") %>% 
              select(c(1,2,3,9,8,7,6))
              
head(catch_clean, 5)
```
A small check - how many observations per species? Relatively few data points for crab and capelin.
```{r}
 catch_clean %>% count(English)
```

# Aggregating catch data by municipality and fish species. Taking total for catches of all types of fleets and for all months.
Some municipalities names were missing in the list of names (these are municipalities that have merged with other muncipalities)
```{r}
catch_final <- catch_clean %>% 
             mutate(Municip_name = replace(Municip_name, Municip_number == 1839, "Beiarn")) %>% 
             mutate(Municip_name = replace(Municip_name, Municip_number == 1842, "Skjerstad_2005")) %>% 
             mutate(Municip_name = replace(Municip_name, Municip_number == 1901, "Harstad_2013")) %>% 
             mutate(Municip_name = replace(Municip_name, Municip_number == 1915, "Bjarkoy_2013")) %>% 
             mutate(Municip_name = replace(Municip_name, Municip_number == 2011, "Guovdageaidnu-Kautokeino")) %>% 
             group_by(Year,Municip_number,Municip_name,English) %>% 
             summarize(Total_catch = sum(Catch_weight))
```


# Merging catch data with ICES (SSB,Bmsy, Blim) data and adding sea area (in case we'll need it)
```{r}
fish_nor <-catch_final %>% 
           mutate(English = replace(English, English == "Atlantic cod", "Cod")) %>% 
           mutate(English = replace(English, English == "Atlantic herring", "Herring")) %>% 
           mutate(English = replace(English, English == "Red king crab", "Kingcrab")) %>% 
           mutate(Spp_year = paste(English,Year, sep = "_")) %>% 
           left_join(ices_fin[,c(3,4,5,6,7)], by = "Spp_year") %>% 
           left_join(seaarea[,c(2,6)], by = c("Municip_name"= "Name")) %>% 
           rename("Species" = "English", "Seaarea"= "Havflate") 
```

Checking the number of data points per species
```{r}
fish_nor %>% group_by(Species) %>% 
             summarize(n = n())
```
Confirming that there are 5 extra municipalities in the landings data:
```{r}
extra_mun<-anti_join(fish_nor[,3], komlist[,2], by=c("Municip_name"="Name"))
unique(extra_mun)
```

###Issues with the data:
Confirming that there are 5 extra municipalities in the landings data:
```{r}
extra_mun<-anti_join(fish_nor[,3], komlist[,2], by=c("Municip_name"="Name"))
unique(extra_mun)
```

###Further data cleaning: removing locations and re-arranging locaitons
Deleting municipalities that are not in the study (Beiarn county and Guovdageaidnu-Kautokeino)
Re-estimating landings for the municipalities that were merged (Bodo + Skjerstad_2005, Bjarkoy_2013 + Harstad_2013) - I sum the landings for these pairs of municipalities before they have merged. 
```{r}
fish_nor_prep <- fish_nor %>% 
                  filter(!Municip_name %in% c("Beiarn", "Guovdageaidnu-Kautokeino")) %>% #mind the correct spelling of Guo... county
                  mutate(newid = case_when(Municip_name %in% c("Harstad_2013", "Bjarkoy_2013") ~ "1",
                                           Municip_name %in% c("Bodo", "Skjerstad_2005") ~ "2",
                                           TRUE ~ Municip_name)) %>% 
                  group_by(newid, Year, Species) %>% 
                  mutate(Total_catch_new = sum(Total_catch))
```

And now update the whole table, where municipalities Bjarkoy_2013 and Skjerstad_2005 do not exist.
In this table, Municip_name_new are the names of the municipalities after they've merged. Total_catch_new are the total fish landings after municipalities merging.
```{r}
fish_nor_prep2<- fish_nor_prep %>% 
                  select(-c("Municip_name", "Total_catch")) %>% 
                  mutate(Municip_name_new = newid) %>% 
                  mutate(Municip_name_new = replace(Municip_name_new, Municip_name_new == "1", "Harstad")) %>% 
                  mutate(Municip_name_new = replace(Municip_name_new, Municip_name_new == "2", "Bodo")) %>% 
                  ungroup() %>% 
                  select(c(1,2,12,3:9,11))
```

###Estimating sustainability score for each of the fish stocks
```{r}
fish_nor_estim <-fish_nor_prep2 %>% 
                 mutate(Blim.Bmsy = Blim/Bmsy)

fish_nor_score <-fish_nor_estim %>% #estimating sustainability score for each stock (each species) - variable stock_score
                 mutate(stock_score = case_when(
                                    SSB.Bmsy >= 1 ~ 1,
                                    SSB.Bmsy < 1 & SSB.Bmsy > Blim.Bmsy ~ (SSB.Bmsy - Blim.Bmsy)/Blim.Bmsy,
                                    SSB.Bmsy <= Blim.Bmsy ~ 0
                                    ))

#write.csv(fish_nor_score, "sustain_score.csv")
```

###Estimating sustainability of fisheries in Northern Norway as score*proportion of catch
```{r}
fish_nor_weight <- fish_nor_score %>% #adding total weight of the catch for all stock per municipality and proportion of each stock
                   group_by(Year, Municip_number, Municip_name_new) %>%
                   mutate(total_tons = sum(Total_catch_new)) %>%
                   ungroup() %>%
                   mutate(catch_prop = Total_catch_new/total_tons,
                   catch_weighted_score = catch_prop*stock_score)

#write.csv(fish_nor_weight, "catch_weighted_score.csv")

final_scores <- fish_nor_weight %>%
                  group_by(Year, Municip_number, Municip_name_new) %>%
                  summarize(fish_score = sum(catch_weighted_score)*100)

#write.csv(final_scores, "fishery_final_score.csv")

```

###Overview of results with ggplots:
Compare total proportions of catch per stock  
```{r, warning=FALSE}
breaks1 <- seq(1994, 2018, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1%%2==0]<- ""

plot_catches <- fish_nor_weight %>% 
                group_by(Year, Species) %>% 
                summarize(catch_spp = sum(Total_catch_new)) %>% 
                group_by(Year) %>% 
                mutate(catch_grandtot = sum(catch_spp)) %>% 
                ungroup() %>% 
                mutate(catch_prop = (catch_spp/catch_grandtot)*100)
                
ggplot(plot_catches, aes(x = Year, y = catch_prop, fill= Species))+
  geom_bar(stat = "identity")+
  scale_x_continuous(breaks = breaks1, labels = labels1)+
  labs(x = "", y = "Total catch proporiton, %")+
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1, vjust = 0.5, face = "bold", colour = "black"),
        axis.text.y = element_text(size = 10, face = "bold",colour = "black"),
        legend.text = element_text(face = "italic"),
        legend.title = element_blank())+
  theme_hc()+
  scale_fill_viridis_d(option = "D") 
```





###Total catch per species and municipality (figures are uploaded as pdf) 
```{r message=FALSE} 
stocks_bycounty <-fish_nor_score %>% 
             mutate(county = case_when(
                                    Municip_number >= 1800 &  Municip_number < 1900 ~ "Nordland",
                                    Municip_number >= 1900 &  Municip_number < 2000 ~ "Troms",
                                    Municip_number >= 2000 ~ "Finnmark")) %>% 
                                    mutate(log_catch = log(Total_catch_new + 1))


counties <- unique(stocks_bycounty$county) 

#redefine breaks if needed, but they are already defined in lines 209 - 211
# breaks1 <- seq(1994, 2018, by = 1)
# labels1 <- as.character(breaks1)
# labels1[!breaks1%%2==0]<- ""

for(i in 1:length(counties))
  {
  cty <- counties[i]
  
  df <- stocks_bycounty %>%
    filter(county == cty)
  
  species <- unique(df$Species)
  
  for(j in 1:length(species)){
  sp <- species[j]
    
  df2 <- df %>% 
    filter(Species == sp)
  
  plot <- ggplot(df2)+
      geom_col(aes(x = Year, y = log_catch), fill = "dodgerblue4")+
      facet_wrap(Municip_name_new  ~ .) +
     # scale_y_continuous(limits = c(0,10))+
      scale_x_continuous(breaks = breaks1, labels = labels1)+
      labs(y = "log(Total catch)", x = "") +
      ggtitle(paste(cty, "_", sp))+
      theme_stata(scheme = "s1color") +
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 6, angle = 90, hjust = 1, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 7, face = "bold", margin = margin(t = 0, r = 5, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))
           
            
ggsave(paste0("figs/catch_", cty, "_", sp, ".pdf"), plot)
  
}
  
}


```

###How SSB per species varied accross years in relation to Blim and Bmsy?
```{r message=FALSE}
#breaks2 <- seq(1994, 2018, by = 1)
# labels1 <- as.character(breaks1)
# labels1[!breaks1%%2==0]<- ""

ices_spp <-unique(ices_clean$Species)

 ices_plotprep <- ices_clean %>% 
       filter(Year == c("1994":"2017")) %>% 
       select(Year, Species, SSB, Bmsy, Blim) %>% 
       mutate(SSB = as.numeric(as.character(SSB))) %>% 
       mutate(Year = as.numeric(as.character(Year)))

 for (i in 1:length(ices_spp)) {
   
 sp <-species[i]  
 df <- ices_plotprep %>% 
       filter(Species == sp)
 bmsy = df$Bmsy[1]
 blim = df$Blim[1]
 
options(scipen = 999) 

plot2 <- ggplot(df)+
      geom_col(aes(x = Year, y = SSB), fill = "dodgerblue")+
      scale_x_continuous(breaks = breaks1, labels = labels1)+
      scale_y_continuous(limits = c(0,max(df$SSB, na.rm = T)))+
      labs(y = "SSB", x = "") +
      geom_hline(aes(yintercept = Blim), col = "darkred", linetype = "dashed", size = 1) +
      geom_hline(aes(yintercept = Bmsy), col = "darkgreen", linetype = "dashed", size = 1) +
      ggtitle(sp) +
      annotate("text", x = 1994, y = (bmsy + 20000), label = "Bmsy", hjust = 0, color = "darkgreen", size = 7)+
      annotate("text", x = 1994, y = (blim + 20000), label = "Blim", hjust = 0, color = "darkred", size = 7)+
      theme_stata(scheme = "s1color") +
      theme(legend.position = "none", 
             axis.title = element_text(size = 15),
             axis.text.x = element_text(size = 12, angle = 45, hjust = 1, vjust = 0.5, face = "bold"),
             axis.text.y = element_text(size = 12, face = "bold", margin = margin(t = 0, r = 0, b = 0, l = 0),
                                        angle = 0, hjust = 1, vjust = 0),
             plot.title = element_text(hjust=0.5, vjust=0.5),
             axis.ticks.x = element_line(size = 0.5),
             axis.ticks.length = unit(0.2, "cm"),
             strip.text = element_text(size = 15))  

ggsave(paste0("figs/SSB_", "_", sp, ".pdf"), plot2)      
        
 } 

```


###How did the fisheries sustainability score varied in Norway temporally in each municipality?
```{r message=FALSE}
plot_fscores <-final_scores %>% 
             mutate(county = case_when(
                                    Municip_number >= 1800 &  Municip_number < 1900 ~ "Nordland",
                                    Municip_number >= 1900 &  Municip_number < 2000 ~ "Troms",
                                    Municip_number >= 2000 ~ "Finnmark"
                                    ))

for(i in 1:length(counties))
  {
  cty <- counties[i]
  
  df <- plot_fscores %>%
    filter(county == cty)
  
  plot <- ggplot(df)+
      geom_line(aes(x = Year, y = fish_score))+
      facet_wrap(Municip_name_new  ~ .) +
      scale_x_continuous(breaks = breaks1, limits = c(1994,2018), labels = labels1)+
      labs(y = "Fisheries sustainability score", x = "") +
      ggtitle(cty)+
      theme_stata(scheme = "s1color") +
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 6, angle = 90, hjust = 1, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 7, face = "bold", angle = 0, hjust = 1, vjust = 0),
            axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.1, "cm"))
           
           
ggsave(paste0("figs/fishery_score","_", cty, ".pdf"), plot)
  
}

```

###Re-estimating fisheries sustainability score by smoothing it with a rolling average (zoo package)
```{r}
municipalities <-unique(final_scores$Municip_name_new)

finallist <-list()
for (i in 1:length(municipalities)) 
  {
  mp <- municipalities[i]
  df <- final_scores %>% 
      filter(Municip_name_new == mp)
   runmean <-data.frame(runmean = rollapply(df$fish_score, width = 3, FUN = mean,  fill=NA))
   df2 <- bind_cols(df, runmean)
  finallist[[i]] <- df2  
}

smoothed_fish_scores <- do.call("rbind", finallist)
colSums(is.na(smoothed_fish_scores)) #a lot of NA were introduced with a rolling mean, because some municipailities have too few data lines. Does it make sense to use rolling mean then_



```


