---
title: "Visualize fisheries score calcualtion"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

# Visualize the idea behind fishereis score
See fig.2 in Afflerbach et al 2019 for the similar example.
```{r}
fishsustain <- tibble(
  SSB.Bmsy = c(seq(0, 0.99, by = 0.01), 1, 2, 3),
  Score = c(seq(0, 0.99, by = 0.01), 1, 1, 1),
  Linetype = ifelse(Score <= 0.9999, "dash", "full"),
 
)
```

```{r}
dfpoly <-tibble(
  x = fishsustain$SSB.Bmsy,
  y = fishsustain$Score,
  id = c(rep(1,100), rep(2,3)),
  value = fishsustain$Linetype
)
```



```{r dpi = 200, fig.height=7, fig.width=8}
p1 <- ggplot(data = fishsustain, aes(x = SSB.Bmsy, y = Score)) +
  geom_line(aes(linetype = Linetype), size = 1) +
  labs(x = "SSB/Bmsy", y = "Stock Score") +
  geom_area(data = dfpoly, aes(x = x, y = y, fill = value, group = id), alpha = 0.8) +
  scale_fill_manual(values = c("rosybrown2", "lightsteelblue2")) +
  theme_classic() +
  theme(
    axis.text.y = element_text(size = 15, margin = margin(t = 0, r = 12, b = 0, l = 0), color = "black"),
    axis.text.x = element_text(size = 15, color = "black"),
    axis.title = element_text(size = 16),
    axis.ticks = element_line(size = 0.5),
    legend.position = "none"
  ) +
   annotate("text", x = 0.5, y = 0.03, label = "Over-exploited", size = 7) +
   annotate("text", x = 2, y = 0.03, label = "Fully or Under-exploited", size = 7) 
p1
```
  theme(legend.position = "none", 
            axis.text.x = element_text(size = 6, angle = 90, hjust = 1, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 7, face = "bold", angle = 0, hjust = 1, vjust = 0),
            axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.1, "cm"))
