---
title: "Calculating status score for water entities"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r include = FALSE}
library(sf)
library(vroom)
library(janitor)
library(rebus)
library(DT)
library(kableExtra)
```

# Load the data
```{r}
status_modified <-read.csv("data/quantitative_status_strongly_modified_areas.csv")
status_unmodified <-read.csv("data/quantitative_status_unmodified_areas.csv")
areas <- read.csv("data/water_entity_n_area_per_mcp.csv")
water_entities <- read.csv("data/municipalities_and_water_entities_table.csv")
```

# Prepare the datasets
Prepare the tables of modified and unmodified water entities - cleaning the water entities names
```{r}
status_unmodified_prep <- status_unmodified %>% 
  mutate(water_entity_id = str_replace_all(water_entity_id,
                                            pattern = "_-_",
                                           replacement = "_")) %>% 
  mutate(water_entity_id = str_replace_all(water_entity_id,
                                            pattern = "-",
                                            replacement = "_"))                     
```


```{r}
status_modified_prep <- status_modified %>% 
  mutate(water_entity_id = str_replace_all(water_entity_id,
                                            pattern = "_-_",
                                           replacement = "_")) %>% 
  mutate(water_entity_id = str_replace_all(water_entity_id,
                                            pattern = "-",
                                            replacement = "_")) 
```

# Calculating the water entity status score for modified and unmodified water entities
STEPS: 
1) Prepare a table like water_entity, its score and reliability score  (3 columns) for modified and unmodified areas separately
2) Join the areas table with both scores tables by water entity id.
3) Create a column "type" that will indicate if water entity is modified or not.

```{r}
unmodified_score <- status_unmodified_prep %>% 
  mutate(score = 
           pmap_dbl(
             list(
    num_status_eco,
    num_status_chem
  ),
  ~ mean(.x, na.rm = T)
  )) %>% 
  select(
    water_entity_id,
    score,
    okologisk_palitelighetsgrad,
    kjemisk_palitelighetsgrad) %>% 
  mutate(reliability_ecological_score = case_when(
    okologisk_palitelighetsgrad == "lav" ~ "low",
    okologisk_palitelighetsgrad == "hoy" ~ "high",
    okologisk_palitelighetsgrad == "middels" ~ "moderate",
  )) %>% 
   mutate(reliability_chemical_score = case_when(
    kjemisk_palitelighetsgrad == "lav" ~ "low",
    kjemisk_palitelighetsgrad == "hoy" ~ "high",
    kjemisk_palitelighetsgrad == "middels" ~ "moderate",
  )) %>% 
  select(-c(
    okologisk_palitelighetsgrad,
    kjemisk_palitelighetsgrad 
  )) %>% 
  mutate(water_entity_class = "unmodified")

```

Here I include reliability score of the assessment also for modified areas. For modified areas, we don't need to include chemical status - for most of entities (34 out of 42) it was not assessed. 
For the 21 entities that do not have ecological potential assessed, we will assume that their potential is 25 - moderate.
```{r}
modified_score <- status_modified_prep %>% 
  select(water_entity_id,
         okologisk_palitelighetsgrad,
         num_status_eco) %>% 
 mutate(reliability_ecological_score = case_when(
    okologisk_palitelighetsgrad == "lav" ~ "low",
    okologisk_palitelighetsgrad == "hoy" ~ "high",
    okologisk_palitelighetsgrad == "middels" ~ "moderate",
  )) %>% 
  mutate(num_status_eco = ifelse(
    is.na(num_status_eco), 
    25,
    num_status_eco
  )) %>% 
  select(-okologisk_palitelighetsgrad) %>% 
  rename(score = num_status_eco) %>% 
  mutate(water_entity_class = "modified")
```


Bind the tables for modified and unmodified status, the column "reliability of chemical score" will have NA for all the modified areas, as I did not include chemical status. 
Also I realized that there are still some bugs in the water entities' names. I fix them here.
```{r}
scores_all <- bind_rows(
  unmodified_score,
  modified_score
) %>% 
  mutate(water_entity_id = 
           str_replace_all(
             water_entity_id,
             pattern = "_/_",
             replacement = "_"
           )) %>% 
   mutate(water_entity_id = 
           str_replace_all(
             water_entity_id,
             pattern = "__",
             replacement = "_"
           )) %>% 
  mutate(water_entity_id = 
           str_replace_all(
  water_entity_id,
  pattern = "/_",
  replacement = "_"))
```

Now I join the table of water entities and their areas with a table of scores. First, I need to change the name of water entities to lower case.
```{r}
scores_per_mcp <- areas %>% 
  mutate(water_entity_id = str_to_lower(water_entity_id)) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = char_class("- "),
    replacement = "_"
  )) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "___",
    replacement = "_")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "__",
    replacement = "_")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "_/_",
    replacement = "_")) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "/_",
    replacement = "_")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = ",",
    replacement = "")) %>% 
    mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "ø",
    replacement = "o")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "å",
    replacement = "a")) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "æ",
    replacement = "ae")) %>% 
  left_join(scores_all, by = "water_entity_id") %>% 
  mutate(score = ifelse(
    is.nan(score),
    NA,
    score
  ))
```

Note that there are 1284 observations in the areas dataframe, but only 1019 in the scores. This is because areas includes water entity per municipality, and some of the water entities spread over several municipalities. So, we have duplicates of water_entity_id in the resultant scores_per_mcp table. The scores_per_mcp table has 1284 variables, because 



Some of the water entities from northern Norway that were in the shape file, were not in the assessment file: there are 2 of such entities. I checked, and it appears that these two entities are lakes that have a connection to fjords. On the portal Vann-Nett they do not have ecological status assessed. I will just remove them. 
```{r}
not_assessed <- scores_per_mcp %>% 
  filter(is.na(reliability_ecological_score)) %>% 
  group_by(water_entity_id) %>% 
  tally()

not_assessed
```

At the same time, there is 1 entity, that is in the status table but is not listed in the shape files.

```{r}
not_in_shapes <-left_join(scores_all, scores_per_mcp, by = "water_entity_id")
print(filter(not_in_shapes, is.na(reliability_ecological_score.y)))
```
I will create another areas file where I will manually add this entity (in the previous run of the cod, there were 10 such entities, because i mistakenly did not include entities in Harstad municipality)
```{r}
areas_addition <- read_excel("data/entities_not_in_shape_file.xlsx")

areas_addition <- areas_addition %>%  
  select(-area_km2) %>%
 filter(municip_name == "Beiarn") %>% 
 rename(area = area_m2) 

areas_prep <- rbind(areas, areas_addition)
```

Clean the water entities names in the areas_prep table, before joining with scores
```{r}
areas_prep2 <- areas_prep %>% 
  mutate(water_entity_id = str_to_lower(water_entity_id)) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = char_class("- "),
    replacement = "_"
  )) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "___",
    replacement = "_")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "__",
    replacement = "_")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "_/_",
    replacement = "_")) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "/_",
    replacement = "_")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = ",",
    replacement = "")) %>% 
    mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "ø",
    replacement = "o")) %>% 
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "å",
    replacement = "a")) %>% 
  mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "æ",
    replacement = "ae")) 
```


And now merge with scores.
Now there will be 1285 observations in total, 1284 from the initail areas table and 1 more water entity, that was not in the shape files.
```{r}
scores_per_mcp_prep2 <-areas_prep2 %>% 
  left_join(scores_all, by = "water_entity_id") %>% 
  mutate(score = ifelse(
    is.nan(score),
    NA,
    score
  ))
```


# Translate reliability scores to numerical
Here I translate the reliability scores to numerical as: low - 0, moderate - 50, high - 100. Then we can present an average reliability score per municipality alongside with the clean waters status score. The 2 water entities that were not assesed (two lakes) are removed at this step.
```{r}
scores_per_mcp_prep3 <-scores_per_mcp_prep2 %>% 
  filter(!is.na(reliability_ecological_score)) %>% 
  mutate(reliability_eco_score_num = case_when(
    reliability_ecological_score == "low" ~ "0",
    reliability_ecological_score == "high" ~ "100",
    reliability_ecological_score == "moderate" ~ "50",
  )) %>% 
  mutate(reliability_chem_score_num = case_when(
    reliability_chemical_score == "low" ~ "0",
    reliability_chemical_score == "high" ~ "100",
    reliability_chemical_score == "moderate" ~ "50",
  )) 
  
```


Of the total 1283 records of status, there are 82 in this table than did not have neither ecological nor chemical status assessed. 
I will gapfill this missing values to have more data and not to distort weighting of water entities within a municipality (because I will take weighted average of scores per entity to have a value per municipality, and weights will be areas of entities within a municipality).


# Gapfilling 82 NA scores 

The list of "empty" records with no ecological or chemical score assessed:
```{r}
no_status_areas <- scores_per_mcp_prep3 %>% 
  filter(is.na(score)) 

datatable(no_status_areas)
```
 
To fill the missing values, I will use mean imputation per river-coast system areas (also called drainage areas). Usually, they contain 15-30 water entities.
According to Norwegian Water entities authority, vassdrag area is the area from which a system of rivers drains into a particular area of a coast. So, if we think about land-based pollution, the status of water entities within one vassdrag area can be expected to be similar, but  resolution maybe too large if there is very local pollution.

```{r}
scores_per_mcp_prep4 <-scores_per_mcp_prep3 %>% 
  split(., .$vassdragomrade) %>% 
  map(., ~ mutate(.x, score = ifelse(
    is.na(score),
    mean(score, na.rm = T),
    score
  ))) %>% 
  do.call("rbind", .)
  
```

However, this approach will solve a problem of NA only where there are more than 1 observation per drainage area.  Example - drainage area 228, entity *storfjorden_0422020300_c* is a single assessed entity in the whole drainage area. Drainage area 227 - all three entities did not have status assessed, same for drainage area 232. There are in total 5 such occasions. For this 11 occasions, I will just replace NA with a score of closest water entity. 

```{r}
print(filter(scores_per_mcp_prep4, is.na(score)))
```


I will have to load the shape file, extract the coordinates from it, sort the coordinates, and then create a table with sorted water entities according to their geometries.

```{r}
entity_shapes <- st_read("data/shapefiles/Kystvann20200415.gdb")
shapes_north <- entity_shapes[entity_shapes$Vannregion %in% c("Finnmark", "Troms", "Nordland", "Norsk-Finsk"), ] 
```

I use approach as described [here](https://gis.stackexchange.com/questions/302692/sort-data-in-an-ascending-order-based-on-the-geometry-of-an-sf-object-in-r-sort)
```{r}
xy = st_coordinates(st_centroid(shapes_north))

shapes_sorted <- shapes_north[order(xy[, "X"], xy[, "Y"]), ] %>%
  select(VannforekomstID, Vannforekomstnavn) %>%
  mutate(water_entity_id = stringr::str_c(Vannforekomstnavn, "_", VannforekomstID)) %>%
  st_set_geometry(., NULL) %>%
  mutate(water_entity_id = str_replace_all(water_entity_id,
    pattern = "_-_",
    replacement = "_"
  )) %>%
  mutate(water_entity_id = str_replace_all(water_entity_id,
    pattern = "-",
    replacement = "_"
  )) %>%
  mutate(water_entity_id = str_to_lower(water_entity_id)) %>%
  mutate(water_entity_id = str_replace_all(water_entity_id,
    char_class(" ,"),
    replacement = "_"
  )) %>%
  mutate(water_entity_id = str_replace_all(water_entity_id,
    pattern = "__",
    replacement = "_"
  )) %>%
   mutate(water_entity_id = str_replace_all(
    water_entity_id,
    pattern = "___",
    replacement = "_"
    )) %>% 
  mutate_if(is.character, ~ str_replace_all(., "ø", replacement = "o")) %>%
  mutate_if(is.character, ~ str_replace_all(., "å", replacement = "a")) %>%
  mutate_if(is.character, ~ str_replace_all(., "æ", replacement = "ae")) %>%
  mutate(water_entity_id = str_replace_all(water_entity_id,
    pattern = "__",
    replacement = "_"
  )) %>%
  mutate(water_entity_id = str_replace_all(water_entity_id,
    pattern = "__",
    replacement = "_"
  )) %>% 
  mutate(water_entity_id = str_replace_all(water_entity_id,
    pattern = "_/_",
    replacement = "_"
  )) %>% 
 mutate(water_entity_id = str_replace_all(water_entity_id,
    pattern = "/_",
    replacement = "_"
  ))
```

Merge sorted water entities with the scores.Because there are duplicate of water entities in score_per_mcp table (due to one entity extends over several municipalities), I group by entity and drainage area and take only one of the duplicates. I use drainage area as a grouping variable, because there are occasions (e.g. Akkarfjord) when entities has same IDs but a located in different drainage areas.
```{r}
entities_wt_scores <- scores_per_mcp_prep4 %>% 
  group_by(water_entity_id, vassdragomrade) %>% 
  slice_head(n=1)

entities_sorted <- shapes_sorted %>% 
  left_join(entities_wt_scores, by = "water_entity_id")
```

Gapfill the 11 missing values using last-observation-carried-forward approach:
```{r}
entities_sorted_gapfilled <- entities_sorted %>% 
  mutate(score_gapfilled = zoo::na.locf(score))
```


Filter out water entities that did not have scores and put them back in the main table:
```{r}
entities_to_gapfill <- filter(scores_per_mcp_prep4, is.na(score))$water_entity_id

entities_to_gapfill_prep <- entities_sorted_gapfilled %>% 
  filter(water_entity_id %in% entities_to_gapfill) %>% 
  mutate(score = score_gapfilled) %>% 
  select(-score_gapfilled)
```

```{r}
scores_per_mcp_prep5 <- scores_per_mcp_prep4 %>% 
  filter(!is.na(score)) %>% 
  bind_rows(., entities_to_gapfill_prep)
```


# Calculating areas of each entity within a municipality

Next, I group the data by municipality and calculate the total area of all water entities within a municipality and proportional area of each water entity per municipality. Area is in m2.
```{r}
scores_per_mcp_prep6 <- scores_per_mcp_prep5 %>% 
  select(-c(VannforekomstID,Vannforekomstnavn)) %>% 
  mutate(area_km2 = area/1000000) %>% 
  group_by(municip_name, municip_number) %>% 
  mutate(total_area_km2 = sum(area_km2)) %>% 
  ungroup() %>% 
  mutate(area_prop = area_km2/total_area_km2)
```

Save the table with water entities and their scores
```{r}
#write.csv(scores_per_mcp_prep6, "data/water_entities_scores.csv", row.names = F)
```


# Caclulating final clean water scores and reliability scores per municipality
```{r}
scores_per_mcp_final <-scores_per_mcp_prep6 %>% 
  mutate_at(vars(reliability_eco_score_num,reliability_chem_score_num),
            as.numeric) %>% 
  group_by(municip_name, municip_number) %>% 
  summarize(score_mcp = sum(score*area_prop, na.rm = T),
            reliability_eco_mcp = mean(reliability_eco_score_num, na.rm = T),
            reliability_chem_mcp = mean(reliability_chem_score_num, na.rm = T)) 
  
```



Save the final table
```{r}
#write.csv(scores_per_mcp_final, "data/clean_waters_score_per_mcp.csv", row.names = F)
```

