---
title: "Calculating water area status score"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# source('~/github/nor-prep/prep/src/common.R')
```

```{r include = FALSE}
library(ggplot2)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(dplyr)
```

# Loading chemical and ecological status shape files
```{r}
komlist <-read.csv("../administrative/komlist.csv", sep = ";")
modified_status <- read.csv("data/quantitative_status_strongly_modified_areas.csv")
unmodified_status <- read.csv("data/quantitative_status_unmodified_areas.csv")
water_mcp <- read.csv("data/municipalities_and_water_entities_table.csv")
```


```{r}
water_shp <- st_read("data/shapefiles/Kystvann20200415.gdb")
all_mcp <-st_read(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
```
# Prepare data layers
Select only the municipalities that we use in the project. 
```{r}
komlist <-filter(komlist, !Name == "NonCoast")
ohikom <-all_mcp[all_mcp$KOMM %in% komlist$Komnum,] 

```

# Plot polygons of water entities and municipalities 
To see how the two layers align
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
mcpsf <- sf::st_as_sf(ohikom, coords= c("long", "lat"),crs = 4326, agr="identity")
```

```{r}
water_shp_north <- water_shp[water_shp$Vannregion %in% c("Finnmark", "Troms", "Nordland"), ] 
```


```{r fig.width=5, fig.height=5}
plot1 <- ggplot(data = world) +
   geom_sf(fill = "antiquewhite1") +
   geom_sf(data = mcpsf, fill = "grey", alpha = 0.001,color = "navy", size = 0.5) + 
   geom_sf(data = water_shp_north, fill = "grey", alpha = 0.001,color = "darkred", size = 0.5) +
   coord_sf(xlim = c(10, 32), ylim = c(65,72), expand = FALSE)+
   labs(x = "", y = "")+
   scale_y_continuous(breaks = 66:72)+
   theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), 
        panel.background = element_rect(fill = "aliceblue"),
        axis.ticks= element_line(size = 1),
        axis.ticks.length = unit(0.2, "cm"),
        axis.text = element_text(size = 15, colour = "black"),
        plot.title = element_text(size = 15, hjust = 0.5, vjust = 0.5),
        legend.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"))
plot1
```
# Calcualte the area of water entity within each municipality
First, make coordinate reference system in the two layers similar.
```{r}
st_crs(mcpsf)
st_crs(water_shp_north)
```
Indeed CRS are different between the two and we need to make them similar.

```{r}
water_shp_north_trans <- st_transform(water_shp_north, crs = 32633)
st_crs(water_shp_north_trans)
```


```{r}
entity_within_mcp <- st_intersection(mcpsf, water_shp_north_trans)

entity_within_mcp_prep <- entity_within_mcp %>% 
  select(NAVN,
         KOMM,
         VannforekomstID,
         Vannforekomstnavn,
         Vassdragsomr책de,
         Vannregion,
         Vannomr책de,
         geometry) %>% 
  rename(vannomrade = Vannomr책de,
         vassdragomrade = Vassdragsomr책de
  )

```

Calcualate the area of each intersection
```{r}
entity_within_mcp_prep$area <-st_area(entity_within_mcp_prep)

plot2 <- ggplot(data = world) +
   geom_sf(fill = "antiquewhite1") +
   geom_sf(data = mcpsf, fill = "grey", alpha = 0.001,color = "navy", size = 0.5) + 
   geom_sf(data = entity_within_mcp_prep, fill = "green", alpha = 0.4,color = "darkgreen", size = 0.1) +
   coord_sf(xlim = c(10, 32), ylim = c(65,72), expand = FALSE)+
   labs(x = "", y = "")+
   scale_y_continuous(breaks = 66:72)+
   theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), 
        panel.background = element_rect(fill = "aliceblue"),
        axis.ticks= element_line(size = 1),
        axis.ticks.length = unit(0.2, "cm"),
        axis.text = element_text(size = 15, colour = "black"),
        plot.title = element_text(size = 15, hjust = 0.5, vjust = 0.5),
        legend.text = element_text(face = "bold"),
        legend.title = element_text(face = "bold"))
plot2
  
```

Prepare a final table with area and clean the column names
```{r}
water_mcp_area <- entity_within_mcp_prep %>% 
  rename(municip_name = NAVN,
         municip_number = KOMM,
         entity_id = VannforekomstID,
         entity_name = Vannforekomstnavn,
         vannregion = Vannregion,
         area_m2 = area) %>% 
 mutate(water_entity_id = stringr::str_c(entity_name, "_", entity_id)) %>% 
  st_set_geometry(., NULL) %>% 
  select(municip_name, 
         municip_number, 
         water_entity_id, 
         vannregion,
         vannomrade,
         vassdragomrade,
         area_m2)

#write.csv(water_mcp_area, "data/water_entity_n_area_per_mcp.csv", row.names = F)
```




##########Example of overlaying shapefiles on top of each other from OHI NE##########
We overlay statistical areas with our regions to select just the statistical areas that overlap with our regions.

```{r ohi_stat_areas}
ohi_stat_areas          <- st_intersection(rgns, stat_shp) #intersects statistical areas with OHI regions
ohi_stat_areas$ohi_area <- st_area(ohi_stat_areas) #calculate area of each overlapped polygon
ggplot(ohi_stat_areas) +
  geom_sf(aes(fill = rgn_name)) +
  geom_sf(data = ne_states, fill = "beige") +
  theme_bw() +
  labs(title = "Statistical areas overlapped with OHI regions") 
```

Calculate proportion of each statistical area in our OHI regions. For statistical areas that overlap with OHI regions, we can use proportional area overlap to adjust catch. We assume that catch is evenly distributed across each statistical area.

```{r prop_stat_areas_in_ohi_rgns}
calc_prop_area <- ohi_stat_areas %>%
  group_by(Id) %>%
  mutate(ohi_rgn_prop_area = ohi_area/stat_area) #this column tells us how much of each OHI sub-region falls within the statistical area in our region

```
###########################################