---
title: "Exploring clean waters data: eutrophication and pollution status assessment"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r include = FALSE}
library(vroom)
library(janitor)
library(rebus)
library(DT)
```



The data is downloaded from https://vann-nett.no/portal/

# Loading data

```{r}
status <- vroom(file.path("data/assessment_of_water_areas.csv"))
types <- vroom(file.path("data/water_types.csv"))
komlist <- read.csv("../administrative/komlist.csv", sep = ";")
komnames_updated <-  read.csv("../administrative/municipalities_new_numbers_2020.csv", sep = ";")
```

# Preliminary cleaning of data
The text variables are quite messy, I will do a thorough cleaning of the data so it will be easier to work with them.
I remove Norwegian special charters, spaces, commas, and upper cases.
```{r}
status_prep <- status %>% 
  clean_names(.) %>% 
  filter(vannkategori == "Kystvann") %>% 
  mutate_if(is.character, ~ str_to_lower(.)) %>% 
  mutate_if(is.character, ~ str_replace_all(., char_class(" ,"), replacement = "_")) %>% 
  filter(vannregion %in% c(
    "nordland_og_jan_mayen",
    "troms_og_finnmark"
  )) %>% 
  mutate_if(is.character, ~ str_replace_all(. ,"ø", replacement = "o")) %>% 
  mutate_if(is.character, ~ str_replace_all(. ,"å", replacement = "a")) %>% 
  mutate_if(is.character, ~ str_replace_all(. ,"æ", replacement = "ae"))  

```

Further cleaning of the data: separate municipalities, create date column (and delete time). 
```{r}
status_prep2 <-
  status_prep %>%
  mutate(date = str_extract(
    endret_dato,
    pattern = START %R% one_or_more(DGT) %R% "/" %R% digit(2) %R% "/" %R% digit(4)
  )) %>% 
  select(-endret_dato) %>% 
  separate(kommunenummer, into = c("kom1", "kom2", "kom3", "kom4", "kom5", "kom6", "kom7", "kom8", "kom9")) 
  
```



# Preparing datasets
First, a bit more data cleaning - removing redundant columns.
Names of water entities are not unique, some of them actually occur several times. I will use a combination of water entity name and its ID number to make a unique identifier.
```{r}
status_prep3 <- status_prep2 %>% 
  mutate(water_entity_id = str_c(vannforekomstnavn, "_", vannforekomst_id)) %>% 
  select(-c(vannforekomst_id,
            vannkategori,
            vanntype,
            risikovurdering,
            kommentar_kjemisk_tilstand,
            kommentar_kjemisk_palitelighetsgrad,
            kommentar_okologisk_palitelighetsgrad,
            kommentar_okologisk_palitelighetsgrad,
            kommentar_kjemisk_palitelighetsgrad,
            kommentar_risikovurdering,
            vannregionkoordinator,
            kommunenavn,
            nasjonal_vanntype,
            kommentar_okoligisk_tilstand_eller_potensial
            )) 
  
```

How many water entities are strongly modified (by industries)?
```{r}
modified <- status_prep3 %>% 
  filter(naturlig_eller_smvf == "sterkt_modifisert")

datatable(modified)
```

# Prepare list of municipalities corresponding to each water entity
The municipalities of the Troms and Finnmark in this data set have new numbers, after administrative changes in January 2020. I will get the names for these municipalities from the table with updated municipalities' numbers. 

For the new number of municipalities, check here:
https://www.regjeringen.no/no/tema/kommuner-og-regioner/kommunereform/nyekommuneogfylkesnummer/id2629203/?expand=factbox2684595


```{r}
komlist_prep <- komlist %>%  
  mutate(Komnum = as.character(Komnum)) %>% 
  select(-Kid) %>% 
  filter(!Name == "NonCoast") %>% 
  filter(Komnum < 1900) %>% 
  rename(municip_name = Name,
         municip_number_new = Komnum) %>% 
 select(municip_name, 
        municip_number_new)

komnames_updated_prep <- komnames_updated %>% 
  select(-municip_number_old) 

komlist_complete <- rbind(komlist_prep, komnames_updated_prep)
```


Separate the list of water entities and the respective municipalities, put the resultant table in a long format.
There is a mess with municipalities' numbers. Some of them are written as 3 digits, because a tailing zero is missing and I have to manually add it. All municipalities' numbers starting with "50**" can be deleted - they belong to Trondelag county.
```{r}
water_mcp <- status_prep3 %>% 
  select(water_entity_id, 
         vannregion, 
         vannomrade,
         vassdragsomrade,
         kom1:kom9) %>% 
  pivot_longer(.,
               cols = kom1:kom9,
               names_to = "municipality",
               values_to = "municip_number") %>% 
  mutate(municip_number = ifelse(
  str_detect(.$municip_number, pattern = "^[[:digit:]]{3}$"), 
  str_c(municip_number, "0"),
  municip_number
  )) %>% 
  filter(!str_detect(municip_number, pattern = "^50[[:digit:]]{2}")) %>% 
  left_join(komlist_complete, by = c("municip_number" = "municip_number_new")) %>% 
  filter(!(is.na(municip_number) & is.na(municip_name)))

```


Further cleaning the municipalities' names
```{r}
water_mcp_prep <- water_mcp %>% 
  mutate(municip_name = str_replace_all(municip_name, 
                                        pattern = " ", 
                                        replacement = "")) %>% 
  mutate(municip_name = str_replace_all(municip_name,
                                        pattern = char_class("áå"),
                                        replacement = "a")) %>% 
  mutate(municip_name = str_replace_all(municip_name,
                                        pattern = "ø",
                                        replacement = "o")) %>% 
  mutate(municip_name = str_replace_all(municip_name,
                                        pattern = "æ",
                                        replacement = "ae")) %>% 
  mutate(municip_name = str_trim(municip_name)) %>% 
  mutate(municip_name = str_replace_all(municip_name,
                                        pattern = "-",
                                        replacement = "_")) %>% 
  mutate(water_entity_id = str_replace_all(water_entity_id,
                                            pattern = "_-_",
                                            replacement = "_")) %>% 
  mutate(water_entity_id = str_replace_all(water_entity_id,
                                            pattern = "-",
                                            replacement = "_"))
 
```


An overview of merged municipalities, that I will have to separate
```{r}
merged_mcp <- water_mcp_prep %>% 
  filter(str_length(.$municip_name) >= 20) %>% 
  group_by(municip_name) %>% 
  tally()

merged_mcp[,1]
```

There are only 3 occasions of merged municipalities and 3 that include names in Norwegian, sami and kven language, I can correct them by hand.

```{r}
water_mcp_prep2 <-water_mcp_prep %>% 
  mutate(municip_name = replace(municip_name,
                                municip_name == "Gaivuotna_Kafjord_Kaivuono",
                                "Kafjord"
                                )) %>% 
  mutate(municip_name = replace(municip_name,
                                municip_name == "Porsanger_Porsaŋgu_Porsanki",
                                "Porsanger"
                                )) %>%   
  mutate(municip_name = replace(municip_name,
                                municip_name == "Storfjord_Omasvuotna_Omasvuono",
                                "Storfjord"
                                )) %>% 
  select(-municipality) %>% 
  separate(col = municip_name,
          into =  c("mp1", "mp2", "mp3", "mp4")) %>% 
  pivot_longer(cols = c("mp1":"mp4"),
               names_to = "municipality_order",
               values_to = "municip_name") %>% 
  filter(! is.na(municip_name)) %>% 
  select(-municipality_order)
```
Still have to correct some of the municipalities' names. 
```{r}
water_mcp_prep3 <- water_mcp_prep2 %>% 
  mutate(municip_name = replace(
    municip_name,
    municip_name %in% c("Sor", "Varanger"),
    "Sor-Varanger"  
  )) %>% 
  mutate(municip_name = replace(
    municip_name,
    municip_name == "Unjarga",
    "Nesseby"  
  )) %>% 
   mutate(municip_name = replace(
    municip_name,
    municip_name == "Habmer",
    "Hamaroy"  
  )) %>% 
    mutate(municip_name = replace(
    municip_name,
    municip_name == "Divtasvuodna",
    "Tysfjord"  
  )) %>% 
    mutate(municip_name = replace(
    municip_name,
    municip_name == "Loabak",
    "Lavangen"  
  )) %>% 
  left_join(komlist, by = c("municip_name" = "Name")) %>% 
  rename(municip_number_old = Komnum) %>% 
  select(-Kid) %>% 
  select(municip_name, 
         municip_number, 
         municip_number_old, 
         water_entity_id,
         vannregion,
         vannomrade,
         vassdragsomrade)
```

Save the table
```{r}
# write.csv(water_mcp_prep3, "data/municipalities_and_water_entities_table.csv",
#          row.names = FALSE)
```


# Translate qualitative status to quantitative and save the tables 
For ecological status: very bad and bad are both 0 score, moderate - 50, good - 75, very good - 100.
For chemical status, good is 100, bad is 0.
For strongly modified areas, ecological potential: bad is 0, moderate is 25, good is 50; chemical status - bad is 0, good is 50.
When the status is undefined or irrelevant, i set NA. In this case, if for example, chemical status is undefined, it will not contribute to the final clean waters score.

I will create functions for translating status in each case to a quantitative scale.

**Ecological status**
```{r}
translate_ecostatus <- function(status) {
  if (status %in% c("svaert_darlig", "darlig")) {
    status_num <- 0
  } else if (status == "moderat") {
    status_num <- 50
  } else if (status == "god") {
    status_num <- 75
  } else if (status == "svaert_god") {
    status_num <- 100
  } else {
    status_num <- NA
  }
  status_num
}  
  
```

**Chemical status**
```{r}
translate_chemstatus <-function(status){
  if (status == "darlig") {
    status_num = 0 
  } else if (status == "god") {
    status_num = 100
  } else {
    status_num = NA
  }
  status_num
}
```

**Status for strongly modified areas - ecological potential**
```{r}
translate_sma_status <-function(status){
  if (status == "darlig") {
    status_num = 0 
  } else if (status == "moderat") {
    status_num = 25
  } else if (status == "godt") {
    status_num = 50 
  } else {
    status_num = NA
  }
  status_num
}
```


**Status for strongly modified areas - chemical status**
```{r}
translate_chemstatus_smvf <-function(status){
  if (status == "darlig") {
    status_num = 0 
  } else if (status == "god") {
    status_num = 50
  } else {
    status_num = NA
  }
  status_num
}
```
I will separate strongly modified areas from the other areas, and translate the status for each separately.

In both cases, the column with quantitative status will be called "num_status" (numerical status). 
First, for a strongly modified areas. There is ecological potential and chemical status, but only 9 out of 42 entities have chemical status assessed. It seems that it is not a routine work to assess a chemical status for modified areas, so we can base the score only on the ecological potential.
```{r}
modified_num_status <- modified %>%  
  select(water_entity_id, 
         okologisk_potensial,
         kjemisk_tilstand,
         okologisk_palitelighetsgrad,
         kjemisk_palitelighetsgrad
         ) %>% 
  mutate(num_status_eco = pmap_dbl(
    list(okologisk_potensial),
    ~ translate_sma_status(.x))) %>% 
  mutate(chem_status = pmap_dbl(
    list(kjemisk_tilstand),
    ~ translate_chemstatus_smvf(.x)
  ))

#write.csv(modified_num_status, file = "data/quantitative_status_strongly_modified_areas.csv", row.names = FALSE)
```


```{r}
unmodified_num_status <- status_prep3 %>% 
  select(water_entity_id, 
         okologisk_tilstand,
         okologisk_palitelighetsgrad,
         kjemisk_tilstand,
         kjemisk_palitelighetsgrad) %>% 
  mutate(num_status_eco = pmap_dbl(
    list(okologisk_tilstand),
    ~ translate_ecostatus(.x)
  )) %>% 
  mutate(num_status_chem = pmap_dbl(
    list(kjemisk_tilstand),
    ~ translate_chemstatus(.x)
  ))

# write.csv(unmodified_num_status, file =  "data/quantitative_status_unmodified_areas.csv", row.names = FALSE)  
```


