---
title: "Mapping the clean waters goal"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r include = FALSE}
library(here)
library(sf)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
```

Here I will create a simple map of the final aquaculture goal (based on the chemical and ecological assessment of coastal water areas, but not on the marine debris).

# Load the data and municipalities shapes and names
```{r}
waters <- read.csv(here("prep", "clean_waters", "data", "clean_waters_score_per_mcp.csv"))
allkommuner <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
komlist <-read.csv("/Users/marinaespinasse/github/nor-prep/prep/administrative/komlist.csv", sep = ";")
```

# Prepare map layers
Filter the municipalities that we use in the project, from the shape file of all Norwegian municipalities.
In the shapefile, Harstad still has an old number - 1901, while 1903 in our list of municipalities for the project.

```{r}
komlist <-filter(komlist, !Name == "NonCoast")

komlist_prep <- komlist %>% 
  mutate(Komnum = replace(Komnum,
                          Komnum == 1903,
                          1901))

ohikom <-allkommuner[allkommuner$KOMM %in% komlist_prep$Komnum,] 
```

Prepare a world map layers, to use as a basis and a make a shapefile of municipalities and `sf` object
```{r }
world <- ne_countries(scale = "medium", returnclass = "sf")

kommunersf <- sf::st_as_sf(ohikom, coords= c("long", "lat"), agr="identity")

kommunersf_prep <- kommunersf %>% 
  mutate(KOMM = replace(KOMM,
                        KOMM == 1901,
                        1903))
```


Merge the table of clean water scores and the shapes of municipalities. There is a Beiarn municipality, but we don't need it, it is not in the project.
```{r}
plot_waters <- waters %>% 
  left_join(
    kommunersf_prep,
    by = c("municip_number" = "KOMM")
  ) %>% 
  filter(!municip_name == "Beiarn") %>% 
  st_as_sf(.)
```

#  Mapping
```{r}
base_map <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_waters,
    mapping = aes(fill = score_mcp),
    color = "dimgray",
    size = 0.4,
    alpha = 0.7
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_gradientn(colors = c("red", "yellow", "green"),
                      limits = c(50,100),
                      breaks = c(50, 75,  100),
                      labels = c("50", "75", "100"),
                      
    name = "Clean waters index scores",
    guide = guide_colorbar(
      title.vjust = 0.6,
      ticks = TRUE,
      ticks.colour = "dimgray",
      ticks.linewidth = 1
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = ""
  ) 
```



```{r}
komnames <- cbind(plot_waters,st_coordinates(st_centroid(plot_waters)))
komnames_prep <- filter(komnames, !municip_number %in% c(1919, 1920))
```
```{r}
main_map <- base_map +
  geom_sf_text(data = komnames_prep, 
               aes(X,Y, label= municip_number),fontface = "bold", size = 2)
  
```

```{r}
ggsave(here("prep","clean_waters", "figs", "clean_waters_scores_map.pdf") ,
       main_map, width = 10, height = 9, dpi = 300)
```

