---
title: "Cleaning & plotting number of vessels"
author: "Sigrid Engen"
date: "`r Sys.Date()`"
output:
  html_document:
    css: 'C:/Users/sigrid.engen/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: 'C:/Users/sigrid.engen/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true

---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```

## Load packages 
```{r}
library(tidyverse)
library(readxl)
library(plyr)
```

## Load data
```{r}
vessels_1<-read_excel(file.path("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/data/Vessels_by_length_2000_1980.xlsx"), skip=2)
vessels_2<-read_excel(file.path("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/data/Vessels_by_length_2001_2019.xlsx"), skip=2, n_max=1861)

munic<-read.csv(file.path("C:/Users/sigrid.engen/github/nor-prep/prep/administrative/komlist.csv"), sep=";")
```
## Data cleaning
### Vessels 1980-2000
```{r}
vessels_1_a <- vessels_1 %>%
  dplyr::rename(
    year = År, 
    county = FYLKE, 
    municip_numb = KOMMUNENR, 
    municipality = KOMMUNE, 
    vessel_length = LENGDE, 
    number_of_vessels = `ANTALL FARTØY`) %>%
  filter(county %in% c("Finnmark", "Troms", "Nordland"))

vessels_1_a$vessel_length <- revalue(vessels_1_a$vessel_length, c("0-9,99m" = "0to9.99m", "10-10,99m" = "10to10.99m", "11-14,99m" = "11to14.99m", "15-20,99m" = "15to20.99m", "21-27,99m" = "21to27.99m", "28 m og over" = "28_and_larger"))
```
### Vessels 2001-2019
```{r}
vessels_2_a <- vessels_2 %>%
  dplyr::rename(vessel_length = ...1, 
         municipality = ...2) %>%
  mutate(vessel_length = 
           zoo::na.locf(vessel_length)) %>%                                               #fills in length names in NA columns 
  separate(municipality, c("municip_numb", "municipality"), sep = 4) %>%                  # separate municipality name and number in two columns
  mutate(county = ifelse(municip_numb %in% 1800:1899, "Nordland",                         #add county variable
                                     ifelse(municip_numb  %in% 1900:1999, "Troms",
                                            ifelse(municip_numb  %in% 2000:2099, "Finnmark", "Missing"
                                              )))) %>%
  filter(county %in% c("Finnmark", "Troms", "Nordland")) %>%
  gather(key = year, 
         value= number_of_vessels, 
         -municipality, 
         -municip_numb, 
         -county, 
         -vessel_length)  %>%                  #reshape format of table
  select(year, county, municip_numb, municipality, vessel_length, number_of_vessels) # reorder columns to match the other dataset

vessels_2_a$municipality <- trimws(vessels_2_a$municipality)   #remove space at beginning of string
  
vessels_2_a$vessel_length <- revalue(vessels_2_a$vessel_length, c(   # revalue vessel length factor levels
    "1. Under 10 meter største lengde" = "0to9.99m", 
    "2. 10-10,99 meter største lengde" = "10to10.99m", 
    "3. 11-14,99 meter største lengde" = "11to14.99m", 
    "4. 15-20,99 meter største lengde" = "15to20.99m",  
    "5. 21-27,99 meter største lengde" = "21to27.99m",  
    "6. 28 meter største lengde og over" = "28_and_larger")) 
    
```
### Merge the two datasets

```{r}
vessels <- rbind(vessels_1_a, vessels_2_a)
```
### Complete the dataset with rows missing
```{r}
#Check how many rows a dataset with all combinations of municipality * year * vessel length should have
#levels(as.factor(vessels$municipality))
#levels(as.factor(vessels$year))
#levels(as.factor(vessels$vessel_length))
# 85*6*40 = 20400

#complete the dataset so that it includes all combinations of municipality * year * vessel length 

vessels <- vessels %>%
  complete(nesting(municipality, 
                  municip_numb, 
                  vessel_length, 
                  county), 
           year, 
           fill = list(number_of_vessels = 0)) # add rows for year not included

#check the number of entries for each combination of year, vessel length and municipality - should be equal but some duplicates are made by municipalities that merged during study period
vessels_check<- vessels %>%
 group_by(vessel_length, 
         year) %>%
 dplyr::summarise(count = n()) 

```
### Pool values from municipalities that merged during study period
#### Hammerfest
```{r}
Hammerfest_old<-filter(vessels, 
                       municip_numb == "2001"
                       )

Soroysund<-filter(vessels, 
                  municip_numb == "2016"
                  )

Hammerfest_new<-filter(vessels, 
                       municip_numb == "2004"
                       ) 

Hammerfest_df<-mutate(Hammerfest_new, 
                      number_of_vessels = 
                        Hammerfest_old$number_of_vessels + 
                        Hammerfest_new$number_of_vessels + 
                        Soroysund$number_of_vessels
                      )                                       #pooling merged municipalities

vessels_final_1<-left_join(vessels, Hammerfest_df, by =c("municip_numb","vessel_length","year")) %>% 
  mutate(number_of_vessels = 
           ifelse(municip_numb == "2004", 
                  number_of_vessels.y, 
                  number_of_vessels.x
                  )) %>% #return number_of_vessels.y for rows with Hammerfest 2004 & number_of_vessels.x if not
  select(-c(number_of_vessels.y, 
            number_of_vessels.x, 
            county.y, 
            municipality.y)) %>% #remove reduntant variables
  dplyr::rename(municipality = municipality.x,
         county = county.x)%>%
  filter(!municip_numb %in% 
           c("2016", "2001")) #remove Sørøysund 2016, Hammerfest 2001 

``` 
#### Harstad
```{r}
Harstad_old<-filter(vessels_final_1, municip_numb == "1901")
Bjarkoy<-filter(vessels_final_1, municip_numb == "1915")
Harstad_new<-filter(vessels_final_1, municip_numb == "1903") 

Harstad_df<-mutate(Harstad_new, 
                   number_of_vessels = 
                     Harstad_old$number_of_vessels + 
                     Harstad_new$number_of_vessels + 
                     Bjarkoy$number_of_vessels) #pooling merged municipalities

vessels_final_2<-left_join(vessels_final_1, 
                       Harstad_df, 
                       by =c("municip_numb","vessel_length","year")) %>%
  mutate(number_of_vessels = 
           ifelse(municip_numb == "1903", 
                  number_of_vessels.y, 
                  number_of_vessels.x)) %>% #return number_of_vessels.y for rows with Harstad 1903 & number_of_vessels.x if not
  select(-c(number_of_vessels.y, 
            number_of_vessels.x, 
            county.y, 
            municipality.y))%>% #remove reduntant variables
  dplyr::rename(municipality = municipality.x,
         county = county.x)%>%
  filter(!municip_numb %in% 
           c("1901", "1915")) #remove Bjarkøy 1915, Harstad 1901
```
#### Bodo
```{r}
Bodo<-filter(vessels_final_2, municip_numb == "1804")
Skjerstad<-filter(vessels_final_2, municip_numb == "1842")

Bodo_df<-mutate(Bodo, 
                   number_of_vessels = 
                     Bodo$number_of_vessels + 
                     Skjerstad$number_of_vessels) #pooling merged municipalities

vessels_final_3<-left_join(vessels_final_2, 
                       Bodo_df, 
                       by =c("municip_numb","vessel_length","year")) %>%
  mutate(number_of_vessels = 
           ifelse(municip_numb == "1804", 
                  number_of_vessels.y, 
                  number_of_vessels.x)) %>% #return number_of_vessels.y for rows with Bodo 1804 & number_of_vessels.x if not
  select(-c(number_of_vessels.y, 
            number_of_vessels.x, 
            county.y, 
            municipality.y))%>% #remove reduntant variables
  dplyr::rename(municipality = municipality.x,
         county = county.x)%>%
  filter(!municip_numb %in% 
           c("1842")) #remove Skjerstad 1842

```
### Final cleaning - select only study municip & rename municipal names
```{r}
#look at which municipalities are in main and not in study municipalities list
#munic$Komnum<-as.factor(munic$Komnum) #change on numeric variable to factor
#f<-anti_join(vessels_final_3, munic, by=c("municip_numb"="Komnum"))
# Beiarn 1839 is in vessels_final_3 but not a study municipality

# remove Beiarn 
vessels_final_4 <- vessels_final_3 %>%
  filter(!municip_numb %in% 
           c("1839")) 

#add new column with new municipality names
munic$Komnum<-as.character(munic$Komnum) #change numeric variable to factor

vessels_final_4$municip_numb<-as.character(vessels_final_4$municip_numb) #change on numeric variable to factor

vessels_final_4<- left_join(vessels_final_4, munic, by = c("municip_numb"="Komnum"))

vessels_final_4<- vessels_final_4 %>%
  select(-municipality, 
         -Kid, 
         ) %>%
  dplyr::rename(
    municipality = "Name"
  ) %>%
  droplevels()   #drop unused levels from data frame (e.g. noncoast)


#Check how many rows a dataset with all combinations of municipality * year * vessel length should have
#levels(as.factor(vessels_final_4$municipality))
#levels(as.factor(vessels_final_4$year))
#levels(as.factor(vessels_final_4$vessel_length))
#should be 19440 rows 

```
## Plotting data
### Plot number of vessels through time by length
```{r}
#collapse lenght categories in new variable
vessels_plot<-vessels_final_4%>%
  dplyr::mutate(vessel_length_2= 
           fct_collapse(vessels_final_4$vessel_length, 
      below11m  = c("0to9.99m", "10to10.99m"),
      from11to15m = c("11to14.99m"),
      from15to21m = c("15to20.99m"),
      above21m = c("21to27.99m","28_and_larger"))
  ) 
  
vessels_plot$year<- as.numeric(vessels_plot$year)
    
county <- unique(vessels_plot$county) 

for(i in 1:length(county))
  {
  cty <- county[i]

  df <- vessels_plot %>%
    filter(county == cty) %>% 
    group_by(year,
             municip_numb,
             municipality,
             vessel_length_2,
             county) %>% 
    dplyr::summarise(number_of_vessels = sum(number_of_vessels)) 
    

plot <- ggplot(df)+
      geom_line(aes(x = year, y = number_of_vessels, color = vessel_length_2)) +
      facet_wrap(municipality  ~ .) +
      scale_color_manual(values = c("darkred", "steelblue", "darkgreen", "black"))+
      labs(y = "Number of vessels", x = "") +
      ggtitle(cty)+
            theme(legend.position = "right",
            legend.title = element_blank(), 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))

ggsave(paste0("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/figs/number_of_vessels_length", cty, ".pdf"), plot)

}

```

### Plot number of vessels through time by length (without below 11m group)
```{r}
#collapse length categories in new variable
vessels_plot<-vessels_final_4%>%
  dplyr::mutate(vessel_length_2= 
           fct_collapse(vessels_final_4$vessel_length, 
      below11m  = c("0to9.99m", "10to10.99m"),
      from11to15m = c("11to14.99m"),
      from15to21m = c("15to20.99m"),
      above21m = c("21to27.99m","28_and_larger"))
  ) %>%
  filter(vessel_length_2 != "below11m" )
  
vessels_plot$year<- as.numeric(vessels_plot$year)
    
county <- unique(vessels_plot$county) 

for(i in 1:length(county))
  {
  cty <- county[i]

  df <- vessels_plot %>%
    filter(county == cty) %>% 
    group_by(year,
             municip_numb,
             municipality,
             vessel_length_2,
             county) %>% 
    dplyr::summarise(number_of_vessels = sum(number_of_vessels)) 
    

plot <- ggplot(df)+
      geom_line(aes(x = year, y = number_of_vessels, color = vessel_length_2)) +
      facet_wrap(municipality  ~ .) +
      scale_color_manual(values = c("darkred", "steelblue", "darkgreen"))+
      labs(y = "Number of vessels", x = "") +
      ggtitle(cty)+
            theme(legend.position = "right",
            legend.title = element_blank(), 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))

ggsave(paste0("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/figs/number_of_vessels_length_minus_below11m", cty, ".pdf"), plot)

}

```