---
title: "Number and age of fishers"
author: "Sigrid Engen"
date: "27 4 2020"
output: html_document
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning=FALSE)
```
## Load packages
```{r, message = FALSE}

library(rgdal)
library(tidyverse)
library(raster)
#library(plyr)
library(scatterpie)
library(ggthemes)
library(forcats)
library(ggplot2)
```
## Load data
```{r, message = FALSE} 
#Fishers
fishers<-read.csv("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/data/number_age_parttime_fulltime_fishers.csv")

#Norway counties
Norway<-readOGR(dsn="C:/Users/sigrid.engen/github/nor-prep/prep/administrative/raw//shapefiles/NO_Fylker", "NO_Fylker_pol")

#Municipalities 
municip<-readOGR(dsn="C:/Users/sigrid.engen/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader", "NO_AdminOmrader_pol")

#Study municipalities
stud_municip<-read.csv(file.path("C:/Users/sigrid.engen/github/nor-prep/prep/administrative/komlist.csv"), sep=";") 

#Study muncipalities centroid - coordinates 
centroid<-read.csv(file.path("C:/Users/sigrid.engen/github/nor-prep/prep/administrative/coordinates_centroid_study_municipalities.csv")) 
```
## Clean fishers data
```{r, message = FALSE}
#rename some long municipality names

fishers<-fishers%>%
  mutate(municip_name = recode(municip_name, 
        "Unjaargga-Nesseby" = "Nesseby", 
        "Gaivuotna-Kåfjord" = "Kåfjord",
        "Herøy i Nordland" = "Herøy", 
        "Bø i Nordland" = "Bø",
        "Deatnu-Tana" = "Tana",
        .default = levels(municip_name)
        )
  )

```

## Plot data 
### Number of fishers 2018 - spatial scatterpie
```{r, message = FALSE} 
fishers_main_2018<-fishers%>% 
  dplyr::group_by(municip_numb,
           year)%>%
  summarise(tot_main = sum(main_occupation))%>%
  dplyr::filter(year == "2018")

# obs group by not working when plyr loaded after dplyr https://stackoverflow.com/questions/26923862/why-are-my-dplyr-group-by-summarize-not-working-properly-name-collision-with 
#detach(package:plyr)

#extract Northern-Norway
counties<-subset(Norway,NAVN=="Troms" | NAVN=="Finnmark" | NAVN=="Nordland")

#extract municipalities in Northern Norway
municip.stud<-crop(municip, counties)

#make map object for use in ggplot https://github.com/tidyverse/ggplot2/wiki/plotting-polygon-shapefiles
municip.stud@data$id = rownames(municip.stud@data) #explicitly identifies attribute rows by the .dbf offset.

#melts the polygons into points, tags each point with the id value of the corresponding attribute row, and tags each point with values from the polygon from which the point was derived.
municip.stud.points = fortify(municip.stud, 
                              region="id")

#joins the points to their corresponding attributes
municip.stud.df = plyr::join(municip.stud.points, 
                             municip.stud@data, 
                             by="id")  

p<-ggplot(municip.stud) + 
  aes(long,lat,
      group=group) + 
  geom_polygon(fill="grey") +
  geom_path(color="white") +
  coord_equal() +
  scale_fill_brewer("Northern Norway")

fishers_main_2018_2<-left_join(fishers_main_2018,centroid, by=c("municip_numb" = "Muni_Number"))%>%
  mutate(radius = tot_main*100)
  
p + geom_scatterpie(data = fishers_main_2018_2, 
                  aes(x, y, r = radius),
                  cols = c("tot_main"))
#good ref to scatter plots on maps https://www.brodrigues.co/blog/2018-10-27-lux_elections_analysis/              

```
##left to do scatterplot- add legend and proportion parttime vs full time fisher

## Plot number of fishers through time - main occupation
```{r}

county <- unique(fishers$county) 

for(i in 1:length(county))
  {
  cty <- county[i]

  df <- fishers %>%
    filter(county == cty) %>% 
    group_by(year,
             municip_numb,
             municip_name,
             county) %>% 
    summarise(main_fishers = sum(main_occupation)) 

    
plot <- ggplot(df)+
      geom_col(aes(x = year, y = main_fishers), fill = "dodgerblue4")+
      facet_wrap(municip_name  ~ .) +
      labs(y = "Number of fishers", x = "") +
      ggtitle(cty)+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))

ggsave(paste0("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/figs/fishers_fulltime", cty, ".pdf"), plot)

}
```
## Plot number of fishers through time - part time occupation
```{r}

county <- unique(fishers$county) 

for(i in 1:length(county))
  {
  cty <- county[i]

  df <- fishers %>%
    filter(county == cty) %>% 
    group_by(year,
             municip_numb,
             municip_name,
             county) %>% 
    summarise(part_fishers = sum(part_time_occupation)) 

    
plot <- ggplot(df)+
      geom_col(aes(x = year, y = part_fishers), fill = "dodgerblue4")+
      facet_wrap(municip_name  ~ .) +
      labs(y = "Number of fishers", x = "") +
      ggtitle(cty)+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))

#ggsave(paste0("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/figs/fishers_parttime", cty, ".pdf"), plot)

}
```

## Plot number of fishers through time - main occupation by age
```{r}
#collapse age categories in new variable
fishers<-fishers%>%
  dplyr::mutate(age2= 
           fct_collapse(fishers$age, 
      below30  = c("<20","20-29"),
      from30to49 = c("30-39","40-49"),
      from50to66 = c("50-59", "60-66"),
      retired = c("67-69", "70 OG OVER"))
  )
    
county <- unique(fishers$county) 

for(i in 1:length(county))
  {
  cty <- county[i]

  df <- fishers %>%
    filter(county == cty) %>% 
    group_by(year,
             municip_numb,
             municip_name,
             age2,
             county) %>% 
    summarise(main_fishers_age = sum(main_occupation)) 

    
plot <- ggplot(df)+
      geom_line(aes(x = year, y = main_fishers_age, color = age2, linetype = age2)) +
      facet_wrap(municip_name  ~ .) +
      scale_color_manual(values = c("darkred", "steelblue", "darkgreen", "black"))+
      labs(y = "Number of fishers", x = "") +
      ggtitle(cty)+
      theme(legend.position = "right",
            legend.title = element_blank(), 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))

#ggsave(paste0("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/figs/fishers_fulltime_age", cty, ".pdf"), plot)

}

```
## Plot number of fishers through time - part time occupation by age
```{r}
#collapse age categories in new variable
#fishers<-fishers%>%
 # dplyr::mutate(age2= 
  #         fct_collapse(fishers$age, 
  #    below30  = c("<20","20-29"),
   #   from30to49 = c("30-39","40-49"),
   #   from50to66 = c("50-59", "60-66"),
     # retired = c("67-69", "70 OG OVER"))
#  )
    
county <- unique(fishers$county) 

for(i in 1:length(county))
  {
  cty <- county[i]

  df <- fishers %>%
    filter(county == cty) %>% 
    group_by(year,
             municip_numb,
             municip_name,
             age2,
             county) %>% 
    summarise(part_fishers_age = sum(part_time_occupation)) 

    
plot <- ggplot(df)+
      geom_line(aes(x = year, y = part_fishers_age, color = age2, linetype = age2)) +
      facet_wrap(municip_name  ~ .) +
      scale_color_manual(values = c("darkred", "steelblue", "darkgreen", "black"))+
      labs(y = "Number of fishers", x = "") +
      ggtitle(cty)+
      theme(legend.position = "right",
            legend.title = element_blank(),
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))

#ggsave(paste0("C:/Users/sigrid.engen/github/nor-prep/prep/small_sc_fisheries/figs/fishers_parttime_age", cty, ".pdf"), plot)

}
```
## Next plot proportions of full time vs part time through time
## also make html file https://kbroman.org/knitr_knutshell/pages/Rmarkdown.html 



