---
title: "Proportion of small fleets to total fleet number"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: "C:/github/nor-prep/prep/templates/norway_banner.html"
  pdf_document:
    toc: true
---
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r libraries, results="hide"}
library(ggthemes)
library(janitor)
```

# Loading data on small fleets and total fleets per municipality and municipalities names
```{r}
smallfleet <-read_excel(file.path(dir_M[2],"Small_scale_fishery/15m_fiskerifartoy per kommune.xlsx"),skip = 3, trim_ws = T) 
totfleet <- read_excel(file.path(dir_M[2],"Small_scale_fishery/Fiskerifartoy per kommune.xlsx"),skip = 3, trim_ws =  T) 
komlist <- read.csv("../administrative/komlist.csv", sep = ";")
```

# Cleaning smallfleets data
```{r}
komlist <- komlist %>%
  mutate(Komnum = as.character(Komnum))

smallfleet <- smallfleet[-c(86:131), ]

smallfleet_prep <- smallfleet %>%
  select(-1) %>%
  rename(Location_Year = X__2) %>%
  gather(key = Year, c("2001":"2018"), value = "smfleet_count") %>%
  mutate(Kommunenr = word(Location_Year, start = 1)) %>%
  mutate(Kommune = word(Location_Year, start = -1)) %>%
  select(-Location_Year) %>%
  left_join(komlist[, c(1, 2)], by = c("Kommunenr" = "Komnum")) %>%
  mutate(Kommune = Name) %>%
  select(-Name) 
```

# Further cleaning small fleets data
The municipalities 1839,1842,1901,1915 did not have name in the komlist dataset. These are the municipalities that were merged with other municipalities since 1990's.
1901 - Harstad, 1915 - former Bjarkoy now Harstad, 1842 - former Skjerskad now Bodo, 1839 - Beiarn that we do not include in the project.
```{r}
smallfleet_prep2 <- smallfleet_prep %>%
  filter(!Kommunenr == "1839") %>%
  mutate(newid = case_when(
    Kommunenr %in% c(1901,1903, 1915) ~ "1",
    Kommunenr %in% c(1804, 1842) ~ "2",
    TRUE ~ Kommunenr
  )) %>%
  group_by(Year, newid) %>%
  summarize(smfleets_count_new = sum(smfleet_count)) 


smallfleet_prep3 <- smallfleet_prep2 %>%
  mutate(Kommunenr_new = newid) %>% 
  mutate(Kommunenr_new = replace(Kommunenr_new, Kommunenr_new == "1", "1903")) %>%
  mutate(Kommunenr_new = replace(Kommunenr_new, Kommunenr_new == "2", "1804")) %>% 
  select(-"newid") %>% 
  mutate(key =  paste(Kommunenr_new, Year, sep = "_")) 
```

26.06 BEGIN HERE
# Cleaning total fleets data
Applying same manupilations to total fleets table as for small fleets
```{r}
totfleet <- totfleet[-c(89:134), ]

totfleet_prep <- totfleet %>%
  rename(Location_Year = X__1) %>%
  gather(key = Year, c("2000":"2018"), value = "allfleets_count") %>%
  mutate(Kommunenr = word(Location_Year, start = 1)) %>%
  mutate(Kommune = word(Location_Year, start = -1)) %>%
  select(-Location_Year) 

totfleet_prep2 <- totfleet_prep %>%
  filter(!Kommunenr == "1839") %>%
  mutate(newid = case_when(
    Kommunenr %in% c(1901, 1903, 1915) ~ "1",
    Kommunenr %in% c(1804, 1842) ~ "2",
    TRUE ~ Kommunenr
  )) %>%
  mutate(allfleets_count = as.numeric(allfleets_count)) %>% 
  group_by(Year, newid) %>%
  summarize(allfleets_count_new = sum(allfleets_count)) 


totfleet_prep3 <- totfleet_prep2 %>%
  mutate(Kommunenr_new = newid) %>% 
  mutate(Kommunenr_new = replace(Kommunenr_new, Kommunenr_new == "1", "1903")) %>%
  mutate(Kommunenr_new = replace(Kommunenr_new, Kommunenr_new == "2", "1804")) %>% 
  select(-"newid") %>% 
  mutate(key =  paste(Kommunenr_new, Year, sep = "_")) 
```

# Merge data on small fleets and total number of fleets and estimate proportion
```{r}
smallfleet_final <-smallfleet_prep3 %>% 
  left_join(totfleet_prep3[, c(2,4)], by = "key") %>% 
  mutate(small_fleet_prop = smfleets_count_new/allfleets_count_new)  %>% 
  select(-key)


#just check if there are NAs
map(smalfleet_final, ~ sum(is.na(.)))

#write_csv(smallfleet_final,"data/prop_small_fleets.csv")
```

