---
title: "Quota ownership"
author: "Sigrid Engen"
date: "29 12 2020"
output: html_document
---

## Load packages 
```{r, echo=T, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(DescTools) #gini coefficient

```

## Load data
```{r}
quota<-readr::read_rds("data/quota_ownership_auditor_general_do_not_share")
m<-readr::read_rds("data/municipality_names_coordinates_from_auditor_general")
municip<-read.csv(file.path("C:/Users/sigrid.engen/github/nor-prep/prep/administrative/komlist.csv"), sep=";")
```

## Clean data
```{r}

##Quota data are individuals by vessels by year. Some individuals may have quota shares in multiple vessels. Must summarise to get total individual share

quota_1 <- quota %>% select('År', 
                          Eier.kommunenr,
                          Eier.kommune,
                          Eier.fylke,
                          Eier.navn,
                          F.nummer,                        
                          'Kjønn',
                          Alder,
                          Eierandel,
                          Organisasjonsform,
                          'Fartøy.ID',
                          'Største.lengde',
                          VCU,
                          kontrollert.andel.u11m,
                          kontrollert.andel.11.15m,
                          kontrollert.andel.15.20m,
                          kontrollert.andel.21.28m
                          ) %>% 
  group_by(F.nummer) %>% 
  mutate(individual = cur_group_id()) %>% 
  ungroup() %>% 
  rename(year = "År", 
         municip_numb = "Eier.kommunenr",
         municipality = "Eier.kommune",
         county = "Eier.fylke",
         name_quota_owner = "Eier.navn",
         social_sec_numb ="F.nummer",
         gender ="Kjønn",
         age = "Alder", 
         ownership_proportion = "Eierandel",
         organization = "Organisasjonsform",
         vessel_id = "Fartøy.ID", 
         cod_b11m = "kontrollert.andel.u11m",
         cod_11_15m = "kontrollert.andel.11.15m",
         cod_15_20m = "kontrollert.andel.15.20m",
         cod_21_28m = "kontrollert.andel.21.28m") 

#check if annual values add to 100%

sum100<-quota_1 %>% 
  group_by(year) %>%  
  summarise(across(cod_b11m:cod_21_28m, sum, na.rm= TRUE))

# The data from 2004-2007 seems to be incomplete
                
 
  # add the names and numbers of municipalities that merged in the study period to the study areas list in order to extract all the relevant population data

municip_1<-municip %>% 
  select(-Kid) %>% 
  add_row(Komnum = 2001, Name = "Hammerfest") %>% 
  add_row(Komnum = 2016, Name = "Soroysund") %>%
  add_row(Komnum = 1901, Name = "Harstad") %>%
  add_row(Komnum = 1915, Name = "Bjarkoy") %>%
  add_row(Komnum = 1842, Name = "Skjerstad") %>% 
  rename(municip_numb = "Komnum")


# check 

g<-anti_join(municip_1, quota, by= c("Komnum" = "Eier.kommunenr"))


# Liste med variabelnavnene for kvoteandel.
varlist1 <- dplyr::vars(c("kontrollert.andel.u11m",
                   "kontrollert.andel.11.15m",
                   "kontrollert.andel.15.20m",
                   "kontrollert.andel.21.28m",
                   "kontrollert.andel.torsketraal",
                   "kontrollert.andel.NVG",
                   "kontrollert.andel.ringnot",
                   "kontrollert.andel.konv.torsk",
                   "kontrollert.andel.NVG.kyst",
                   "kontrollert.andel.makrell.not.u13m",
                   "kontrollert.andel.makrell.not.13.21m",
                   "kontrollert.andel.makrell.garn.u13m",
                   "kontrollert.andel.makrell.garn.13.21m"))


# Hvis du vil ha et paneldatasett med alle kvoteandelene (i prosent) for hver kommune, for alle år

dt <- eierdata.komplett %>% 
  group_by(År, Eier.kommunenr) %>% 
  summarise_at(varlist1, 
               sum, 
               na.rm = TRUE) %>% 
  


# Fletter inn kommunenavn og posisjon.
dt2 <- dt %>% left_join(kommunenavn, by = c("Kommunenummer"))


# Eksempel på utregning hvis du bare vil se på kvoteandel torsk < 11 m

test <- eierdata.komplett %>% 
  group_by(År, Eier.kommunenr) %>% 
  summarise(kontrollert.andel.u11m = sum(kontrollert.andel.u11m, na.rm = TRUE))

x <- c(541, 1463, 2445, 3438, 4437, 5401, 6392, 8304, 11904, 22261)
# compute Gini coefficient
gini(x)
# generate a vector of weights.
w <- runif(n=length(x))
gini(x,w)

#https://www.r-bloggers.com/2013/01/calculating-a-gini-coefficients-for-a-number-of-locales-at-once-in-r/  
# https://www.khanacademy.org/economics-finance-domain/ap-microeconomics/ap-consumer-producer-surplus/inequality/v/gini-coefficient-and-lorenz-curve 


```

## Cod quota shares by municipality through time (2008-2018)
```{r}
varlist1 <- dplyr::vars(c("cod_b11m",
                           "cod_11_15m",
                           "cod_15_20m",
                           "cod_21_28m"))
               
dt <- quota_1 %>% 
  filter(county %in% c("Finnmark", "Troms", "Nordland")) %>% 
  group_by(year, municip_numb) %>% 
  summarise_at(varlist1, 
               sum, 
               na.rm = TRUE) %>% 
  mutate(municip_numb = as.numeric(municip_numb)) %>% 
  left_join(municip_1, by = c("municip_numb"))

# Karasjok and Kautokeino are in the data - no qouta share -remove?

  
```
## Plot 
```{r, fig.width = 11, fig.height = 8}

ggplot(dt)+
      geom_col(aes(x = year, y = cod_b11m), fill = "dodgerblue4")+
      facet_wrap(Name  ~ .) +
      labs(y = "", x = "") +
      ggtitle("test")+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))
```
## Plot 
```{r, fig.width = 11, fig.height = 8}

ggplot(dt)+
      geom_col(aes(x = year, y = cod_11_15m), fill = "dodgerblue4")+
      facet_wrap(Name  ~ .) +
      labs(y = "", x = "") +
      ggtitle("test")+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))
```
## Plot 
```{r, fig.width = 11, fig.height = 8}

ggplot(dt)+
      geom_col(aes(x = year, y = cod_15_20m), fill = "dodgerblue4")+
      facet_wrap(Name  ~ .) +
      labs(y = "", x = "") +
      ggtitle("test")+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))
```
## Plot 
```{r, fig.width = 11, fig.height = 8}

ggplot(dt)+
      geom_col(aes(x = year, y = cod_21_28m), fill = "dodgerblue4")+
      facet_wrap(Name  ~ .) +
      labs(y = "", x = "") +
      ggtitle("test")+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))
```
## Quota shares individuals within municipality through time (gini coefficient)
```{r}
#grouping on social sec number but what about those values without social sec number

dt_2 <- quota_1 %>% 
  filter(county %in% c("Finnmark", "Troms", "Nordland")) %>% 
  group_by(year, municip_numb, social_sec_numb) %>% 
  summarise_at(varlist1, 
               sum, 
               na.rm = TRUE) %>% 
  ungroup() %>% 
  mutate(municip_numb = as.numeric(municip_numb)) %>% 
  left_join(municip_1, by = c("municip_numb"))

dt_3 <- dt_2 %>% select(-c(cod_11_15m:cod_21_28m)) %>% 
  filter(!cod_b11m == 0) %>%   # remove zero values - only look at those with cod quota
  group_by(year, municip_numb) %>% 
  summarise(count =n()) %>% 
  left_join(municip_1, by = c("municip_numb"))

dt_gini_codb11<- dt_2 %>% 
  select(-c(cod_11_15m:cod_21_28m)) %>% 
  filter(!cod_b11m == 0) %>%
  group_by(year, municip_numb) %>% 
  summarise(gini = Gini(cod_b11m))%>% 
  left_join(municip_1, by = c("municip_numb"))

dt_gini_codb11_15<- dt_2 %>% 
  select(-c(cod_b11m, cod_15_20m:cod_21_28m)) %>% 
  filter(!cod_11_15m == 0) %>%
  group_by(year, municip_numb) %>% 
  summarise(gini = Gini(cod_11_15m))%>% 
  left_join(municip_1, by = c("municip_numb"))
  
```
## Plot 
```{r, fig.width = 11, fig.height = 8}

ggplot(dt_3)+
      geom_col(aes(x = year, y = count), fill = "dodgerblue4")+
      facet_wrap(Name  ~ .) +
      labs(y = "", x = "") +
      ggtitle("test")+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))


```

## Plot 
```{r, fig.width = 11, fig.height = 8}

ggplot(dt_gini)+
      geom_col(aes(x = year, y = gini), fill = "dodgerblue4")+
      facet_wrap(Name  ~ .) +
      labs(y = "", x = "") +
      ggtitle("test")+
      theme(legend.position = "none", 
            axis.text.x = element_text(size = 5, angle = 90, hjust = 0.5, vjust = 0.5, face = "bold"),
            axis.text.y = element_text(size = 6, face = "bold", hjust = 0.5, vjust = 0.5, margin = margin(t = 0, r = 2, b = 0, l = 0)),
            plot.title = element_text(hjust=0.5, vjust=0.5),
            axis.ticks.x = element_line(size = 0.5),
            axis.ticks.length = unit(0.2, "cm"))


```
###
2.2.2. Gini coefficient
We use the Gini coefficient to measure changes in the distributionoftheuse ofquota interms of catchshare revenue among active
vessels. The Gini coefficient is a measure of the difference between
the cumulative distribution of catch share revenue and the cumulative distribution of the uniform distribution. Since revenue data
were available for all active vessels in each catch share program,
the Gini coefficient for a census can be calculated as follows:
G =
(2 × i − n − 1) x1
n2u ,
where i = 1 to n; i is the vessel’s rank order in ascending order; x is
the annual catch share species revenue for vessel i; n is the number
of active vessels; u is the mean revenue (Bellú and Liberati, 2006).

Performance of federally managed catch share fisheries in the UnitedStatesAyeisha A. Brinsona,∗, Eric M. ThunbergbaNOAA 

###
https://stats.stackexchange.com/questions/339972/gini-coefficients-accounting-for-firm-exits 

###
In many applications the Gini coefficient is estimated from grouped
data with 5 to 30 categories instead of the microdata (Gastwirth, 1972;
Abounoori and McCloughan, 2003). For example, income or tax statistics
are often grouped for confidentiality reasons (Van Ourti and Clarke, 2011).
Empirical research has shown that the Gini coefficient may have a nonnegligible downward bias when data are grouped (Lerman and Yitzhaki, 1989;
Davies and Shorrocks, 1989; Kwok, 2010). Vice versa, Kwok (2010) noted
that the Gini coefficient increases if a combined household is split into several smaller households or people living alone. Thus, the Gini coefficient
may produce different results for income when the units of analysis are individuals instead of households (Deininger and Squire, 1996). Therefore, in
interpreting the Gini coefficient the demographic structure of a country or
region should be taken into account.

On the Negative Bias of the Gini Coefficient
due to Grouping 

###
The Gini coefficient is a downwardly biased measure of inequality in small populations when
income is generated by one of three common distributions. The paper discusses the sources of
bias and argues that this property is far more general. This has implications for (i) the comparison
of inequality among sub-samples, some of which may be small, and (ii) the use of the Gini in
measuring firm size inequality in markets with a small number of firms. The small sample bias
has often lead to mis-perceptions about trends in industry concentration. THE SMALL SAMPLE BIAS OF THE GINI COEFFICIENT: 