---
title: "Demo of some text matching functionality"
author: "Jamie Afflerbach"
date: "4/25/2019"
output: html_document
---

This script shows some functionality for matching strings between two aquaculture datasets. This primarily uses the `stringr` package from `tidyverse`.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
```

# Load data

```{r}
a <- read_excel("~/github/nor-prep/prep/food_provision/Mariculture/raw/Akvakulturregisteret.xlsx", skip = 1) #skip first row
b <- read_excel("~/github/nor-prep/prep/food_provision/Mariculture/raw/Escapees.xlsx", skip = 2) #skip first 2 rows
```

There are around `r length(unique(a$NAVN))` companies in Akvakulturregisteret and `r length(unique(b$Selskapsnavn))` in the Escapees data.

# Clean data

```{r}
#companies data
comp <- a %>%
  mutate(name = tolower(NAVN)) %>% #turning all letters to lower case
  select(name, POSTNR) %>%
  distinct() %>%
  mutate(name_fix = str_replace_all(name, " as$", ""), #removes all the " as" at the end of company names
         name_fix = str_replace_all(name_fix, " a/s$", "")) 

#escapees data
escape <- b %>%
  mutate(company = tolower(Selskapsnavn)) %>% #turn all letters to lowercase
  select(company, Lok.nr.) %>%
  distinct() %>%
  mutate(name_fix = str_replace_all(company, " as$", ""),
         name_fix = str_replace_all(name_fix, " asa$", ""),
         name_fix = str_replace_all(name_fix, " as, processing", ""))

```


First try to match

```{r}
match1 <- full_join(comp, escape, by = "name_fix") %>%
  filter(!is.na(company), 
         !is.na(name))
```

We successfully matched `r length(unique(match1$name_fix))` companies. That leaves a maximum of `r length(unique(b$Selskapsnavn)) - length(unique(match1$name_fix))` companies to match.


Continuing to clean up `comp` and `escape`

```{r}
escape2 <- escape %>%
  mutate(name_fix2 = case_when(
    name_fix == "aqua gen" ~ "aquagen",
    name_fix == "aqua gen norway" ~ "aquagen",
    name_fix == "fyllingnes fisk" ~ "fyllingsnes fisk",
    name_fix == "grieg seaf. finnmark" ~ "grieg seafood finnmark",
    name_fix == "pan fish norway" ~ "panfish norway",
    TRUE ~ as.character(name_fix)))
```









