---
title: "Plots of aquaculture goal scores and related plots"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true=
---

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```


```{r include = FALSE}
library(readxl)
library(ggthemes)
library(scales)
library(readr)
``` 

Here I will plot the final scores of the aquaculture subgoal.
The scores will be based on maximal yield in northern Norway.

# Load the data 
```{r}
scores_final <- read_csv("data/aquaculture_goal_final.csv")

#there is a column inserted in the front when reading the file, probably read_csv feature
scores_final_prep <- scores_final[,-1]
```

Loading also the Norwegian names of the municipalities, for plotting
```{r}
komnames_norw <- read_excel(file.path("~/github/aqua_fish_paper/data/raw/komlist_norw.xlsx"))
```

# Prepare the data
Here I merge the column of Norwegian municipalities names with aquaculture scores table. Also, i keep just the needed columns


```{r}
scores_final_prep2 <-scores_final_prep %>% 
  left_join(komnames_norw[c("Komnum", "Norw_name")], by = c("municip_number" = "Komnum")) %>% 
  select(c(year:municip, sustainability, aquaculture_score, Norw_name))
```

# Plotting the final scores
Ggplot theme for food production goal:

```{r}
theme_food <- function(...) {
  theme_stata(scheme = "s1color") +
    theme(
      axis.text.x = element_text(
        size = 8, angle = 90, hjust = 1, vjust = 0.5,
        face = "bold"
      ),
      axis.text.y = element_text(size = 10, angle = 0, face = "bold"),
      axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0), face = "bold", size = 10),
      plot.title = element_text(hjust = 0.5, vjust = 0.5, face = "bold"),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2, "cm"),
      strip.text = element_text(size = 10, face = "bold"),
      strip.background = element_rect(fill = "bisque", color ="bisque"),
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "white"),
      panel.grid.major.y = element_line(color = "grey", size = 0.7),
      legend.background = element_rect(color = "white")
    )
}

```

Define breaks for y-axis
```{r}
breaks <- seq(2005, 2018, by = 1)
labels <- as.character(breaks)
labels[!breaks%%2==0]<- ""
```

# Plotting aquaculture goal scores per municipality and county

```{r}
sc_counties <- unique(scores_final_prep2$county) 

for(i in 1:length(sc_counties))
  {
  cty <- sc_counties[i]
  
  df <- scores_final_prep2 %>%
    filter(county == cty) 
    
    
plot <- ggplot(df)+
      geom_line(aes(x = year, y = aquaculture_score), size = 0.7, color = "darkblue") +
      facet_wrap(Norw_name  ~ .) +
      scale_x_continuous(breaks = breaks, labels = labels)+
      labs(y = "Aquacultlure subgoal score", x = "") +
      ggtitle(cty)+
      theme_stata(scheme = "s1color") +
      theme_food()
           
           
            
ggsave(paste0("figs/aquaculture_final_score_", cty, ".pdf"), plot)
  
}
```

# Plotting aquacutlure sustainability scores per municipality
Here we plot just the sustainability scores (that are multiplied by production/max production)
```{r}
sc_counties <- unique(scores_final_prep2$county) 

for(i in 1:length(sc_counties))
  {
  cty <- sc_counties[i]
  
  df <- scores_final_prep2 %>%
    filter(county == cty) 
    
    
plot <- ggplot(df)+
      geom_line(aes(x = year, y = sustainability), size = 0.7, color = "black") +
      facet_wrap(Norw_name  ~ .) +
      scale_x_continuous(breaks = breaks, labels = labels)+
      labs(y = "Aquacultlure sustainability score", x = "") +
      ggtitle(cty)+
      theme_stata(scheme = "s1color") +
      theme_food()
           
           
            
ggsave(paste0("figs/aquaculture_sustainability_score_", cty, ".pdf"), plot)
  
}
```

