---
title: "Calculating final aquaculture goal"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('~/github/nor-prep/prep/src/common.R')
```

Here I combine the sustainability indicators of aquaculture (lice, momb, efcr), and production into one table.
Then I estimate the proportion of total production per year and municipality, that is sustainable.
# Load data: lice, momb, efcr and production
```{r}
lice <- read.csv("data/lice_below_thr_gapfilled.csv")
momb <-read.csv("data/mean_momb_mp_yr_gapfilled.csv")
efcr <-read.csv("data/regional_efcr_score_on_gapfilled_prod.csv")
```
And the sea area table:
```{r}
seaarea <-read_excel(file.path(dir_M[2], "Fishery/DeliveryDataFiskDir/KommArea.xlsx"))
```


# Homogenizing the data
```{r}
lice_prep <- lice %>% 
  rename(prop_below_lice = below_lice,
         lice_score = score_final) %>% 
          select(-gf) %>% 
          rename_all(., tolower)
      

momb_prep <- momb %>% 
  select(-c(momb_score, gapfilled)) %>% 
  rename(municip_number = Komnum_new,
         municip = Kommune_new,
         momb_score = momb_gapfilled,
         proportion_biomass_bad_momb = totprop_biomass_polluted) %>% 
         rename_all(., tolower)

```

# Merging the three tables
```{r}
aqua_goal_prep <- efcr %>%
  select(-gapfilled) %>%
  rename(reference_efcr = ref_efcr) %>% 
  left_join(lice_prep, by = c(
    "year",
    "municip",
    "municip_number"
  )) %>%
  left_join(momb_prep, by = c(
    "year",
    "municip",
    "municip_number"
  )) 
  
```

# Define function to calculate geometric mean of sustaianbility scores
Geometric mean does not tolerate zeros: when any terms is zero, the whole mean becomes zero. A workaround is to add 1 to each term and then subtract 1 from the resultant geometric mean. But because the scores are on the scale 0-1,  I add and subtract 0.01.
```{r}
gm_mean <- function(...) {
  x <- c(...)
  if (all(is.na(x))) {
    return(NA)
   } else if (any(x == 0, na.rm = T)) {
     mod_x <- x + 0.01
   gmmean  <- prod(mod_x, na.rm = TRUE)^(1 / sum(!is.na(mod_x))) - 0.01
   } else {
    gmmean  <- prod(x, na.rm = TRUE)^(1 / sum(!is.na(x)))
   }
   return(gmmean)
 }
```

Calculating geometric mean of sustainability scores and defining the reference - maximal produciton within the whole northern Norway in a given year.

$Score = \frac{Production_i}{max(Production_i)}*sutainability \ score$

```{r}
aqua_goal_final <-
  aqua_goal_prep %>%
  mutate(mean_sustain_score = pmap_dbl(
    list(
      efcr_score,
      lice_score,
      momb_score
    ),
    gm_mean
  )) %>% 
  group_by(year) %>% 
  mutate(maxprod = max(production_final)) %>% 
  ungroup() %>% 
  mutate(prod_prop_mp = production_final/maxprod) %>% 
  mutate(aqua_score = prod_prop_mp*mean_sustain_score*100)
           
```

# Check the distribution of the aquaculture sustainability score:
```{r}
library(ggthemes)

ggplot(data = aqua_goal_final) +
  geom_density(
    mapping = aes(x = aqua_score),
   fill = "yellow", color = "darkgoldenrod", alpha = 0.6
  ) +
  ggtitle("Distribution of the aquaculture goal score") +
  #scale_x_continuous(labels = scales::comma) +
  labs(x = "score") +
  theme(axis.text = element_text(size = 12)) +
  theme_gdocs()
```


# Calculate production per seaarea for each municipality and year

```{r}
aqua_goal_final_seaarea <-
  aqua_goal_prep %>%
  mutate(mean_sustain_score = pmap_dbl(
    list(
      efcr_score,
      lice_score,
      momb_score
    ),
    gm_mean
  )) %>% 
  left_join(seaarea[,c(2,3)], 
            by = c(
    "municip_number" = "Komnum"
  )) %>% 
  rename(seaarea = Havflate) %>% 
  mutate(prod_area = production_final / seaarea) %>% 
  group_by(year) %>% 
  mutate(maxprod_area = max(prod_area)) %>% 
  ungroup() %>% 
  mutate(prod_prop_area = prod_area / maxprod_area) %>% 
  mutate(aqua_score_area = prod_prop_area * mean_sustain_score * 100)
```

# Check the distribution of the aquaculture score, when the production is per seaarea 
```{r}
library(ggthemes)

ggplot(data = aqua_goal_final_seaarea) +
  geom_density(
    mapping = aes(x = aqua_score_area),
   fill = "darkorange", color = "darkgoldenrod", alpha = 0.5
  ) +
  ggtitle("Distribution of the aquaculture goal score") +
  #scale_x_continuous(labels = scales::comma) +
  labs(x = "score") +
  theme(axis.text = element_text(size = 12)) +
  theme_gdocs()
```

Save the table with the score based on produdction per sea area
```{r}
#write.csv(aqua_goal_final_seaarea, "./data/production_per_area_and_aqua_score.csv")
```

# Calculating aquaculture sub-goal final scores
```{r}

```




##############
Just another way to avoid zeroes in the geometric mean, the code is for python and I have not translated it to R.
https://github.com/RobertoCM/geomMeanExt/blob/master/geomMeanExt.py


	

	
