---
title: "Escapees counts per municipality"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

# Loading libraries
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r include=FALSE}
library(readr)
library(janitor)
library(stringr)
library(qdap)
library(kableExtra)
library(docxtractr)
```

# Reading escapees data tables from Word document
```{r}
escap_docx <-read_docx(file.path(dir_M[2], "Aquaculture/Escapees/Rømninger2006-2016.docx"))
escap_tbl <- docx_extract_all_tbls(escap_docx)
komlist<- read.csv("../../administrative/komlist.csv", sep = ";")
licenses <-read.csv("data/aquaculture_lisences_northnorway_2010_2018.csv")
```


```{r}
selected_tables <- map_lgl(escap_tbl, ~ nrow(.x) >= 10) 
escap_tbl_prep <- escap_tbl[selected_tables]

escap_tbl_prep2 <-  escap_tbl_prep[map_lgl(escap_tbl_prep, ~ ncol(.x) == 9)]
escap_tbl_prep2_gjenfangst <-  escap_tbl_prep[map_lgl(escap_tbl_prep, ~ ncol(.x) == 10)]
```

```{r}

escap_tbl_prep3 <- escap_tbl_prep2 %>%
  map(., ~ setNames(.x, c(
    "county", "company_name", "location_number",
    "species", "size", "escape_date",
    "escapees_count_prod1", "escapees_count_fishdir",
    "escapees_count_prod2"
  ))) %>%
  map(., ~ mutate(.x, recatched_count = rep(NA, length(county))))
```

```{r}
escap_tbl_prep3_gjenfangst <- escap_tbl_prep2_gjenfangst %>% 
  map(.,  ~ setNames(.x, c("county", "company_name",  "location_number", 
                           "species", "size", "escape_date",
                           "escapees_count_prod1", "escapees_count_fishdir",
                           "escapees_count_prod2", "recatched_count")))
```


```{r}
escap_prep4 <- do.call("rbind", c(escap_tbl_prep3, escap_tbl_prep3_gjenfangst))
```

```{r}
catch_year <- function(text) {
  lastdig <- str_extract(text, "([:digit:]){2}$")
  year = str_c("20", lastdig)
  year
}
```

```{r}
escap_prep5 <- escap_prep4 %>%
  filter(company_name != "" & county != "") %>%
  filter(county != "Fylke") %>%
  filter(county %in% c("N", "T", "F")) %>%
  mutate(year = catch_year(escape_date)) %>%
  filter(species %in% 
           c(
    str_extract("Laks", fixed(species, ignore_case = T)),
    str_extract("Rb.ørret", fixed(species, ignore_case = T))
    )) %>%
  mutate_at(
    .vars = vars(c(
      "location_number",
      "escapees_count_prod1",
      "escapees_count_fishdir",
      "escapees_count_prod2",
      "recatched_count"
    )),
    .funs = ~ str_replace(., "\\s", "")
  ) %>%
  mutate_at(
    .vars = vars(c(
      "escapees_count_prod1",
      "escapees_count_fishdir",
      "escapees_count_prod2",
      "recatched_count"
    )),
    .funs = funs(as.numeric)
  ) %>%
  rowwise() %>%
  mutate(escap_final = sum(escapees_count_prod1,
    escapees_count_prod2,
    escapees_count_fishdir,
    -recatched_count,
    na.rm = T
  ))
```


I used aquaculture registry to get the geogprahical locaiton of all the companies in the escapees data.
However, when I tired to match aquaculture registry with escapees data by location numbers, I got 13 mismatches (locaitons, not occuring in aquaculture registry, one of them is transportation boat).
All these locaitons have escapees data for 2006 - 2015. So, I assume that those are companies that shut down or were sold to other companies, or have moved to other locaitons.


# Merge escapees data with lisences data
I identified duplicates after merging escapees with municipalities in lice data: <span style="colour:blue">Håløy Havservice AS</span> (municipality must be 1903 - Harstad), 
and <span style="color:blue">Nordlaks Oppdrett AS</span> (municipality must be 1903 - Harstad). In both cases duplicates were due to Harstad merged with Bjarkoy in 2013. 
One of the <span style="color:blue">Jøkelfjord Laks AS</span> data rows did not have location and municipality, I will assume they are the same as in another data row (loc 10808 and municip 1943) 


```{r}
#get a list of unique location names
names(licenses) <- gsub("Ø", names(licenses), replacement = "O") %>%
    gsub("Å", ., replacement = "A") %>%
    str_to_lower(.)

totlocs_mp <- licenses %>%
  filter(vannmiljo == "SALTVANN") %>%
  group_by(lok_nr,lok_komnr, lok_kom) %>%
  summarize(nlocs = n_distinct(lok_nr)) 

escap_prep6 <- escap_prep5 %>% 
  mutate(location_number = as.integer(location_number)) %>% 
  left_join(totlocs_mp[,c(1,2,3)], by = c("location_number" = "lok_nr")) 
  
```
There are duplicated rows after merging with liceses data frame. The duplicates are row 65 and 71 (Håløy Havservice AS and Nordlaks Oppdrett AS).
Both companies are in Harstad which after 2013 was not merged with Bjarkoy municipality. That's why duplicates occured.
```{r}
foo <-escap_prep6[,-14]
which(duplicated(foo) == 1) 
#[1] 65 71
```


```{r}
escap_prep7 <- escap_prep6 %>% 
  select(-lok_kom) %>% 
  distinct(., location_number, escape_date, .keep_all = T) %>% 
  mutate(lok_komnr = replace(lok_komnr, location_number == 32257, "1903" )) %>% 
  mutate(lok_komnr = replace(lok_komnr, location_number == 11378, "1903" )) %>% 
  mutate(location_number = replace(location_number, company_name == "Jøkelfjord Laks AS", "10808")) %>% 
  mutate(lok_komnr = replace(lok_komnr, company_name == "Jøkelfjord Laks AS", "1943"))
```


##  Exploring missing data: which companies did not have municipality number after merging escapees and lice data?
```{r}
nas <- filter(escap_prep7, is.na(lok_komnr)) %>% 
      group_by(company_name, location_number) %>% 
      summarize(n = n()) #12 companies did not have info on locaiton, I get info by  google serach.
```

##  Loading back the missing companies data 
I have found the locations for these companies on internet.
```{r}
missing <-data.table::fread("./data/missing_companies.csv",select = c(2,3,5), encoding = "UTF-8")
```

## Getting municipality number for companies with no municipality location registered
I remove  <span style="color:blue">Hammerfest Lakseslakteri</span> and  <span style="color:blue">Victoria Lady (brønnbåt)</span> and because they don't have location numbers. 
```{r}
missing_prep <-left_join(missing, komlist[,c(1,2)], by = c("municip" = "Name")) %>% 
                mutate(location_number = as.character(location_number)) %>% 
                mutate(company = as.character(company)) %>% 
                filter(!company %in% c("Victoria Lady (brønnbåt)","Hammerfest Lakseslakteri")) 
              
          
missing_final <-filter(escap_prep7, is.na(lok_komnr)) %>% 
     group_by(company_name, location_number) %>% 
     left_join(missing_prep, by = c("company_name" = "company", "location_number")) %>% 
     ungroup() %>% 
     mutate(lok_komnr = Komnum) %>% 
     mutate(lok_komnr = as.character(lok_komnr)) %>% 
     select(-c(municip, Komnum))
```

Merging did not work for <span style="color:blue">Sjølaks AS</span> ,I suspect due to special character ø which dplyr sometimes does not recognize
```{r}

#I delete Victoria Lady (brønnbåt) - it does not have municipality
#I add Hammerfest municipality to the row of Hammerfest Lakseslakteri, but locaiton number for this company is unknown.
komlist <- komlist %>% 
  mutate(Komnum = as.character(Komnum))

escap_fin <-bind_rows(filter(escap_prep7, !is.na(lok_komnr)), 
                             missing_final) %>% 
          mutate(lok_komnr = replace(lok_komnr, company_name == "Sjølaks AS", "1943")) %>% 
          mutate(lok_komnr = replace(lok_komnr, company_name == "Hammerfest Lakseslakteri", "2004")) %>% 
          filter(!company_name == "Victoria Lady (brønnbåt)") %>% 
          left_join(komlist[,c(1,2)], by = c("lok_komnr"  = "Komnum")) %>% 
          rename("kommunenavn" = "Name") 
```

## How many locations per municipality for each year  and what is the range of escapees count?
```{r}
#Most of the municipalities had just one ovservation from a single locaiton each year. 
#Number of escapees within same municipality may vary considerabely, thus we should not average them.

escap_range <-escap_fin  %>% 
             group_by(year, lok_komnr, species) %>% 
             summarize(locperyear = n(),min_escap = min(escap_final), max_escap = max(escap_final)) 

kable(escap_range) %>% 
    kable_styling("striped", full_width = T) %>% 
     column_spec(1:6,bold = T) 
```

# Aggreagte escapees counts per municiaplity per year  
```{r}
escap_pmp <-escap_fin  %>% 
             #mutate_at(c(8:12), as.numeric) %>% 
             group_by(year, lok_komnr, kommunenavn) %>% 
             summarize(escaped_permp = sum(escap_final))

#write_csv(escap_pmp,"./data/escapees_per_yrmpspp.csv")
```
