---
title: "Lice counts per week and municipality"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source("~/github/nor-prep/prep/src/common.R")
```

```{r}
library(varhandle)
library(stringr)
```

# Load data
```{r}
lus <- read.csv(file.path(dir_M[2], "Aquaculture/Lice/lakselus_per_fisk_utf8.csv"), header = TRUE, encoding = "UTF-8")
municip <- read.table("../../administrative/komlist.csv", header = TRUE, sep = ";")
municip <-filter(municip, !Name == "NonCoast")
lus_2005 <-read.csv("./data/lice_vers2_2005_2011_formatted.csv", sep = ",")
lisences <-read.csv("./data/aquaculture_lisences_northnorway_2010_2018.csv")
```


# Finding municipality and lice threshold info for some of the locations
There are 89 locations for which no municipality and region specified, also the lice threshold is missing for those.
I will check if the Aquaculture registry have information on these locaitons ("lisences" dataframe).

```{r}
#list of the locaitons with missing info:
lus <- lus %>%  
  rename(Ar = X.r)

missing_mppinfo <-lus %>% 
  filter(is.na(Kommunenummer)) 

foo <-lus %>% 
  group_by(Ar) %>% 
  mutate(lusegrense = as.numeric(as.character(Lusegrense.uke)))

#cheking if lice threshold changed during 2012-2019: yes, since 2017 lice threshold is 0.2 (in spring) and 0.5 otherwise
foo2 <- foo %>% 
  group_by(Ar) %>% 
  summarize(grensemin = min(lusegrense, na.rm = T),
            grensemax = max(lusegrense, na.rm = T))
foo2
```


Merging lus_prep with lisences dataframe - to find municipaities where they are missing. 
I could find info only for 2 out of 89 locations with missing municipality and I add municipalities for them mannually in the code "Preliminary data cleaning".

```{r}
locsmpp <- lisences  %>% 
  group_by(Year, LOK_NR, LOK_KOMNR, LOK_KOM) %>% 
  summarize(count=n()) 

missing_mppinfo_filled <- missing_mppinfo %>% 
  group_by(Ar,Lokalitetsnummer) %>% 
  summarize(locscount = n()) %>% 
  left_join(locsmpp[,c("Year","LOK_NR", "LOK_KOMNR", "LOK_KOM")], by = c("Lokalitetsnummer" = "LOK_NR","Ar"="Year"))

map(missing_mppinfo_filled, ~ sum(is.na(.)))
```

# Preliminary data cleaning
```{r}
colnames(lus)[c(2, 19, 20)] <- c("Ar", "ProduksjonsomradeID", "Produksjonsomrade")
lus_prep <- lus %>%
  mutate(Kommunenummer = replace(Kommunenummer, Lokalitetsnummer == "10893", "1835" )) %>% 
  mutate(Kommunenummer = replace(Kommunenummer, Lokalitetsnummer == "11087", "1836" )) %>% 
  filter(Kommunenummer >= 1800 & Kommunenummer < 2100) %>%
  mutate(Kommunenummer = replace(Kommunenummer, Kommunenummer == "1915", "1903")) %>%
  mutate(Kommunenummer = replace(Kommunenummer, Kommunenummer == "1901", "1903")) %>%
  mutate(Kommunenummer = as.numeric(Kommunenummer)) %>%
  left_join(municip[, c(1, 2)], by = c("Kommunenummer" = "Komnum")) %>%
  mutate(Kommune = Name) %>%
  select(-Name)
```


```{r}
#Lice threshold is 0.2 from week 21 to week 26 in Northern Norway since 2017.
replace_licethr <- function(num, year, week=NULL) {
  if (year < 2017) {
    if (num < 0.5) {
      overgrense <- "Nei"
    } else {
      overgrense <- "Ja"
    }
  } else if (year >= 2017 & week < 21|week > 26 ) {
    if (num < 0.5 ) {
      overgrense <- "Nei"
    } else {
      overgrense <- "Ja"
    }
  } else {
    if (num < 0.2) {
      overgrense <- "Nei"
    } else {
      overgrense <- "Ja"
    }
  }
  overgrense
}

```


```{r}
#filter out rows where lice threshold is unknown.
#(the responsible engineer said it was probably a mistake in data logging)
foo3 <-filter(lus_prep, Over.lusegrense.uke == "Ukjent" & !is.na(Lus.i.bevegelige.stadier))
foo4 <- foo3 %>% 
  mutate(Over.lusegrense.uke = pmap_chr(list(Lus.i.bevegelige.stadier, Ar, Uke), ~ replace_licethr(..1, ..2, ..3)))


#and rbind with the other:
lus_prep2 <-lus_prep %>% 
  filter(!is.na(Lus.i.bevegelige.stadier)) %>%
  filter(Over.lusegrense.uke != "Ukjent") %>% 
  bind_rows(foo2)
```


# Further data cleaning: remove inactive localities and those that did not count lice
How many localities are there in total per municipality per year?
We may want to know how many locations were active per municiality, year and week, to be aware how many locations actually report lice count.
```{r}
total_locations <- lus_prep %>%
  group_by(Kommune, Kommunenummer,Ar, Uke) %>%
  summarize(tot_loc = n_distinct(Lokalitetsnummer)) %>%
  mutate(key = paste(Kommune,"_" , Ar, "_", Uke))

#write.csv(total_locations,"data/number_lice_observations_per_mp.csv") 
```


Select only locations that were counting lice. 
Change column names to English.
```{r}
lus_prep3 <- lus_prep2 %>%
  filter(Har.telt.lakselus == "Ja") %>%
  select(-c(
    "Voksne.hunnlus", "Lus.i.bevegelige.stadier", "Fastsittende.lus", "Brakklagt", "Sj.temperatur",
    "Har.telt.lakselus", "Lusegrense.uke", "ProduksjonsomradeID", "Produksjonsomrade"
  )) %>%
  rename(
    Week = Uke, Year = Ar, Locality_number = Lokalitetsnummer, Locality_name = Lokalitetsnavn,
    Municip_number = Kommunenummer, Municip_name = Kommune, County_number = Fylkesnummer, County = Fylke,
    Above_threshold = Over.lusegrense.uke
  )
```


# Estimate average % of weeks localities are below lice threshold (new data)
```{r}
below_threshold_new <- lus_prep3 %>%
  mutate(Above_thr_new = case_when(
    Above_threshold == "Nei" ~ "1",
    Above_threshold == "Ja" ~ "0"
  )) %>%
   mutate(Above_thr_new = as.numeric(as.character(Above_thr_new))) %>%
  group_by(Year, Locality_number, Municip_number, Municip_name) %>% 
  summarise(totweeks = n(),
            below_weeks = sum(Above_thr_new),
            prop_weeks_below = below_weeks/totweeks)
```
About halth of the locations reported salmon lice only once a year. Barentswatch.no explains that missing lice count maybe because the locality did not have fish in the pens during those weeks or they just did not report. I will select only the locations that reported at least 25 weeks over a year (about half of all locations).

```{r}
below_threshold_new2 <- below_threshold_new %>% 
  filter(totweeks >= 25) 

#check how many locations left per municipality: most of municipalities had 1-5 locations,
#counting lice at least 25 weeks a year.
locs_counted_25weeks <- below_threshold_new2 %>% 
  group_by(Year, Municip_name) %>% 
  summarize(locscount = n())

#Final table
below_lice_mpp_new <-below_threshold_new2 %>% 
  group_by(Year, Municip_number, Municip_name) %>% 
  summarize(below_lice = mean(prop_weeks_below))
```


# Cleaning older lice data (2005-2011)
```{r}
lus_old_prep <- lus_2005 %>% 
  select(-c("Farmname", "Sitting_lice"))

#the list of locaitons with no municipality info
lus_old_missinglocs <-lus_old_prep %>% 
  filter(Latlon == "(-999.0,-999.0)") %>% 
  group_by(Location) %>% 
  summarize(count = n())

#count the number of locations in the new data that have lice counts
lus_new_locs <-lus %>% 
  group_by(Lokalitetsnummer, Lokalitetsnavn, Lat, Lon) %>% 
  summarise(count = n())

#merge the old and new datasets to see for which locations I can find lat and lon info - only for 63 (extra check - for another colleague)
locnames_merged <-lus_old_missinglocs %>% 
  left_join(lus_new_locs, by = c("Location" = "Lokalitetsnummer")) %>% 
  filter(!is.na(Lokalitetsnavn)) %>% 
  select(-c("count.x", "count.y"))
```


```{r}
#location 10933 had no municipality info, it is not in the Aquaculture registry
lus_old_prep2 <- lus_old_prep %>% 
  filter(!is.na(Municipality)) %>% 
  select(-c(What, Latlon))
```

# Estimate average % of months when localities are below lice threshold (old data)
```{r}
#check NAs in the old data
map(lus_old_prep2, ~sum(is.na(.)))

#delete all rows where count of adult females is NA,
#remove duplicated rows 
#summarize per location months below lice threshold (0.5)

below_thr_old <-lus_old_prep2 %>% 
  filter(!is.na(Adult_females)) %>% 
  filter(!duplicated(.)) %>% 
  mutate(Avobe_threshold = ifelse(Adult_females >= 0.5, 0, 1)) %>% 
  group_by(Year, Municipality, Location) %>% 
   summarise(totmonths = n(),
            below_months = sum(Avobe_threshold),
            prop_months_below = below_months/totmonths)
  
```

Again, about halth of the locations reported lice less than 6 times a year (once a month).
I will select only the locations that reported at least 6 times a year.

```{r}
below_thr_old2 <- below_thr_old %>%
  filter(totmonths >= 6)

# check how many locations left per municipality: most of municipalities had 1-6 locations,
# counting lice at least 6 months in a year.

locs_counted_6months <- below_thr_old2 %>%
  group_by(Year, Municipality) %>%
  summarize(locscount = n())

# Final table
below_lice_mpp_old <- below_thr_old2 %>%
  group_by(Year, Municipality) %>%
  summarize(below_lice = mean(prop_months_below)) %>%
  left_join(municip[, c(1, 2)], by = c("Municipality" = "Komnum")) %>%
  rename(Municip_number = Municipality, Municip_name = Name) %>%
  select(Year, Municip_number, Municip_name, below_lice)
```


# Merge old and new data tables
```{r}
below_lice_full <- bind_rows(below_lice_mpp_old, below_lice_mpp_new)
#check how many observations per municipality we get:
years_per_mpp <- below_lice_full %>%  group_by(Municip_name) %>% summarize(count = n())
#none of the municipalities had data for all 2005-2018 years, at most we get observations for 14 years. That makes sense since some of the industries came into existince after 2005.
```


# Gapfilling missing data in lice dataset
## Check which counties did not have lice data at all
```{r}
anti_join(municip, below_lice_full, by = c("Name" = "Municip_name")) %>% 
  arrange(Name)
#13 counties 
```
All these municipalities  had no information in lice, and they also did not have data on momb, so likely there was no produciton at all.

## Define a function to gapfill

```{r inlcude=FALSE}
#Version 1: take the mean of all available years


# gapfill <- function(df, miss_col) {
#     if (all(is.na(df[, miss_col]))) {
#     warning("there are no numeric values in the data", call. = FALSE)
#     finaldf <-df 
#     } else {
#       cleandat <- drop_na(df, miss_col)
#       replacement <- mean(cleandat[,miss_col], na.rm = T)
#       replaced <- replace_na(df[,miss_col], replacement)
#       finaldf <- cbind(df, replaced)
#     }
#   finaldf
# }

```

Version 2: if more than 7 years of data are available, take the mean of the latest 5
```{r}
gapfill <- function(df, miss_col) {
  finaldf <- df
  cleandat <- drop_na(df, miss_col)
  if (all(is.na(df[, miss_col]))) {
    warning("there are no numeric values in the data", call. = FALSE)
  } else if (nrow(cleandat >= 7)) {
    replacement <- mean(tail(cleandat, 5)[, miss_col], na.rm = T)
    replaced <- replace_na(df[, miss_col], replacement)
    finaldf <- cbind(df, replaced)
  } else {
    replacement <- mean(cleandat[, miss_col], na.rm = T)
    replaced <- replace_na(df[, miss_col], replacement)
    finaldf <- cbind(df, replaced)
  }
  finaldf
}
```

## Replacing missing years of data
```{r warning=FALSE} 
lice_municips <- below_lice_full %>% 
  mutate(Komname = droplevels(Municip_name)) %>% 
  split(., .$Komname)
```


## Compiling final dataframe
```{r}
allyears <-data.frame(Year = seq(2005, 2019, by = 1))

lice_replaced_prep <- map(lice_municips, ~ as.data.frame(.x)) %>%
  map(., ~ full_join(.x, allyears, by = "Year")) %>% 
  map(., ~ mutate(.x, Komnum_new = rep(Municip_number[1], length(Municip_number)))) %>% 
  map(., ~ mutate(.x, Kommune_new = rep(Komname[1], length(Komname)))) %>% 
  map(., ~ select(.x,-c(Municip_number, 
                        Municip_name, 
                        Komname
                        ))) %>% 
  map(., ~ gapfill(.x,2)) 

lice_gapfilled <- do.call("rbind", lice_replaced_prep) %>%
  rename(Municip_number = Komnum_new, Municip = Kommune_new, score_final = replaced) %>%
  mutate(gf = ifelse(is.na(below_lice), "y", "n"))

#write_csv(lice_gapfilled, "data/lice_below_thr_gapfilled.csv")
```
