---
title: "Proporiton of localities with above-thresholdp lice count"
author: "Marina Espinasse"
date: "`r Sys.Date()`"
output:
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
source("~/github/nor-prep/prep/src/common.R")
```

```{r}
library(varhandle)
library(stringr)
```
# Load data
```{r}
# lice counts
lus <- read.csv(file.path(dir_M[2], "Aquaculture/Lice/lakselus_per_fisk_utf8.csv"), header = TRUE, encoding = "UTF-8")
municip <- read.table("../../administrative/komlist.csv", header = TRUE, sep = ";")
```

# Preliminary data cleaning
```{r}
colnames(lus)[c(2, 19, 20)] <- c("Ar", "ProduksjonsomradeID", "Produksjonsomrade")
lus_prep <- lus %>%
  filter(Kommunenummer >= 1800 & Kommunenummer < 2100) %>%
  mutate(Kommunenummer = replace(Kommunenummer, Kommunenummer == "1915", "1903")) %>%
  mutate(Kommunenummer = replace(Kommunenummer, Kommunenummer == "1901", "1903")) %>%
  mutate(Kommunenummer = as.numeric(Kommunenummer)) %>%
  left_join(municip[, c(1, 2)], by = c("Kommunenummer" = "Komnum")) %>%
  mutate(Kommune = Name) %>%
  select(-Name)
```

# Further data cleaning: remove inactive localities and those that did not count lice
## How many localities are there in total per municipality per year
We may want to know how many locations were active per municiality, year and week to be aware how many locations actually report lice count.
```{r}
total_locations <- lus_prep %>%
  group_by(Kommune, Kommunenummer,Ar, Uke) %>%
  summarize(tot_loc = n_distinct(Lokalitetsnummer)) %>%
  mutate(key = paste(Kommune,"_" , Ar, "_", Uke))
```

# Prepare final dataset with total number of localities with over-threshold lice count
Select only locations that were counting lice. I do not filter out inactive locaitons, because obviously, they did not count lice.
Change column names to English.
```{r}
lus_final <- lus_prep %>%
  filter(Har.telt.lakselus == "Ja") %>%
  select(-c(
    "Voksne.hunnlus", "Lus.i.bevegelige.stadier", "Fastsittende.lus", "Brakklagt", "Sj.temperatur",
    "Har.telt.lakselus", "Lusegrense.uke", "ProduksjonsomradeID", "Produksjonsomrade"
  )) %>%
  rename(
    Week = Uke, Year = Ar, Locality_number = Lokalitetsnummer, Locality_name = Lokalitetsnavn,
    Municip_number = Kommunenummer, Municip_name = Kommune, County_number = Fylkesnummer, County = Fylke,
    Above_threshold = Over.lusegrense.uke
  )
```

# Estimate total number and proportion of locaitons above lice threshold per week
```{r}
above_threshold_locs <- lus_final %>%
  mutate(Above_thr_new = case_when(
    Above_threshold == "Nei" ~ "0",
    Above_threshold == "Ja" ~ "1"
  )) %>%
  mutate(Above_thr_new = as.numeric(as.character(Above_thr_new))) %>% 
  group_by(Week, Year, Municip_number, Municip_name) %>%
  summarize(locs_above_thr_count = sum(Above_thr_new), locs_active = n(), locs_above_thr_prop = locs_above_thr_count / locs_active) 
  
#write.csv(above_threshold_locs, "data/count_locations_above_lice_threshold_weekly.csv")
```

# Estimate average number of localities with high lice count per municipality*year
```{r}
above_threshold_yearly_average <- above_threshold_locs %>%
  group_by(Municip_number, Municip_name, Year) %>%
  summarize(average_locs_above_thr = mean(locs_above_thr_count), average_prop_locs_above_thr = mean(locs_above_thr_prop))

#write.csv(above_threshold_yearly_average, "data/locs_above_thr_yearly_averages.csv")
```

# Add the column of total number of localities (including inactive that did not count lice)
```{r}
above_threhold_locs_tototal <- above_threshold_locs %>%
  mutate(key = paste(Municip_name, "_", Year, "_", Week)) %>%
  left_join(total_locations[, c(5, 6)], by = "key")

#write.csv(above_threhold_locs_tototal, "data/locations_above_lice_thr_compare_tototal.csv")
```

