---
title: "Calculating the proportion of fish lost during the produciton"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r libraries, results="hide"}
library(ggthemes)
library(janitor)
library(DT)
```

# Loading data
```{r}
akva_final <- read.table("data/total_aquaculture_production_n_efcr.csv", header = TRUE, sep = ",")
tap_0518 <-read.csv(file.path(dir_M[2],"Aquaculture/Production/production_unlocked_concatenated_2005_2018/akva_tap_formatted_2005_2018.csv"))
```


# Calculate total amount of lost fish
```{r}
losttotal_prep <- tap_0518 %>% 
  clean_names() %>% 
  rename(year = aar) %>% 
  select(-c(fylkesnr, fylke, kommune)) %>% 
  mutate(utkast_kg = tap_utkast*5)
```


# Calculate the amount and the proportion of lost fish 
We calculate the proporiton of lost fish as follows: 
(Discard + Sum of lost in cages) / (Harvest of both species - Discard).

I assumed that

1) When the harvest is zero but total lost fish is above zero, the proportion is 1.

2) When the final harvest was negative (4 occasions when discard was larger then harvest),  or the amount of lost was negative, the proportion of lost is NA.

3) When the sum of lost fishes was larger then harvest (this is possible: fish was not yet ready for harvest but got lost), the proportion of lost fish is 1.

```{r}
prop_lost <- function(lost, prod){
  if (prod == 0 & lost == 0) {
    proportion = NA 
  } else if (prod < 0 | lost < 0) {
    proportion = NA
  } else if (prod == 0 & lost > 0) {
    proportion = 1
  } else {
    proportion = lost/prod
  }
  proportion
}
```


```{r}
lost_permp <- akva_final %>%
  left_join(losttotal_prep, by = c("year", "kommunenr")) %>%
  mutate_at(
    c(
      "utkast_kg",
      "tap_d_dfisk",
      "tap_r_mming",
      "tap_annet"
    ),
    funs(replace(., is.na(.), 0))
  ) %>% 
mutate(sum_lost_cages = pmap_dbl(list(
    tap_d_dfisk,
    tap_r_mming,
    tap_annet
  ), sum, na.rm = T) * 2) %>%
  mutate(sum_lost_total = pmap_dbl(list(
    sum_lost_cages,
    utkast_kg), sum, na.rm = T)) %>% 
  select(
    year,
    kommune,
    harvested_salmon,
    harvested_trout,
    utkast_kg,
    sum_lost_cages,
    sum_lost_total
  ) %>%

  mutate(harvested_final = pmap_dbl(list(
    harvested_salmon,
    harvested_trout
  ), sum) - utkast_kg) %>% 
  mutate(prop_lost = pmap_dbl(list(
    sum_lost_total,
    harvested_final
    ),  prop_lost))


```


