---
title: "Mapping the escaped fish sustainability index"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r, include = FALSE}
library(here)#using library "here" to specify paths to files
library(sf)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
```

# Loading escapeed index data and municipalities shapes and names
```{r}
escapees <- read.csv(here("prep","food_provision", "mariculture", "data", "escapees_score_per_mcp.csv"))
allkommuner <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
komlist <-read.csv("/Users/marinaespinasse/github/nor-prep/prep/administrative/komlist.csv", sep = ";")
```

# Preparing map layers
Prepare the world data for the base map, and transform the municipalities shapes to an sf object.
```{r }
komlist <-filter(komlist, !Name == "NonCoast")

komlist_prep <- komlist %>% 
  mutate(Komnum = replace(Komnum,
                          Komnum == 1903,
                          1901))

ohikom <-allkommuner[allkommuner$KOMM %in% komlist_prep$Komnum,] 
```


```{r }
world <- ne_countries(scale = "medium", returnclass = "sf")

kommuner.pol <- spTransform(ohikom, CRS("+proj=longlat +ellps=GRS80"))
kommunersf <- sf::st_as_sf(ohikom, coords= c("long", "lat"), agr="identity")

kommunersf_prep <- kommunersf %>% 
  mutate(KOMM = replace(KOMM,
                        KOMM == 1901,
                        1903))
```
Now I need to merge shapes of municipalities and the scores of escapees index per municipality
```{r}
plot_escapees <- escapees %>% 
  left_join(
    kommunersf_prep,
    by = c("municip_number" = "KOMM")
  ) %>% 
  st_as_sf(.)
```

# Maps
```{r}
base_map <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_escapees,
    mapping = aes(fill = escap_score_mcp),
    color = "white",
    size = 0.2,
    alpha = 0.8
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_gradient(low = "red", 
                      high = "green",
                      limits = c(0.6,1),
    name = "Escapees index scores",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top"
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = ""
  ) 
```

Just another version - using a 3-color gradient:
```{r}
base_map2 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_escapees,
    mapping = aes(fill = escap_score_mcp),
    color = "dimgray",
    size = 0.4,
    alpha = 0.7
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_gradientn(colors = c("red", "yellow", "green"),
                      limits = c(0.6,1.00),
                      breaks = c(0.60, 0.7,0.8,0.9, 1),
                      labels = c("0.64", "0.70", "0.80", "0.90","1.00"),
                      
    name = "Escapees index scores",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top"
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = ""
  ) 
```

Add municipalities'numbers. I remove two muicipalities, because their numbers cannot be visitble on the map, they are too small
```{r}
komnames <- cbind(plot_escapees,st_coordinates(st_centroid(plot_escapees)))
komnames_prep <- filter(komnames, !municip_number %in% c(1919, 1920))
```


```{r}
main_map <- base_map2 +
  geom_sf_text(data = komnames_prep, 
               aes(X,Y, label= municip_number),fontface = "bold", size = 2)
  
```

```{r}
ggsave(here("prep","food_provision", "mariculture", "figs", "escapees_scores_map.pdf") ,
       main_map, width = 10, height = 9, dpi = 300)
```

