---
title: "Mapping aquaculture production"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, 
    out.width = "100%",
    dpi = 300,
    fig.width = 8,
    fig.height = 6,
    strip.white = T)
source('~/github/aqua_fish_paper/src/common.r')
```

```{r}
library(mapdata)
library(maptools)
library(rgdal)
library(rgeos)
library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
```


# Loading data
## Municipalities shape files
```{r}
allkommuner <-readOGR(dsn = "/Users/marinaespinasse/github/nor-prep/prep/administrative/raw/shapefiles/NO_AdminOmrader")
komlist <-read.csv("/Users/marinaespinasse/github/nor-prep/prep/administrative/komlist.csv", sep = ";")
```

# Aquaculture annual production data
NB! Aquaculture production files are stored in the closed repository *nor-data*". 
I use the production data that is gapfilled: negative production values are replaced using LOCF procedure.
Load also the list of municipalities with their Norwegian names (with Norwegian special characters).

```{r}
aqua_prod <- vroom::vroom(file.path("~/github/nor-data/aquaculture/gapfilled_aquaculture_production.csv"))
komnames_norw <- read_excel(file.path("../data/raw/komlist_norw.xlsx"))
```


# Preparing mapping layers 
In the list of municipalities, Harstad has number 1903 (after merging with Bjarkoy) but in the shapefile, the number is old - 1903. I fix that manually.Later for mapping, I will need to change this number back to 1903.
```{r }
komlist <-filter(komlist, !Name == "NonCoast")
komlist_prep <- komlist %>% 
  mutate(Komnum = replace(Komnum,
                          Komnum == 1903,
                          1901))

ohikom <-allkommuner[allkommuner$KOMM %in% komlist_prep$Komnum,] 
```

Now ohikom object has 81 rows = 81 municipalities, as it should.

```{r }
world <- ne_countries(scale = "medium", returnclass = "sf")
kommuner.pol <- spTransform(ohikom, CRS("+proj=longlat +ellps=GRS80"))
kommunersf <- sf::st_as_sf(ohikom, coords= c("long", "lat"), agr="identity")

kommunersf_prep <- kommunersf %>% 
  mutate(KOMM = replace(KOMM,
                        KOMM == 1901,
                        1903))
```

## Preparing the data for plotting
I divide the production (kg) by 1000 to have it in tons.
```{r }
aquaprod05 <-filter(aqua_prod, year == "2005")

plot_aquaprod05_prep <- aquaprod05 %>% 
  mutate(production_tons = production_final/1000) %>% 
  left_join(
  kommunersf_prep, 
  by = c("municip_number"="KOMM")) %>% 
  st_as_sf(.)
 
```

There are 60 municipalities that had aquaculture production in 2005.


Production in 2005 ranges from 177 to 18557 tons.
```{r}

intervals05 <-c(0, 4000, 8000, 12000, 16000, 20000)#intervals for year 2005 only

labels05 <- imap_chr(intervals05, function(., id){
  return(paste0(formatC(intervals05[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals05[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels05 <- labels05[1:length(labels2) - 1]


plot_aquaprod05_prep2  <- plot_aquaprod05_prep %>% 
  mutate(prod_quantiles = cut(production_tons,
                               breaks = intervals05,
                               labels = labels05,
                               include.lowest = T))
```

For adding the numbers of municipalities on the map, we'll need centroids of municipalities' polygons.
Municipalities 1919 and 1920 are too small, I will denote them with asterisks later.
```{r}
komnames05 <- cbind(plot_aquaprod05_prep2,st_coordinates(st_centroid(plot_aquaprod05_prep2)))
komnames_prep05 <- filter(komnames05, !municip_number %in% c(1919, 1920))
```

Adding color palette and alpha level 0.7 to it:
```{r}
addalpha <- function(colors, alpha=0.8) {
  r <- col2rgb(colors, alpha=T)
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  return(rgb(r[1,], r[2,], r[3,], r[4,]))
}

```

```{r}
cols <- addalpha(brewer.pal(5,"Reds"))
```

## Map for year 2005
```{r}
base_map05 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_aquaprod05_prep2,
    mapping = aes(fill = prod_quantiles),
    color = "white",
    size = 0.2
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols,
    name = "Aquaculture yield, tons",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top"
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2005"
  ) 

```

# Adding municipalities borders on land
```{r}
main_map05 <- base_map05 +
geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4) +
geom_sf_text(data = komnames_prep05, aes(X,Y, label= municip_number),fontface = "bold", size = 2) 
main_map05
```
## Map for year 2018
```{r }
aquaprod18 <-filter(aqua_prod, year == "2018")

plot_aquaprod18_prep <- aquaprod18 %>% 
  mutate(production_tons = production_final/1000) %>% 
  left_join(
  kommunersf_prep, 
  by = c("municip_number"="KOMM")) %>% 
  st_as_sf(.)
 
```

There are 65 municipalities that had aquaculture production in 2018.


Production in 2018 ranges from 358 to  31650 tons.
```{r}
intervals18 <-c(0, 8000, 16000, 24000, 32000)#intervals for year 2018 only

labels18 <- imap_chr(intervals18, function(., id){
  return(paste0(formatC(intervals18[id], format="f", big.mark="," , digits = 0),
                             "",
                             " – ",
                             formatC(intervals18[id + 1], format="f", big.mark="," , digits = 0),
                             ""))
})

labels18 <- labels18[1:length(labels18) - 1]


plot_aquaprod18_prep2  <- plot_aquaprod18_prep %>% 
  mutate(prod_quantiles = cut(production_tons,
                               breaks = intervals18,
                               labels = labels18,
                               include.lowest = T))
```

```{r}
komnames18 <- cbind(plot_aquaprod18_prep2,st_coordinates(st_centroid(plot_aquaprod18_prep2)))
komnames_prep18 <- filter(komnames18, !municip_number %in% c(1919, 1920))
```

```{r}
base_map18 <- ggplot() +
  geom_sf(
   data = world, 
   fill = "antiquewhite1", 
   color = "dimgray", 
   size = 0.3) +
   #coord_sf(xlim = c(10, 32), ylim = c(65, 72)
  theme(axis.text = element_text(size = 17),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15,  face = "bold"),
        panel.background = element_rect(fill = "aliceblue") ,
        panel.grid.major = element_line(color = "#dbdbd9", linetype = "dashed"),
        plot.title = element_text(hjust = 0.5, vjust = 0.5, size = 17, face = "bold")
  ) +
  geom_sf(
    data = plot_aquaprod18_prep2,
    mapping = aes(fill = prod_quantiles),
    color = "white",
    size = 0.2
  ) +
  coord_sf(xlim = c(8, 32), ylim = c(64, 72), expand = FALSE) +
  labs(x = "Longitude", y = "Latitude") +
  scale_fill_manual(
    values = cols,
    name = "Aquaculture yield, tons",
    guide = guide_legend(
      keyheight = unit(8, units = "mm"),
      title.position = "top"
    )
  ) +
  labs(
    x = NULL,
    y = NULL,
    title = "2018"
  ) 

```

```{r, dpi = 250, fig.height=8, fig.width=8}
main_map18 <- base_map18 + 
geom_path(kommuner.pol, mapping = aes(x = long, y = lat, group = id), color = 'dimgray',  size = .4) + geom_sf_text(data = komnames_prep18, aes(X,Y, label= municip_number),fontface = "bold", size = 2) 
main_map18
```
## Save both maps
```{r}
library(patchwork)
```


```{r}
grDevices::cairo_pdf("../figs/aquaculture_yield_2005_2018.pdf", width = 10, height = 12, fallback_resolution = 300)
main_map05/main_map18
```
