---
title: "Mom-b examination of aquaculture localities"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r message=FALSE, warning=FALSE, include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r libraries, results="hide"}
library(ggthemes)
library(janitor)
library(DT)
library(zoo)
```

# Load mom-b examination data and municipalities' names
NB!I use only the dataset send by Fisheries directorate on July 2nd, 2019.
```{r}
load_by_sheet <-function(x, page = 1) {
  read_excel(file.path(x),sheet = page)
}
path <-file.path(dir_M[2],"Aquaculture/MOMB_2019/B-undersøkelser_Nordland_Troms_Finnmark_kommune.xlsx")
page <- list(5,6,7)
allcounties <- map2(path, page, load_by_sheet) 


momb_prep <-  allcounties %>% 
  do.call("rbind", .) %>% 
  clean_names()
```

```{r}
akvareg <-load_by_sheet(path, page = 4) 
akvareg <- akvareg %>%  mutate(LOK_NR = as.numeric(LOK_NR))
licenses <-read.csv("data/aquaculture_lisences_northnorway_2010_2018.csv")
komlist <-read.csv("../../administrative/komlist.csv", sep = ";")
komlist <-filter(komlist, !Name=="NonCoast") %>%  
    mutate(Name = fct_drop(as.character(Name)))
```

# Further preparation of momb data
```{r}
momb_prep2 <-  momb_prep %>% 
  mutate(Year = lubridate::year(dato_provetaking)) %>% 
  select(-c("mom_type", "reference")) %>% 
  rename("Kommune" = "akvareg_lok_kom", "Komnum" = "akvareg_lok_komnr") %>% 
  left_join(komlist[,c(1,2)], by = "Komnum") %>% 
  mutate(Kommune = Name) %>% 
  select(-Name)
  

map(momb_prep2, ~sum(is.na(.x))) #Still 93 observations with missing info on the municipality 

momb_prep3 <- momb_prep2 %>% 
  filter(!is.na(Kommune)) 
```
Regarding localities with no municipality info: they are either shut down or were not yet given licence to start the production (commented by Fisheries directorate). 
Thus, I delete these locations. 

## Check how many examined aquaculture localities per municipality each year and compare to total
```{r}
names(licenses) <- gsub("Ø", names(licenses), replacement = "O") %>%
    gsub("Å", ., replacement = "A") %>%
    str_to_lower(.)

totlocs_mp <- licenses %>%
  filter(vannmiljo == "SALTVANN") %>%
  group_by(year, till_komnr, till_kom) %>%
  summarize(nlocs = n_distinct(lok_nr)) %>% 
  rename(Year = year, Komnum = till_komnr,Kommune = till_kom)

#Examinations per year per municipality
examinations_year <- momb_prep3 %>% 
  group_by(Year, Kommune, Komnum) %>% 
  summarize(examinations = n_distinct(loknr)) %>% 
  left_join(totlocs_mp, by = c("Year", "Komnum")) #NB!! merge here by locaiton mp, not tillateslse mp

#Compare the number of examinations to the total number of registered localities
#There are no cases when all registered localities were examined.
examinations_filtered <- filter(examinations_year, !is.na(nlocs))
examinations <-examinations_filtered$examinations
totals <-examinations_filtered$nlocs

for (i in length(examinations)) {
 print(examinations[examinations == totals])
}

#write.csv(examinations_filtered, "data/number_momb_examinations_per_mp.csv")

```

# Which counties are not in the momb data?
```{r}
anti_join(komlist[,c(1,2)], momb_prep3, by="Komnum") %>% 
            datatable(. , options=list(pageLength=15), rownames = FALSE,
                      class = 'cell-border stripe compact')
```

# How many years of momb obsersvations per municipality?
```{r}
momb_prep3 %>% 
  group_by(Kommune) %>% 
  summarize(nyears = n_distinct(Year)) %>% 
  datatable(rownames = FALSE,  class =  "stripe compact")
```

Also checking data density for municipality and year in a dataframe
```{r}
bympp <-split(momb_prep3, momb_prep3$Kommune) 
 datatable(bympp[["Vevelstad"]])
```

# Finding maximal allowed biomass for each locality
To find MAB, i use licenses table again.
```{r}

locs_unique <- licenses %>% 
  clean_names(.) %>% 
  group_by(year, lok_nr, lok_kap, lok_enhet) %>% 
  summarize(n = n())
  
#remove years 2004 and 2019 from the momb dataset, those years have incomplete data.
#There is no data on maximal capacity before 2010,and for 2016,so  years 2005-2010 and 2016 need to be gapfilled then. 
#Even in 2010-2018, about 220 localities did not have information on the capacity (not sure why, bad quality of data I think)

#So, I delete these years from the df.
momb_prep4 <- momb_prep3 %>%
  filter(Year >=2010 & Year < 2019) %>% 
  left_join(locs_unique[, c(1,2,3,4)], by = c("Year"="year", "loknr" = "lok_nr")) %>% 
  filter(!is.na(lok_kap))
```

# Preparing final  table: average momb per municipality and year

```{r}
momb_prep5 <- momb_prep4 %>% 
  group_by(Komnum,Kommune, Year) %>% 
  mutate(tot_capacity_mp = sum(lok_kap)) %>% 
  ungroup() %>% 
  mutate(prop_biomass_polluted = ifelse(miljo_tilstand %in% c(3,4), lok_kap/tot_capacity_mp, 0)) %>% 
  group_by(Komnum,Kommune, Year) %>% 
  summarise(totprop_biomass_polluted = sum(prop_biomass_polluted),
            momb_score = (1-totprop_biomass_polluted))
```



# Estimating running mean of the momb score per municipality
Remove municipalities with no data
```{r}
momb_bymp <- split(momb_prep5, momb_prep5$Kommune) 


#check if there are municipalities with empty rows.
#There are 16 of such municipalities (same as for lice plus Leirfjord, Evenes, and Vaeroy)
empties<- map(momb_bymp, ~ isTRUE(length(.x$momb_score)==0))
empty_municips <-unlist(empties) 

#the list of municipalities that did not have any momb data
empties_list <- unlist(empties) %>% 
 stack(.) %>% 
 filter(values==TRUE)

momb_bymp_cleaned <-momb_bymp[!empty_municips]
```


Filter out municipalities with more than 5 years of data
```{r}
long_data <- map(momb_bymp_cleaned, ~ length(.x$momb_score)) %>% 
  unlist(.) %>% 
  stack(.) %>% 
  filter(values >= 5) %>% 
  select(-values) %>% 
  mutate(ind = as.character(ind))

short_data <- map(momb_bymp_cleaned, ~ length(.x$momb_score)) %>% 
  unlist(.) %>% 
  stack(.) %>% 
  filter(values < 5) %>% 
  select(-values) %>% 
  mutate(ind = as.character(ind))

long_momb_data <-momb_bymp_cleaned[long_data$ind]
short_momb_data <-momb_bymp_cleaned[short_data$ind]
```

Apply 3-year rolling mean to municipalities with at least 5 years of data
```{r}
smooth_df <-function(df, colnum){
  col <- df[,colnum]
  smoothed <-rollapply(col, width = 3, FUN = mean,  fill=NA)
  df_smoothed <-cbind(df, smoothed)
  df_smoothed
}


long_momb_smoothed <- long_momb_data %>% 
  map(.,~ as.data.frame(.x)) %>% 
  map(., ~ smooth_df(.x, 5)) %>% 
  map(., ~ mutate(.x, momb_score=smoothed)) %>% 
  map(., ~ select(.x, -smoothed))

short_momb_data <-map(short_momb_data,~ as.data.frame(.x))
momb_smoothed <- c(long_momb_smoothed, short_momb_data)
```

# Gapfilling missing years of momb data
## Creating a function to gapfill
The idea is to take an average of the 5 recent years with data. 
Often, there will only be less than 6 years with data - then we just take average of those.
```{r include = FALSE}
#Just another version of the funciton
# gapfill <- function(df) {
#   if (all(is.na(df[, 2]))) {
#     warning("there are no numeric values in the data", call. = FALSE)
#     finaldf <-df
#   } else if (sum(is.na(df[, 2])) <= 5) {
#     cleandat <- drop_na(df)
#     cleandat_sort <- arrange(cleandat, cleandat[, 1])
#     replacement <- mean(tail(cleandat, 3)[, 2], na.rm = T)
#     replaced <- replace_na(df[, 2], replacement)
#     finaldf <- cbind(df[, 1], replaced)
#     colnames(finaldf) <- names(df)
#   } else {
#     cleandat <- drop_na(df)
#     cleandat_sort <- arrange(cleandat, cleandat[, 1])
#     replacement <- mean(tail(cleandat, 5)[, 2], na.rm = T)
#     replaced <- replace_na(df[, 2], replacement)
#     finaldf <- cbind(df[, 1], replaced)
#     colnames(finaldf) <- names(df)
#   }
# 
#   finaldf
# }
```


```{r}
#Gapfilling with the last 5 years of data if there are at least 7 years of data, otherwise - gapfilling with all available years.
gapfill <- function(df, id_col, miss_col) {
     cleandat <- drop_na(df, miss_col)
     finaldf <-df 
    if (all(is.na(df[, miss_col]))) {
    warning("there are no numeric values in the data", call. = FALSE)
    
    } else if (nrow(cleandat) >= 7) {
      cleandat_sort <- arrange(cleandat, cleandat[,id_col])
      replacement <- mean(tail(cleandat, 5)[,miss_col], na.rm = T)
      replaced <- replace_na(df[,miss_col], replacement)
      finaldf <- cbind(df, replaced)
   
    } else {
      cleandat_sort <- arrange(cleandat, cleandat[,id_col])
      replacement <- mean(cleandat[,miss_col], na.rm = T)
      replaced <- replace_na(df[,miss_col], replacement)
      finaldf <- cbind(df, replaced)
 
    }

  finaldf
}

```


## Compiling final dataframe
```{r}
allyears <-data.frame(Year = seq(2005, 2018, by = 1))

gapfilled_prep <- momb_smoothed %>%
  map(., ~ full_join(.x, allyears, by = "Year")) %>% 
  map(., ~ mutate(.x, Komnum_new = rep(Komnum[1], length(Komnum)))) %>% 
  map(., ~ mutate(.x, Kommune_new = rep(Kommune[1], length(Kommune)))) %>% 
  map(., ~ select(.x,-c(Komnum, Kommune))) %>% 
  map(., ~ gapfill(.x, 1,3)) %>% 
  map(., ~ mutate(.x, momb_gapfilled = replaced)) %>% 
  map(., ~ select(.x, -replaced)) 


momb_gapfilled <- do.call("rbind", gapfilled_prep) 
  
momb_gapfilled <- momb_gapfilled %>% 
  mutate(gapfilled = case_when(
    is.na(momb_score)  ~ "y",
    TRUE ~ "n"
  ))
```
Almost half of the municipalities by year observations (411 out of 910) had momb index of 1. This is because momb grade of 3 or 4 is very rare in northern Norway, so in most cases, every municipality has localities with momb grade 1 or 2. 

Save the resultant table:
```{r}
#write_csv(momb_gapfilled, "data/mean_momb_mp_yr_gapfilled.csv")
```

