---
title: "Mom-b examination of aquaculture localities"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r libraries, results="hide"}
library(ggthemes)
library(janitor)
```

# Load mom-b examination data and municipalities names
```{r}
momb_early <-read_excel(file.path(dir_M[2],"Aquaculture/MOMB_2019/MOM_BFiskDira/v_m_akva_lokalitet_mom_b.xlsx"))
momb_latest <-read_excel(file.path(dir_M[2],"Aquaculture/MOMB_2019/B_undersøkelser_Nordland_Troms_Finn_ Fdir_june2019.xls"))
komlist <-read.csv("../../administrative/komlist.csv", sep = ";")
```


# Further preparation of momb data
```{r}
#There is no indicaiton of county and municipality in the new momb data. To find out where the locations were, I will merge this dataset with our previous momb for years 2011-2018.

early_locations <-momb_early %>% 
  select(c("loknr", "fylke", "kommunenr")) %>% 
  mutate(loknr = as.character(loknr)) %>% 
  filter(fylke %in% c("Nordland", "Troms", "Finnmark")) %>% 
  group_by(loknr, kommunenr) %>% 
  summarize(n = n())

momb_latest_prep2 <-  momb_latest_prep %>% 
  mutate(Year = lubridate::year(dato_provetaking)) %>% 
  left_join(early_locations[,c(1,2)], by = "loknr")

map(momb_latest_prep2, ~sum(is.na(.x))) # Still 35 locations did not have info on the municipality where they are located
```



# Preparing final  table: proportion and count of localities with mom b below 2 per municipality
Proportion of localities below mom b 2 in relation to all localities categorized by mom b that year in a municipality.
```{r}
momb_final <- momb_latest_prep2 %>%
  mutate(newindex = case_when(
    momb_index > 2 ~ "1",
    momb_index <= 2 ~ "0"
  )) %>%
  mutate(newindex = as.numeric(newindex)) %>% 
  group_by(Year, kommunenr) %>%
  summarize(
    tot_locs = n(),
    locs_momb_low_count = sum(newindex, na.rm = TRUE),
    locs_momb_low_prop = locs_momb_low_count / tot_locs
  )
```
I see that at maximum there are 12 -17 localities examined per municipality (in the last years). Before 2011 - there are only 1 to 6 localities examined, which confirms insufficient data. I suggest to leave momb estimations aside for now. 


-------------------------------------------------------------------------------
# Extra info (maybe redundant)
## If getting momb info based on earlier dataset  (momb_early) 
```{r}
map_chr(komlist, typeof)#check the type of municp number, usually varies between integer and character

momb_prep_early <- momb_early %>%
  select(-c("prodorgnav", "lokalitet", "loktilstan", "dato_frien", "status_lok", "mu_type", "rapportid", "lat_lokali", "lon_lokali", "url", "vannmiljo_", "fylkeskode", "dato_fri_1", "prodorgn_1")) %>%
  filter(fylke %in% c("Nordland", "Troms", "Finnmark")) %>%
  filter(vannmilj_1 == "SALT") %>%
  rename(loknavn = navn) %>%
  left_join(komlist[, c(1, 2)], by = c("kommunenr" = "Komnum")) %>%
  mutate(kommune = Name) %>%
  select(-"Name") %>%
  mutate(mu_dato = excel_numeric_to_date(as.numeric(mu_dato))) 
```

# Further cleaning mom b 2001-2010 data
```{r}
momb_prep2_early <- momb_prep_early %>% 
  select(-c("TilstandKlasse__1", "vannmiljo","vannmilj_1","plassering", "status_l_1")) %>% 
  clean_names()
```

# Cleaning mom b data for 
```{r}
momb_latest_prep <- momb_latest %>% 
  select(-c("Reference","Dato-ferdig rapport","Rapporterings Dato", "MOM-type")) %>% 
  clean_names()
```








