---
title: "Plotting trends in lice and momb"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---

```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r libraries, results="hide"}
library(ggthemes)
library(ghibli)
```

# Import data
```{r}
lice <-read.csv("data/lice_below_thr_gapfilled.csv") 
momb <-read.csv("data/mean_momb_mp_yr_gapfilled.csv")
komnames_norw <- read_excel(file.path("~/github/aqua_fish_paper/data/raw/komlist_norw.xlsx"))
```


# Lice plots
```{r}
lice_plot <- lice %>% 
  mutate(county = case_when(
    Municip_number >= 1800 & Municip_number < 1900 ~ "Nordland",
    Municip_number >= 1900 & Municip_number < 2000 ~ "Troms",
    Municip_number >= 2000 ~ "Finnmark"
  ))
```

Add names of municipalities with their special Norwegian letters
```{r}
lice_plot_norw <- lice_plot %>% 
  left_join(komnames_norw[c("Komnum", "Norw_name")], by = c("Municip_number" = "Komnum"))
```


```{r}
theme_lice <- function(...) {
  theme_stata(scheme = "s1color") +
    theme(
      legend.position = "none",
      axis.text.x = element_text(
        size = 6, angle = 90, hjust = 1, vjust = 0.5,
        face = "bold"
      ),
      axis.text.y = element_text(size = 7, angle = 0),
      axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
      plot.title = element_text(hjust = 0.5, vjust = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2, "cm"),
      strip.text = element_text(size = 12, face = "bold"),
      strip.background = element_rect(fill = "mistyrose", color ="mediumvioletred"),
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "seashell"),
      panel.grid.major.x = element_line(color = "grey97", size = 0.7),
    )
}
```

## Trend in lice index: proporiton of week a locality is below lice threshold
```{r}
breaks1 <- seq(2004, 2018, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1%%2==0]<- ""


county_names <- list("Finnmark", "Nordland", "Troms")

lice_df <- split(lice_plot, lice_plot$county)

plots_lice <- map2(
  lice_df, 
  names(lice_df),
   ~ ggplot(data = .x, mapping = aes(x = Year, y = score_final)) +
  geom_point(color = "mediumslateblue") +
  geom_smooth() +
  ggtitle(.y)  +
  facet_wrap(Municip ~ .) +
  scale_x_continuous(breaks = breaks1, labels = labels1) +
  labs(y = "Lice index", x = "") +
  theme_lice())

 file_names <- paste0("figs/lice_score_trends", "_", county_names, ".pdf")
 map2(file_names, plots_lice, ggsave)
```


## Same theme but for publication, white backgorund everywhere
```{r}
theme_lice_publ <- function(...) {
  theme_stata(scheme = "s1color") +
    theme(
      legend.position = "none",
      axis.text.x = element_text(
        size = 6, angle = 90, hjust = 1, vjust = 0.5,
        face = "bold"
      ),
      axis.text.y = element_text(size = 7, angle = 0),
      axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
      plot.title = element_text(hjust = 0.5, vjust = 0.5, face = "bold"),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2, "cm"),
      strip.text = element_text(size = 10, face = "bold"),
      strip.background = element_rect(fill = "bisque", color ="bisque"),
      panel.background = element_rect(fill = "white"),
     #panel.border = element_rect(color = "grey97"),
      plot.background = element_blank(),
      panel.grid.major.x = element_line(color = "grey97", size = 0.5),
    )
}
```


```{r}
breaks1 <- seq(2004, 2018, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1%%2==0]<- ""


county_names <- list("Finnmark", "Nordland", "Troms")

lice_df <- split(lice_plot_norw, lice_plot$county)

plots_lice <- map2(
  lice_df, 
  names(lice_df),
   ~ ggplot(data = .x, mapping = aes(x = Year, y = score_final)) +
  geom_point(color = "mediumslateblue", size = 0.6) +
  geom_smooth(color = "darkblue",
              size = 0.4,
              fullrange = FALSE,
              method = "loess",
              span = 0.6) +
  ggtitle(.y)  +
  facet_wrap(Norw_name ~ .) +
  scale_x_continuous(breaks = breaks1, labels = labels1) +
  scale_y_continuous(limits = c(0, 1.2), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  labs(y = "Lice index", x = "") +
  theme_lice_publ())

 file_names <- paste0("figs/lice_score_trends_publ", "_", county_names, ".pdf")
 map2(file_names, plots_lice, ggsave)
```

# Trends in momb  index: the proportion of MAB with momb score 3 and 4

```{r}
momb_plot <- momb %>% 
  mutate(county = case_when(
    Komnum_new >= 1800 & Komnum_new < 1900 ~ "Nordland",
    Komnum_new >= 1900 & Komnum_new < 2000 ~ "Troms",
    Komnum_new >= 2000 ~ "Finnmark"
  ))
```

And adding Norwegian special letters to municipalities's names
```{r}
momb_plot_norw <- momb_plot %>% left_join(komnames_norw[c("Komnum", "Norw_name")], by = c("Komnum_new" = "Komnum"))
```

## Theme for momb colourfull plots
```{r}
theme_momb <- function(...) {
  theme_stata(scheme = "s1color") +
    theme(
      legend.position = "none",
      axis.text.x = element_text(
        size = 6, angle = 90, hjust = 1, vjust = 0.5,
        face = "bold"
      ),
      axis.text.y = element_text(size = 7, angle = 0),
      axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
      plot.title = element_text(hjust = 0.5, vjust = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2, "cm"),
      strip.text = element_text(size = 12, face = "bold"),
      strip.background = element_rect(fill = "moccasin", color ="salmon4"),
      panel.background = element_rect(fill = "white"),
      plot.background = element_rect(fill = "oldlace"),
      panel.grid.major.x = element_line(color = "grey97", size = 0.7),
    )
}
```


Plotting trends in momb index:
```{r}
breaks1 <- seq(2004, 2018, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1%%2==0]<- ""

county_names <- list("Finnmark", "Nordland", "Troms")

momb_df <- split(momb_plot, momb_plot$county)

plots_momb <- map2(
  momb_df, 
  names(momb_df),
   ~ ggplot(data = .x, mapping = aes(x = Year, y = momb_gapfilled)) +
  geom_point(color = "orangered4") +
  geom_smooth(color = "red",
              size = 0.4) +
  ggtitle(.y)  +
  facet_wrap(Kommune_new ~ .) +
  scale_x_continuous(breaks = breaks1, labels = labels1) +
  labs(y = "MOM B index", x = "") +
  theme_momb())

 file_names <- paste0("figs/momb_score_trends", "_", county_names, ".pdf")
 map2(file_names, plots_momb, ggsave)
 
```
## Plotting momb trends for publications
Here I will use same theme as for lice - lice_theme_publ, but will use different colours for geom_smooth.

```{r}
breaks1 <- seq(2004, 2018, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1%%2==0]<- ""

county_names <- list("Finnmark", "Nordland", "Troms")

momb_df <- split(momb_plot_norw, momb_plot$county)

plots_momb <- map2(
  momb_df, 
  names(momb_df),
   ~ ggplot(data = .x, mapping = aes(x = Year, y = momb_gapfilled)) +
  geom_point(color = "orangered4", size = 0.6) +
  geom_smooth(color = "red",
              size = 0.4,
              fullrange = FALSE,
              method = "loess",
              span = 0.6) +
  ggtitle(.y)  +
  facet_wrap(Norw_name ~ .) +
  scale_x_continuous(breaks = breaks1, labels = labels1) +
  scale_y_continuous(limits = c(0, 1.2), breaks = c(0, 0.2, 0.4, 0.6, 0.8,1)) +
  labs(y = "MOM B index", x = "") +
  theme_lice_publ())

 file_names <- paste0("figs/momb_score_trends_publ", "_", county_names, ".pdf")
 map2(file_names, plots_momb, ggsave)
```

