---
title: "Visualizing aquaculture production"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    css: '~/github/nor-prep/prep/templates/style.css'
    toc: true
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/nor-prep/prep/templates/norway_banner.html'
  pdf_document:
    toc: true
---
```{r message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning= FALSE, message = FALSE)
source('~/github/nor-prep/prep/src/common.R')
```

```{r libraries, results="hide"}
library(ggthemes)
library(janitor)
library(DT)
```

# Loading data
```{r}
akva_final <- read.table("data/total_municipality_aquaculture__production_and_efcr.csv", header = TRUE, sep = ",")
tap_0518 <-read.csv(file.path(dir_M[2],"Aquaculture/Production/production_unlocked_concatenated_2005_2018/akva_tap_formatted_2005_2018.csv"))
biomass_0518 <-read.csv(file.path(dir_M[2],"Aquaculture/Production/production_unlocked_concatenated_2005_2018/akva_biomass_2005_2018_formatted.csv")) 
municip <- read.table("../../administrative/komlist.csv", header = TRUE, sep = ";")
```


# Check the overall distribution of aquaculture production
```{r}
mosaic::fav_stats(akva_final$production_final)
```

```{r}
dens_plot <- ggplot(data = akva_final) +
  geom_density(
    mapping = aes(x = production_final),
    fill = "salmon", color = "darkred", alpha = 0.7
  ) +
  ggtitle("Distribution of the total aquaculture production") +
  scale_x_continuous(labels = scales::comma) +
  labs(x = "Total aquaculture production, kg") +
  theme(axis.text = element_text(size = 12)) +
  theme_gdocs()

dens_plot
```

# How did total production vary in the whole region annually?
```{r fig.width=8}

ggplot(akva_final, aes(x = reorder(as.factor(year), production_final, FUN=sum), y = production_final)) +
  geom_boxplot() +
  geom_jitter(aes(color = county), alpha = 0.5, size = 2,
              position = position_jitter(width = 0.3, height = 0)) +
  labs(x = "", y = "Production, kg") +
  theme_gdocs() +
  scale_color_manual(values = c("#ed48ff", "#486dff", "#ffc848"))

```


# Plotting total acquaculture production 
## Creating ggplot theme for total aquaculture production
```{r}
theme_akva <- function(...) {
  theme_stata(scheme = "s1color") +
    theme(
      legend.position = "none",
      axis.text.x = element_text(
        size = 6, angle = 90, hjust = 1, vjust = 0.5,
        face = "bold"
      ),
      axis.text.y = element_text(size = 7, angle = 0),
      axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
      plot.title = element_text(hjust = 0.5, vjust = 0.5),
      axis.ticks.x = element_line(size = 0.5),
      axis.ticks.length = unit(0.2, "cm"),
      strip.text = element_text(size = 12)
    )
}
```


# Plots
```{r}

breaks1 <- seq(1994, 2018, by = 1)
labels1 <- as.character(breaks1)
labels1[!breaks1%%2==0]<- ""


county_names <- list("Finnmark", "Nordland", "Troms")

counties_df <- split(akva_final, akva_final$county)

plots_akva <- map2(
  counties_df, 
  names(counties_df),
   ~ ggplot(data = .x, mapping = aes(x = as.numeric(year), y = as.numeric(production_final))) +
  geom_col(fill = "dodgerblue4") +
  ggtitle(.y)  +
  facet_wrap(municip ~ .) +
  scale_x_continuous(breaks = breaks1, labels = labels1) +
  scale_y_continuous(trans = "log10") +
  labs(y = "Log(Total Mariculture prod (tons))", x = "") +
  theme_akva())

 file_names <- paste0("figs/tot_akva_prod", "_", county_names, ".pdf")
 map2(file_names, plots_akva, ggsave)

```

# Check how fish biomass was changing over a year
An example from Hammerfest municipailty, year 2006
```{r}
biom_change0506 <- data.frame(
  year = rep(c(2005, 2006), each = 12),
  month = rep(c(1:12), 2),
  salmon = c(
    unname(unlist(biomass_0518[1, c(6:17)])),
    unname(unlist(biomass_0518[61, c(6:17)]))
  )
)
```

```{r}
ggplot(data = biom_change0506) +
  geom_line(aes(x = as.factor(month), y = salmon, color = as.factor(year), group = year), size = 2) +
  labs(x = "Month",
       y = "Biomass, kg") +
  scale_color_manual(values = c("darkolivegreen", "darkorchid"), name = "Year") +
      theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    legend.text = element_text(size = 10)
  ) +
  annotate("text", x = 4, y = 1000000, label = "Uttak", color = "orchid", face = "bold") +
  annotate("text", x = 11, y = 1800000, label = "Uttak", color = "darkolivegreen4", face = "bold")
``` 
 
# Which municipalities did not have aquaculture?
```{r}
anti_join(municip, biomass_0518[,c(4,5)], municip, by = c("Komnum"="KOMMUNENR")) %>% 
  arrange(Name) %>% 
  filter(Name != "NonCoast") %>% 
  datatable()
```